### 2021-05-18 paper/paper.Rmd
Make input arguments takes more general data structure.

---
title: "paper"
author: "Mingze Huang, Irina Gaynanova, Christian L. Muller"
date: "5/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2021-05-18 R/GenerateData.R

@@ -1,3 +1,4 @@
+
 #' Construct a correlation matrix
 #' Functions to create autocorrelation matrix (p by p) with parameter rho and block correlation matrix (p by p) using group index (of length p) and (possibly) different parameter rho for each group.
 #' @rdname CorrStructure
@@ -80,7 +81,6 @@ blockcor <- function(blockind, rho){
 #'       \item{Sigma: }{true latent correlation matrix of \code{Z1} and \code{Z2} ((p1+p2) by (p1+p2)).}
 #' }
 #' @export
-#'
 #' @importFrom MASS mvrnorm
 #' @example man/examples/GenerateData_ex.R
 #'

### 2021-05-18 R/KendallTau.R

@@ -16,7 +16,8 @@
 #'
 #' ### vector input
 #' # Data generation (X1: truncated continuous, X2: continuous)
-#' Z <- mvrnorm(n, mu = c(0, 0), Sigma = matrix(c(1, r, r, 1), nrow = 2))
+#' library(MASS)
+#' Z <- MASS::mvrnorm(n, mu = c(0, 0), Sigma = matrix(c(1, r, r, 1), nrow = 2))
 #' X1 <- Z[,1]
 #' X1[Z[,1] < 1] <- 0
 #' X2 <- Z[,2]

### 2021-05-18 R/bridge.R

@@ -45,8 +45,8 @@ R_sol <- function(type1, type2, tau, zratio1, zratio2, method, tol, ratio) {
   if (length(inside) > 0) {
     out[inside] <- r_ml(type1 = type1, type2 = type2, tau = tau[inside], zratio1 = matrix(zratio1[inside, ], nrow=length(inside)), zratio2 = matrix(zratio2[inside, ], nrow = length(inside)))
   }
-  for(ind in outside){
-    out[ind] = r_sol(type1 = type1, type2 = type2, tau = tau[ind], zratio1 = zratio1[ind, ], zratio2 = zratio2[ind, ], tol = tol)
+  if (length(outside) > 0) {
+    out[outside] = sapply(outside, function(x){r_sol(type1 = type1, type2 = type2, tau = tau[x], zratio1 = zratio1[x, ], zratio2 = zratio2[x, ], tol = tol)})
   }
   return(out)
 }

### 2021-05-18 R/estimateR.R

@@ -1,21 +1,29 @@
-
+#'
+#'
+#'
 #' @title Estimate latent correlation matrix
-#'
 #' @description Estimation of latent correlation matrix from observed data of (possibly) mixed types (continuous/biary/truncated continuous) based on the latent Gaussian copula model.
-#'
-#' @aliases estimateR estimateR_mixed
-#' @param X A numeric data matrix (n by p), n is the sample size and p is the number of variables.
-#' @param type A type of variables in \code{X}, must be one of "continuous", "binary" or "trunc".
+#' @rdname estR
+#' @aliases estR
+#' @param X1 A numeric data matrix (n by p1).
+#' @param X2 A numeric data matrix (n by p2).
+#' @param type1 A type of variables in \code{X1}, must be one of "continuous", "binary" or "trunc".
+#' @param type2 A type of variables in \code{X2}, must be one of "continuous", "binary" or "trunc".
 #' @param method The calculation method of latent correlation. Either "original" method or "approx". If \code{method = "approx"}, multilinear approximation method is used, which is much faster than the original method. If \code{method = "original"}, optimization of the bridge inverse function is used. The default is "approx".
 #' @param use.nearPD A logical value indicating whether to use \link[Matrix]{nearPD} or not when the resulting correlation estimator is not positive definite (have at least one negative eigenvalue).
 #' @param nu Shrinkage parameter for correlation matrix, must be between 0 and 1, the default value is 0.01.
 #' @param tol Desired accuracy when calculating the solution of bridge function.
 #' @param verbose If \code{verbose = FALSE}, printing information whether nearPD is used or not is disabled. The default value is FALSE.
 #' @param ratio The maximum ratio of Kendall's tau and boundary to implement multilinear interpolation.
-#' @return \code{estimateR} returns
+#'
+#' @return \code{estR} returns
 #' \itemize{
-#'       \item{type: }{Type of the data matrix \code{X}}
[…]

### 2021-05-18 R/fromKtoR.R

@@ -51,16 +51,21 @@ fromKtoR <- function(K, zratio, type, method, tol, ratio) {
 fromKtoR_mixed <- function(K12, zratio1, zratio2, type1, type2, method, tol, ratio) {
   K12 = as.matrix(K12)
   d1 <- nrow(K12);  d2 <- ncol(K12)
-  if (!(is.null(zratio1))) {zratio1 = apply(zratio1, 2, function(x){rep(x, d2)})}
-  if (!(is.null(zratio2))) {zratio2 = apply(zratio2, 2, function(x){rep(x, each = d1)})}
-  if ((type1 == "continuous" & type2 == "binary") | (type1 == "continuous" & type2 == "trunc") |
-      (type1 == "binary" & type2 == "trunc") | (type1 == "continuous" & type2 == "ternary") |
-      (type1 == "binary" & type2 == "ternary") | (type1 == "trunc" & type2 == "ternary")) {
-    hatR = R_sol(type1 = type2, type2 = type1, tau = c(t(K12)), zratio1 = matrix(zratio2, nrow = length(K12)), zratio2 = matrix(zratio1, nrow = length(K12)), method = method, tol = tol, ratio = ratio)
-    return(matrix(hatR, d1, d2, byrow = TRUE))
-  } else {
-    hatR = R_sol(type1 = type1, type2 = type2, tau = c(K12), zratio1 = matrix(zratio1, nrow = length(K12)), zratio2 = matrix(zratio2, nrow = length(K12)), method = method, tol = tol, ratio = ratio)
+  if (type1 == "continuous" & type2 == "continuous") {
+    hatR <- sin(pi/2 * K12)
     return(matrix(hatR, d1, d2))
+  } else {
+    if (!(is.null(zratio1))) {zratio1 = apply(zratio1, 2, function(x){rep(x, d2)})}
+    if (!(is.null(zratio2))) {zratio2 = apply(zratio2, 2, function(x){rep(x, each = d1)})}
+    if ((type1 == "continuous" & type2 == "binary") | (type1 == "continuous" & type2 == "trunc") |
+       (type1 == "binary" & type2 == "trunc") | (type1 == "continuous" & type2 == "ternary") |
+       (type1 == "binary" & type2 == "ternary") | (type1 == "trunc" & type2 == "ternary")) {
[…]

### 2021-05-18 man/examples/estimateR_ex.R

@@ -1,4 +1,5 @@
 
+library(MASS)
 library(ggplot2)
 ### Data setting
 n <- 1000; p1 <- 15; p2 <- 10 # sample size and dimensions for two datasets.
@@ -27,10 +28,10 @@ simdata <- GenerateData(n=n, trueidx1 = trueidx1, trueidx2 = trueidx2,
 )
 X1 <- simdata$X1; X2 <- simdata$X2; Sigma12_tt <- simdata$Sigma12
 # Estimate latent correlation matrix with original method
-R1_tt_org <- estimateR(X1, type = "binary", method = "original")$R
-R2_tt_org <- estimateR(X2, type = "continuous", method = "original")$R
-R12_tt_org <- estimateR_mixed(X1, X2, type1 = "binary", type2 = "continuous",
-                              method = "original")$R12;
+R1_tt_org <- estR(X1, "binary", method = "original")
+R2_tt_org <- estR(X2, "continuous", method = "original")
+R12_tt_org <- estR(X1, type1 = "binary", X2, type2 = "continuous",
+                              method = "original");
 cp1 <- "exp"; cp2 <- "cube"
 for (tp1 in c("continuous", "binary", "ternary", "trunc")) {
   for (tp2 in c("continuous", "binary", "ternary", "trunc")) {
@@ -55,11 +56,11 @@ for (tp1 in c("continuous", "binary", "ternary", "trunc")) {
                                     c1 = c1, c2 = c2)
       X1 <- simdata$X1; X2 <- simdata$X2; Sigma12 <- simdata$Sigma12
       assign(paste("R1", cp1, cp2, tp1, tp2, md, sep = "_"),
-             estimateR(X1, type = tp1, method = md)$R)
+             estR(X1, tp1, method = md))
       assign(paste("R2", cp1, cp2, tp1, tp2, md, sep = "_"),
-             estimateR(X2, type = tp2, method = md)$R)
+             estR(X2, tp2, method = md))
       assign(paste("R12", cp1, cp2, tp1, tp2, md, sep = "_"),
-             estimateR_mixed(X1, X2, type1 = tp1, type2 = tp2, method = md)$R12)
+             estR(X1 = X1, type1 = tp1, X2 = X2, type2 = tp2, method = md))
       PlotCompare(list(cbind(c(Sigma1), c(get(paste("R1", cp1, cp2, tp1, tp2, md, sep = "_")))),
[…]

### 2021-05-18 paper/paper.Rmd
Create JOSS paper folder

@@
 ---
-title: "paper"
-author: "Mingze Huang, Irina Gaynanova, Christian L. Muller"
-date: "5/18/2021"
-output: html_document
+# Example from https://joss.readthedocs.io/en/latest/submitting.html
+title: 'Latentcor: An R Package for Latent Correlation Estimation'
+tags:
+  - R
+  - Statistics
+  - Latent Correlation
+authors:
+  - name: Mingze Huang
+    orcid: 0000-0003-3919-1564
+    affiliation: "1, 2" # (Multiple affiliations must be quoted)
+  - name: Irina Gaynanova
+    orcid: 0000-0000-0000-0000
+    affiliation: 2
+  - name: Christian L. Muller
+affiliations:
+ - name: Department of Statistics, Texas A& M University
+   index: 1
+ - name: Department of Economics, Texas A& M University
+   index: 2
+ - name: Department of Statistics, University of Munich
+   index: 3
+date: 18 May 2021
+year: 2021
+bibliography: paper.bib
+output: rticles::joss_article
+csl: apa.csl
+journal: JOSS
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
+# Summary
 
-## R Markdown
+# Mathematics
 
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+# Citations
 
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
+# Rendered R Figures
 
-```{r cars}
-summary(cars)
-```
+# Acknowledgements
 
-## Including Plots
-
-You can also embed plots, for example:
-
-```{r pressure, echo=FALSE}
-plot(pressure)
-```
-
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+# References

### 2021-05-18 paper/paper.tex
Create test folder to meet JOSS requirement.

\documentclass[10pt,a4paper,onecolumn]{article}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{tikz}
\usepackage{hyperref}
\hypersetup{colorlinks,breaklinks,
            urlcolor=[rgb]{0.0, 0.5, 1.0},
            linkcolor=[rgb]{0.0, 0.5, 1.0}}
\usepackage{caption}
\usepackage{tcolorbox}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{seqsplit}
\usepackage{fixltx2e} % provides \textsubscript
\usepackage[
  backend=biber,
%  style=alphabetic,
%  citestyle=numeric
]{biblatex}
\bibliography{paper.bib}

% --- Page layout -------------------------------------------------------------
\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.0cm,
            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}

% --- Default font ------------------------------------------------------------
% \renewcommand\familydefault{\sfdefault}

% --- Style -------------------------------------------------------------------
\renewcommand{\bibfont}{\small \sffamily}
\renewcommand{\captionfont}{\small\sffamily}
\renewcommand{\captionlabelfont}{\bfseries}

% --- Section/SubSection/SubSubSection ----------------------------------------
\titleformat{\section}
  {\normalfont\sffamily\Large\bfseries}
  {}{0pt}{}
\titleformat{\subsection}
  {\normalfont\sffamily\large\bfseries}
  {}{0pt}{}
\titleformat{\subsubsection}
  {\normalfont\sffamily\bfseries}
  {}{0pt}{}
\titleformat*{\paragraph}
  {\sffamily\normalsize}

% --- Header / Footer ---------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
%\renewcommand{\headrulewidth}{0.50pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{\hspace{-0.75cm}\includegraphics[width=5.5cm]{C:/Users/sharkmanhmz/Documents/R/R-4.1.0/library/rticles/rmarkdown/templates/joss/resources/JOSS-logo.png}}
\fancyhead[C]{}
\fancyhead[R]{}
\renewcommand{\footrulewidth}{0.25pt}

\fancyfoot[L]{\footnotesize{\sffamily , (2021). Latentcor: An R Package
for Latent Correlation
Estimation. \textit{Journal of Open Source Software}, (), . \href{https://doi.org/}{https://doi.org/}}}

\fancyfoot[R]{\sffamily \thepage}
\makeatletter
\let\ps@plain\ps@fancy
\fancyheadoffset[L]{4.5cm}
\fancyfootoffset[L]{4.5cm}

% --- Macros ---------

\definecolor{linky}{rgb}{0.0, 0.5, 1.0}

\newtcolorbox{repobox}
   {colback=red, colframe=red!75!black,
     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}

\newcommand{\ExternalLink}{%
   \tikz[x=1.2ex, y=1.2ex, baseline=-0.05ex]{%
       \begin{scope}[x=1ex, y=1ex]
           \clip (-0.1,-0.1)
               --++ (-0, 1.2)
               --++ (0.6, 0)
               --++ (0, -0.6)
               --++ (0.6, 0)
               --++ (0, -1);
           \path[draw,
               line width = 0.5,
               rounded corners=0.5]
               (0,0) rectangle (1,1);
       \end{scope}
       \path[draw, line width = 0.5] (0.5, 0.5)
           -- (1, 1);
       \path[draw, line width = 0.5] (0.6, 1)
           -- (1, 1) -- (1, 0.6);
       }
   }

% --- Title / Authors ---------------------------------------------------------
% patch \maketitle so that it doesn't center
\patchcmd{\@maketitle}{center}{flushleft}{}{}
\patchcmd{\@maketitle}{center}{flushleft}{}{}
% patch \maketitle so that the font size for the title is normal
\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
% patch the patch by authblk so that the author block is flush left
\def\maketitle{{%
  \renewenvironment{tabular}[2][]
    {\begin{flushleft}}
    {\end{flushleft}}
  \AB@maketitle}}
\makeatletter
\renewcommand\AB@affilsepx{ \protect\Affilfont}
%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
\makeatother
\renewcommand\Authfont{\sffamily\bfseries}
\renewcommand\Affilfont{\sffamily\small\mdseries}
\setlength{\affilsep}{1em}

\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}

\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}

\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}

\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Latentcor: An R Package for Latent Correlation Estimation},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% Pandoc citation processing

\title{Latentcor: An R Package for Latent Correlation Estimation}

        \author[1, 2]{Mingze Huang}
          \author[2]{Irina Gaynanova}
          \author{Christian L. Muller}

      \affil[1]{Department of Statistics, Texas A\& M University}
      \affil[2]{Department of Economics, Texas A\& M University}
      \affil[3]{Department of Statistics, University of Munich}
  \date{\vspace{-5ex}}

\begin{document}
\maketitle

\marginpar{
  %\hrule
  \sffamily\small

  {\bfseries DOI:} \href{https://doi.org/}{\color{linky}{}}

  \vspace{2mm}

  {\bfseries Software}
  \begin{itemize}
    \setlength\itemsep{0em}
    \item \href{}{\color{linky}{Review}} \ExternalLink
    \item \href{}{\color{linky}{Repository}} \ExternalLink
    \item \href{}{\color{linky}{Archive}} \ExternalLink
  \end{itemize}

  \vspace{2mm}

  {\bfseries Submitted:} \\
  {\bfseries Published:}

  \vspace{2mm}
  {\bfseries License}\\
  Authors of papers retain copyright and release the work under a Creative Commons Attribution 4.0 International License (\href{http://creativecommons.org/licenses/by/4.0/}{\color{linky}{CC-BY}}).
}

\hypertarget{summary}{%
\section{Summary}\label{summary}}

\hypertarget{mathematics}{%
\section{Mathematics}\label{mathematics}}

\hypertarget{citations}{%
\section{Citations}\label{citations}}

\hypertarget{rendered-r-figures}{%
\section{Rendered R Figures}\label{rendered-r-figures}}

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

\hypertarget{references}{%
\section{References}\label{references}}

\end{document}

### 2021-05-18 tests/testthat.R

@@ -1,0 +1,4 @@
+library(testthat)
+library(latentcor)
+
+test_check("latentcor")

### 2021-05-30 paper/paper.Rmd
Adding content on paper

@@
 ---
-# Example from https://joss.readthedocs.io/en/latest/submitting.html
 title: 'Latentcor: An R Package for Latent Correlation Estimation'
@@
 - Latent Correlation
+date: "18 May 2021"
+output:
+  rticles::joss_article: default
+  rticles::joss_article(): default
 authors:
@@
   orcid: 0000-0003-3919-1564
-    affiliation: "1, 2" # (Multiple affiliations must be quoted)
+  affiliation: 1, 2
 - name: Irina Gaynanova
@@
 - name: Christian L. Muller
+year: 2021
+bibliography: paper.bib
 affiliations:
@@
   index: 3
-date: 18 May 2021
-year: 2021
-bibliography: paper.bib
-output: rticles::joss_article
 csl: apa.csl
@@
 
-# Mathematics
+The R package *latentcor* provides estimation for latent correlation with mixed data types (continuous, binary, truncated and ternary).
+Comparing to *MixedCCA*, which estimates latent correlation for canonical correlation analysis, our new package provides a standalone version for latent correlation estimation.
+Also we add new functionality for latent correlation between ternary/continous, ternary/binary, ternary/truncated and ternary/ternary cases.
+Compare to MixedCCA, standalone, new functionality, memory footprint.
 
+# Statement of need
+No package deal with latent correlation across mixed data type.
+
+# Usage
+ Create a table
+ type of variable (reference to papers)
+ Definition for Kendall tau.
+ Describe the theorem states expected value of kendall tau = F(r). refer to table for reference of formula.
+ A sentence to describe two methods: slow original vs interpolation.
+ Table to show memory improvement compare to mixedCCA.
+ Generate data for pairs.
+ One example of original call, one for interpolation call.
 # Citations

### 2021-05-30 paper/paper.tex
Adding content on paper

@@
 
-\hypertarget{mathematics}{%
-\section{Mathematics}\label{mathematics}}
+The R package \emph{latentcor} provides estimation for latent
+correlation with mixed data types (continuous, binary, truncated and
+ternary). Comparing to \emph{MixedCCA}, which estimates latent
+correlation for canonical correlation analysis, our new package provides
+a standalone version for latent correlation estimation. Also we add new
+functionality for latent correlation between ternary/continous,
+ternary/binary, ternary/truncated and ternary/ternary cases. Compare to
+MixedCCA, standalone, new functionality, memory footprint.
 
-\hypertarget{citations}{%
-\section{Citations}\label{citations}}
+\hypertarget{statement-of-need}{%
+\section{Statement of need}\label{statement-of-need}}
+
+No package deal with latent correlation across mixed data type.
+
+\hypertarget{usage}{%
+\section{Usage}\label{usage}}
+
+Create a table type of variable (reference to papers) Definition for
+Kendall tau. Describe the theorem states expected value of kendall tau =
+F(r). refer to table for reference of formula. A sentence to describe
+two methods: slow original vs interpolation. Table to show memory
+improvement compare to mixedCCA. Generate data for pairs. One example of
+original call, one for interpolation call. \# Citations
 

### 2021-05-31 paper/paper.Rmd
Create reference table.

@@
 Also we add new functionality for latent correlation between ternary/continous, ternary/binary, ternary/truncated and ternary/ternary cases.
+
 Compare to MixedCCA, standalone, new functionality, memory footprint.
@@
 # Statement of need
+Currently there is no standalone package dealing with latent correlation for mixed data type like we did in *latentcor*.
+The R package *stats* [@team2013r] have some functionality to calculate different type of correlations (Pearson, Kendall and Spearman).
+The R package *pcaPP* [@croux2013robust] provides a fast calculation for Kendall's \tau.
+The R package *MixedCCA* [@yoon2020sparse] have functionality for latent correlation estimation as an intermediate step for canonical correlation analysis on mixed data.
 No package deal with latent correlation across mixed data type.
@@
 # Usage
- Create a table
- type of variable (reference to papers)
+
+ |Type | continuous | binary | truncated | ternary |
+|-----|----------|
+|continuous | @liu2009nonparanormal | @fan2017high | @yoon2020sparse | @quan2018rank |
+|binary | @fan2017high | @fan2017high | @yoon2020sparse | @quan2018rank |
+|truncated | @yoon2020sparse | @yoon2020sparse | @yoon2020sparse | This paper |
+|ternary | @quan2018rank | @quan2018rank | This paper | @quan2018rank |
+
  Definition for Kendall tau.
@@
  One example of original call, one for interpolation call.
-# Citations
 
@@
 
-# Acknowledgements
-
 # References

### 2021-05-31 paper/paper.tex
Create reference table.

@@
 \urlstyle{same}  % don't use monospace font for urls
+\usepackage{longtable,booktabs}
 \usepackage{graphicx,grffile}
@@
 % Pandoc citation processing
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+% for Pandoc 2.8 to 2.10.1
+\newenvironment{cslreferences}%
+  {}%
+  {\par}
+% For Pandoc 2.11+
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
+  % set entry spacing
+  \ifnum #2 > 0
+  \setlength{\parskip}{#2\baselineskip}
+  \fi
+ }%
+ {}
+\usepackage{calc} % for calculating minipage widths
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
@@
 functionality for latent correlation between ternary/continous,
-ternary/binary, ternary/truncated and ternary/ternary cases. Compare to
-MixedCCA, standalone, new functionality, memory footprint.
+ternary/binary, ternary/truncated and ternary/ternary cases.
+
+Compare to MixedCCA, standalone, new functionality, memory footprint.
 
@@
 
-No package deal with latent correlation across mixed data type.
+Currently there is no standalone package dealing with latent correlation
+for mixed data type like we did in \emph{latentcor}. The R package
+\emph{stats} (Team \& others, 2013) have some functionality to calculate
+different type of correlations (Pearson, Kendall and Spearman). The R
+package \emph{pcaPP} (Croux, Filzmoser, \& Fritz, 2013) provides a fast
+calculation for Kendall's \tau. The R package \emph{MixedCCA} (Yoon,
+Carroll, \& Gaynanova, 2020) have functionality for latent correlation
+estimation as an intermediate step for canonical correlation analysis on
+mixed data. No package deal with latent correlation across mixed data
+type.
 
@@
 
-Create a table type of variable (reference to papers) Definition for
-Kendall tau. Describe the theorem states expected value of kendall tau =
-F(r). refer to table for reference of formula. A sentence to describe
-two methods: slow original vs interpolation. Table to show memory
-improvement compare to mixedCCA. Generate data for pairs. One example of
-original call, one for interpolation call. \# Citations
+Create a table
+
+\begin{longtable}[]{@{}ll@{}}
+\toprule
+Type & continuous \\
+\midrule
+\endhead
+continuous & (\textbf{davis2006reliable?}) \\
+binary & (\textbf{dice1945measures?}) \\
+truncated & (\textbf{dunne2002solutions?}) \\
+ternary & (\textbf{bommert2020adjusted?}) \\
+stabilityIntersectionGreedy & (\textbf{bommert2020adjusted?}) \\
+stabilityIntersectionMBM & (\textbf{bommert2020adjusted?}) \\
+stabilityIntersectionMean & (\textbf{bommert2020adjusted?}) \\
+stabilityJaccard & (\textbf{jaccard1901etude?}) \\
+stabilityKappa & (\textbf{carletta1996assessing?}) \\
+stabilityLustgarten & (\textbf{lustgarten2009measuring?}) \\
+stabilityNogueira & (\textbf{nogueira2018stability?}) \\
+stabilityNovovicova & (\textbf{novovicova2009new?}) \\
+stabilityOchiai & (\textbf{ochiai1957zoogeographic?}) \\
+stabilityPhi & (\textbf{nogueira2016measuring?}) \\
+stabilitySechidis & (\textbf{sechidis2020stability?}) \\
+stabilitySomol & (\textbf{somol2008evaluating?}) \\
+stabilityUnadjusted & (\textbf{bommert2020adjusted?}) \\
+stabilityWald & (\textbf{wald2013stability?}) \\
+stabilityYu & (\textbf{yu2012stable?}) \\
+stabilityZucknick & (\textbf{zucknick2008comparing?}) \\
+\bottomrule
+\end{longtable}
+
+type of variable (reference to papers) Definition for Kendall tau.
+Describe the theorem states expected value of kendall tau = F(r). refer
+to table for reference of formula. A sentence to describe two methods:
+slow original vs interpolation. Table to show memory improvement compare
+to mixedCCA. Generate data for pairs. One example of original call, one
+for interpolation call.
 
@@
 
-\hypertarget{acknowledgements}{%
-\section{Acknowledgements}\label{acknowledgements}}
+\hypertarget{references}{%
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
 
-\hypertarget{references}{%
-\section{References}\label{references}}
+\hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
+\leavevmode\hypertarget{ref-croux2013robust}{}%
+Croux, C., Filzmoser, P., \& Fritz, H. (2013). Robust sparse principal
+component analysis. \emph{Technometrics}, \emph{55}(2), 202--214.
+
+\leavevmode\hypertarget{ref-team2013r}{}%
+Team, R. C., \& others. (2013). R: A language and environment for
+statistical computing.
+
+\leavevmode\hypertarget{ref-yoon2020sparse}{}%
+Yoon, G., Carroll, R. J., \& Gaynanova, I. (2020). Sparse semiparametric
+canonical correlation analysis for data of mixed types.
+\emph{Biometrika}, \emph{107}(3), 609--625.
+
+\end{CSLReferences}
 

### 2021-06-01 paper/paper.Rmd
Definition of tau.
Describe bridge function

@@
 
- Definition for Kendall tau.
+*Definition 1* Fan et al.
+(2017) considered the problem of estimating $\Sigma$ for the latent Gaussian copula model based on Kendall's $\tau$.
+Given the observed data $(X_{1j}, X_{1k}), ..., (X_{nj}, X_{nk})$ for variables $X_{j}$ and $X_{k}$, Kendall's $\tau$ is defined as
+$$
+\hat{\tau}_{jk}=\frac{2}{n(n-1)}\sum_{1\leq i <i'\leq n} sign(X_{ij}-X_{i'j})sign(X_{ik}-X_{i'k})
+$$
+*Theorem 1* Let $W_{1}\in\cal{R}^{p_1}$, $W_{2}\in\cal{R}^{p_2}$, $W_{3}\in\cal{R}^{p_3}$ be such that $W=(W_{1}, W_{2}, W_{3})\sim NPN(0,\Sigma,f)$ with $p=p_{1}+p_{2}+p_{3}$.
+Let $X=(X_{1}, X_{2}, X_{3})\in\cal{R}^{p}$ satisfy $X_{j}=W_{j}$ for $j=1,...,p_{1}$, $X_{j}=I(W_{j}>c_{j})$ for $j=p_{1}+1,...,p_{1}+p_{2}$ and $X_{j}=I(W_{j}>c_{j})W_{j}$ for $j=p_{1}+p_{2}+1,...,p$ with $\Delta_{j}=f(c_{j})$.
+The rank-based estimator of $\Sigma$ based on the observed $n$ realizations of $X$ is the matrix $\hat{R}$ with $\hat{r}_{jj}=1$, $\hat{r}_{jk}=\hat{r}_{kj}=F^{-1}(\hat{\tau}_{jk})$ with block structure
+$$
+\hat{R}=\left(\begin{array}
+F^{-1}_{CC}(\hat{\tau})\hspace{.2in} F^{-1}_{CB}(\hat{\tau})\hspace{.2in} F^{-1}_{CT}(\hat{\tau})\\
+F^{-1}_{BC}(\hat{\tau})\hspace{.2in} F^{-1}_{BB}(\hat{\tau})\hspace{.2in} F^{-1}_{BT}(\hat{\tau})\\
+F^{-1}_{TC}(\hat{\tau})\hspace{.2in} F^{-1}_{TB}(\hat{\tau})\hspace{.2in} F^{-1}_{TT}(\hat{\tau})
+\end{array}\right)
+$$
  Describe the theorem states expected value of kendall tau = F(r). refer to table for reference of formula.

### 2021-06-01 paper/paper.Rmd
Typo

@@
 $$
-\hat{R}=\left(\begin{array}
+\hat{R}=\left(\begin{array}\\
 F^{-1}_{CC}(\hat{\tau})\hspace{.2in} F^{-1}_{CB}(\hat{\tau})\hspace{.2in} F^{-1}_{CT}(\hat{\tau})\\
@@
 $$
- Describe the theorem states expected value of kendall tau = F(r). refer to table for reference of formula.
+refer to table for reference of formula.
  A sentence to describe two methods: slow original vs interpolation.

### 2021-06-01 paper/paper.Rmd
Describe two methods.

@@
 $$
+The original method is taking estimated Kendall's $\hat{\tau}$ and other parameters to calculate latent correlation $\hat{r}$.
+Whereas the approximated method is using multilinear interpolation to approximate latent correlation $\hat{r}$ via pre-calculated grid values [@yoon2021fast].
+
 refer to table for reference of formula.
- A sentence to describe two methods: slow original vs interpolation.
+
  Table to show memory improvement compare to mixedCCA.

### 2021-06-03 paper/paper.tex
Roughly fill up the paper.Rmd.

@@
 \urlstyle{same}  % don't use monospace font for urls
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{248,248,248}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
+\newcommand{\ExtensionTok}[1]{#1}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{#1}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\RegionMarkerTok}[1]{#1}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
 \usepackage{longtable,booktabs}
@@
 
-Create a table
-
 \begin{longtable}[]{@{}ll@{}}
@@
 \endhead
-continuous & (\textbf{davis2006reliable?}) \\
-binary & (\textbf{dice1945measures?}) \\
-truncated & (\textbf{dunne2002solutions?}) \\
-ternary & (\textbf{bommert2020adjusted?}) \\
-stabilityIntersectionGreedy & (\textbf{bommert2020adjusted?}) \\
-stabilityIntersectionMBM & (\textbf{bommert2020adjusted?}) \\
-stabilityIntersectionMean & (\textbf{bommert2020adjusted?}) \\
-stabilityJaccard & (\textbf{jaccard1901etude?}) \\
-stabilityKappa & (\textbf{carletta1996assessing?}) \\
-stabilityLustgarten & (\textbf{lustgarten2009measuring?}) \\
-stabilityNogueira & (\textbf{nogueira2018stability?}) \\
-stabilityNovovicova & (\textbf{novovicova2009new?}) \\
-stabilityOchiai & (\textbf{ochiai1957zoogeographic?}) \\
-stabilityPhi & (\textbf{nogueira2016measuring?}) \\
-stabilitySechidis & (\textbf{sechidis2020stability?}) \\
-stabilitySomol & (\textbf{somol2008evaluating?}) \\
-stabilityUnadjusted & (\textbf{bommert2020adjusted?}) \\
-stabilityWald & (\textbf{wald2013stability?}) \\
-stabilityYu & (\textbf{yu2012stable?}) \\
-stabilityZucknick & (\textbf{zucknick2008comparing?}) \\
+continuous & Liu, Lafferty, \& Wasserman (2009) \\
+binary & Fan, Liu, Ning, \& Zou (2017) \\
+truncated & Yoon, Carroll, \& Gaynanova (2020) \\
+ternary & Quan, Booth, \& Wells (2018) \\
 \bottomrule
@@
 
-type of variable (reference to papers) Definition for Kendall tau.
-Describe the theorem states expected value of kendall tau = F(r). refer
-to table for reference of formula. A sentence to describe two methods:
-slow original vs interpolation. Table to show memory improvement compare
-to mixedCCA. Generate data for pairs. One example of original call, one
-for interpolation call.
+\emph{Definition 1} Fan et al.~(2017) considered the problem of
+estimating \(\Sigma\) for the latent Gaussian copula model based on
+Kendall's \(\tau\). Given the observed data
+\((X_{1j}, X_{1k}), ..., (X_{nj}, X_{nk})\) for variables \(X_{j}\) and
+\(X_{k}\), Kendall's \(\tau\) is defined as \[
+\hat{\tau}_{jk}=\frac{2}{n(n-1)}\sum_{1\leq i <i'\leq n} sign(X_{ij}-X_{i'j})sign(X_{ik}-X_{i'k})
+\] \emph{Theorem 1} Let \(W_{1}\in\cal{R}^{p_1}\),
+\(W_{2}\in\cal{R}^{p_2}\), \(W_{3}\in\cal{R}^{p_3}\),
+\(W_{4}\in\cal{R}^{p_4}\) be such that
+\(W=(W_{1}, W_{2}, W_{3}, W_{4})\sim NPN(0,\Sigma,f)\) with
+\(p=p_{1}+p_{2}+p_{3}+p_{4}\). Let
+\(X=(X_{1}, X_{2}, X_{3}, X_{4})\in\cal{R}^{p}\) satisfy \(X_{j}=W_{j}\)
+for \(j=1,...,p_{1}\), \(X_{j}=I(W_{j}>c_{j})\) for
+\(j=p_{1}+1,...,p_{1}+p_{2}\), \(X_{j}=I(W_{j}>c_{j})W_{j}\) for
+\(j=p_{1}+p_{2}+1,...,p\) and
+\(X_{j}=I(W_{j}>c_{j}^{1})+I(W_{j}>c_{j}^{2})\) with
+\(\Delta_{j}=f(c_{j})\), \(\Delta_{j}^{1}=f(c_{j}^{1})\) and
+\(\Delta_{j}^{2}=f(c_{j}^{2})\). The rank-based estimator of \(\Sigma\)
+based on the observed \(n\) realizations of \(X\) is the matrix
+\(\hat{R}\) with \(\hat{r}_{jj}=1\),
+\(\hat{r}_{jk}=\hat{r}_{kj}=F^{-1}(\hat{\tau}_{jk})\) with block
+structure \[
+\hat{R}=\left(\begin{array}\\
+F^{-1}_{CC}(\hat{\tau})\hspace{.2in} F^{-1}_{CB}(\hat{\tau})\hspace{.2in} F^{-1}_{CT}(\hat{\tau})\hspace{.2in} F^{-1}_{CN}(\hat{\tau})\\
+F^{-1}_{BC}(\hat{\tau})\hspace{.2in} F^{-1}_{BB}(\hat{\tau})\hspace{.2in} F^{-1}_{BT}(\hat{\tau})\hspace{.2in} F^{-1}_{BN}(\hat{\tau})\\
+F^{-1}_{TC}(\hat{\tau})\hspace{.2in} F^{-1}_{TB}(\hat{\tau})\hspace{.2in} F^{-1}_{TT}(\hat{\tau})\hspace{.2in} F^{-1}_{TN}(\hat{\tau})\\
+F^{-1}_{NC}(\hat{\tau})\hspace{.2in} F^{-1}_{NB}(\hat{\tau})\hspace{.2in} F^{-1}_{NT}(\hat{\tau})\hspace{.2in} F^{-1}_{NN}(\hat{\tau})
+\end{array}\right)
+\] The original method is taking estimated Kendall's \(\hat{\tau}\) and
+other parameters to calculate latent correlation \(\hat{r}\). Whereas
+the approximated method is using multilinear interpolation to
+approximate latent correlation \(\hat{r}\) via pre-calculated grid
+values (Yoon, Müller, \& Gaynanova, 2021).
+
+refer to table for reference of formula.
+
+Table to show memory improvement compare to mixedCCA.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{library}\NormalTok{(latentcor)}
+\DocumentationTok{\#\#\# Data setting}
+\NormalTok{n }\OtherTok{\textless{}{-}} \DecValTok{1000}\NormalTok{; p1 }\OtherTok{\textless{}{-}} \DecValTok{1}\NormalTok{; p2 }\OtherTok{\textless{}{-}} \DecValTok{1} \CommentTok{\# sample size and dimensions for two datasets.}
+\NormalTok{maxcancor }\OtherTok{\textless{}{-}} \FloatTok{0.9} \CommentTok{\# true canonical correlation}
+
+\DocumentationTok{\#\#\# Correlation structure within each data set}
+\FunctionTok{set.seed}\NormalTok{(}\DecValTok{0}\NormalTok{)}
+\NormalTok{perm1 }\OtherTok{\textless{}{-}} \FunctionTok{sample}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\NormalTok{p1, }\AttributeTok{size =}\NormalTok{ p1);}
+\NormalTok{Sigma1 }\OtherTok{\textless{}{-}} \FunctionTok{autocor}\NormalTok{(p1, }\FloatTok{0.7}\NormalTok{)[perm1, perm1]}
+\NormalTok{blockind }\OtherTok{\textless{}{-}} \FunctionTok{sample}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\DecValTok{3}\NormalTok{, }\AttributeTok{size =}\NormalTok{ p2, }\AttributeTok{replace =} \ConstantTok{TRUE}\NormalTok{);}
+\NormalTok{Sigma2 }\OtherTok{\textless{}{-}} \FunctionTok{blockcor}\NormalTok{(blockind, }\FloatTok{0.7}\NormalTok{)}
+\NormalTok{mu }\OtherTok{\textless{}{-}} \FunctionTok{rbinom}\NormalTok{(p1}\SpecialCharTok{+}\NormalTok{p2, }\DecValTok{1}\NormalTok{, }\FloatTok{0.5}\NormalTok{)}
+
+\DocumentationTok{\#\#\# true variable indices for each dataset}
+\NormalTok{trueidx1 }\OtherTok{\textless{}{-}} \DecValTok{1}
+\NormalTok{trueidx2 }\OtherTok{\textless{}{-}} \DecValTok{1}
+
+\CommentTok{\# Data generation}
+\NormalTok{simdata }\OtherTok{\textless{}{-}} \FunctionTok{GenerateData}\NormalTok{(}\AttributeTok{n=}\NormalTok{n, }\AttributeTok{trueidx1 =}\NormalTok{ trueidx1, }\AttributeTok{trueidx2 =}\NormalTok{ trueidx2,}
+                        \AttributeTok{maxcancor =}\NormalTok{ maxcancor,}
+                        \AttributeTok{Sigma1 =}\NormalTok{ Sigma1, }\AttributeTok{Sigma2 =}\NormalTok{ Sigma2,}
+                        \AttributeTok{copula1 =} \StringTok{"exp"}\NormalTok{, }\AttributeTok{copula2 =} \StringTok{"cube"}\NormalTok{,}
+                        \AttributeTok{muZ =}\NormalTok{ mu,}
+                        \AttributeTok{type1 =} \StringTok{"binary"}\NormalTok{, }\AttributeTok{type2 =} \StringTok{"continuous"}\NormalTok{,}
+                        \AttributeTok{c1 =} \FunctionTok{rep}\NormalTok{(}\DecValTok{1}\NormalTok{, p1), }\AttributeTok{c2 =}  \ConstantTok{NULL}
+\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## Warning in GenerateData(n = n, trueidx1 = trueidx1, trueidx2 = trueidx2, : Same
+## threshold is applied to the all variables in the first set.
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{X1 }\OtherTok{\textless{}{-}}\NormalTok{ simdata}\SpecialCharTok{$}\NormalTok{X1; X2 }\OtherTok{\textless{}{-}}\NormalTok{ simdata}\SpecialCharTok{$}\NormalTok{X2; Sigma12\_tt }\OtherTok{\textless{}{-}}\NormalTok{ simdata}\SpecialCharTok{$}\NormalTok{Sigma12}
+\CommentTok{\# Estimate latent correlation matrix with original method}
+\NormalTok{R1\_tt\_org }\OtherTok{\textless{}{-}} \FunctionTok{estR}\NormalTok{(X1, }\StringTok{"binary"}\NormalTok{, }\AttributeTok{method =} \StringTok{"original"}\NormalTok{)}
+\NormalTok{R2\_tt\_org }\OtherTok{\textless{}{-}} \FunctionTok{estR}\NormalTok{(X2, }\StringTok{"continuous"}\NormalTok{, }\AttributeTok{method =} \StringTok{"original"}\NormalTok{)}
+\NormalTok{R12\_tt\_org }\OtherTok{\textless{}{-}} \FunctionTok{estR}\NormalTok{(X1, }\AttributeTok{type1 =} \StringTok{"binary"}\NormalTok{, X2, }\AttributeTok{type2 =} \StringTok{"continuous"}\NormalTok{,}
+                              \AttributeTok{method =} \StringTok{"original"}\NormalTok{)}\SpecialCharTok{$}\NormalTok{R12}
+\CommentTok{\# Estimate latent correlation matrix with original method}
+\NormalTok{R1\_tt\_org }\OtherTok{\textless{}{-}} \FunctionTok{estR}\NormalTok{(X1, }\StringTok{"binary"}\NormalTok{, }\AttributeTok{method =} \StringTok{"approx"}\NormalTok{)}
+\NormalTok{R2\_tt\_org }\OtherTok{\textless{}{-}} \FunctionTok{estR}\NormalTok{(X2, }\StringTok{"continuous"}\NormalTok{, }\AttributeTok{method =} \StringTok{"approx"}\NormalTok{)}
+\NormalTok{R12\_tt\_org }\OtherTok{\textless{}{-}} \FunctionTok{estR}\NormalTok{(X1, }\AttributeTok{type1 =} \StringTok{"binary"}\NormalTok{, X2, }\AttributeTok{type2 =} \StringTok{"continuous"}\NormalTok{,}
+                              \AttributeTok{method =} \StringTok{"approx"}\NormalTok{)}\SpecialCharTok{$}\NormalTok{R12}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
+\leavevmode\hypertarget{ref-fan2017high}{}%
+Fan, J., Liu, H., Ning, Y., \& Zou, H. (2017). High dimensional
+semiparametric latent graphical model for mixed data. \emph{Journal of
+the Royal Statistical Society. Series B: Statistical Methodology},
+\emph{79}(2), 405--421.
+
+\leavevmode\hypertarget{ref-liu2009nonparanormal}{}%
+Liu, H., Lafferty, J., \& Wasserman, L. (2009). The nonparanormal:
+Semiparametric estimation of high dimensional undirected graphs.
+\emph{Journal of Machine Learning Research}, \emph{10}(10).
+
+\leavevmode\hypertarget{ref-quan2018rank}{}%
+Quan, X., Booth, J. G., \& Wells, M. T. (2018). Rank-based approach for
+estimating correlations in mixed ordinal data. \emph{arXiv preprint
+arXiv:1809.06255}.
+
 \leavevmode\hypertarget{ref-team2013r}{}%
@@
 
+\leavevmode\hypertarget{ref-yoon2021fast}{}%
+Yoon, G., Müller, C. L., \& Gaynanova, I. (2021). Fast computation of
+latent correlations. \emph{Journal of Computational and Graphical
+Statistics}, 1--8.
+
 \end{CSLReferences}

### 2021-06-03 R/GenerateData.R

@@ -117,8 +117,8 @@ GenerateData <- function(n, trueidx1, trueidx2, Sigma1, Sigma2, maxcancor,
   }
   dat <- MASS::mvrnorm(n, mu = muZ, Sigma = JSigma) # generate a data matrix of size: n by length(muZ). length(muZ) should match with ncol(JSigma)=nrow(JSigma).
 
-  Z1 <- dat[, 1:p1]
-  Z2 <- dat[, (p1+1):p]
+  Z1 <- as.matrix(dat[, 1:p1])
+  Z2 <- as.matrix(dat[, (p1+1):p])
 
   # Three different types of copula
   if(copula1 != "no"){
@@ -154,9 +154,6 @@ GenerateData <- function(n, trueidx1, trueidx2, Sigma1, Sigma2, maxcancor,
   } else if (type1 == "ternary") {
     X1 <- ifelse(Z1 >= matrix(apply(c1, 2, max), nrow = nrow(Z1), ncol = ncol(Z1), byrow = T), 2, 1)
     X1[Z1 <= matrix(apply(c1, 2, min), nrow = nrow(Z1), ncol = ncol(Z1), byrow = T)] = 0
-  } else if (type1 == "dtrunc") {
-    X1 <- ifelse(Z1 >= matrix(apply(c1, 2, max), nrow = nrow(Z1), ncol = ncol(Z1), byrow = T), 1, Z1)
-    X1[Z1 <= matrix(apply(c1, 2, min), nrow = nrow(Z1), ncol = ncol(Z1), byrow = T)] = 0
   }
 
   if(type2 == "continuous") {
@@ -167,9 +164,6 @@ GenerateData <- function(n, trueidx1, trueidx2, Sigma1, Sigma2, maxcancor,
     X2 <- ifelse(Z2 > matrix(c2, nrow = nrow(Z2), ncol = ncol(Z2), byrow = T), 1, 0)
   } else if (type2 == "ternary") {
     X2 <- ifelse(Z2 >= matrix(apply(c2, 2, max), nrow = nrow(Z2), ncol = ncol(Z2), byrow = T), 2, 1)
-    X2[Z2 <= matrix(apply(c2, 2, min), nrow = nrow(Z2), ncol = ncol(Z2), byrow = T)] = 0
-  } else if (type2 == "dtrunc") {
[…]

### 2021-06-03 R/bridge.R

@@ -30,8 +30,6 @@ zratio = function(X, type) {
     }
   } else if (type == "ternary") {
     zratio <- cbind(colMeans(X == 0), 1 - colMeans(X == 2))
-    # } else if (type == "dtrunc") {
-    #   zratio <- cbind(colMeans(X == 0), 1 - colMeans(X == 1))
   }
   return(zratio)
 }
@@ -81,8 +79,6 @@ bridge <- function(type1, type2, r, zratio1, zratio2) {
     out <- bridgeF_nt(r = r, zratio1 = zratio1, zratio2 = zratio2)
   } else if (type1 == "ternary" & type2 == "ternary") {
     out <- bridgeF_nn(r = r, zratio1 = zratio1, zratio2 = zratio2)
-  } else if (type1 == "dtrunc" & type2 == "continuous") {
-    out <- bridgeF_dc(r = r, zratio1 = zratio1)
   } else {
     stop("Unrecognized type of variables. Should be one of continuous, binary, trunc or ternary.")
   }

### 2021-06-03 R/estimateR.R

@@ -36,12 +36,11 @@
 #' @example man/examples/estimateR_ex.R
 #'
 
-estR <- function(X1, type1, X2 = NULL, type2 = NULL, Xlist = NULL, types = NULL, method = "approx", use.nearPD = TRUE, nu = 0.01, tol = 1e-6, verbose = FALSE, ratio = 0.9){
+estR <- function(X1, type1, X2 = NULL, type2 = NULL, method = "approx", use.nearPD = TRUE, nu = 0.01, tol = 1e-6, verbose = FALSE, ratio = 0.9){
   # shrinkage method
   if(nu < 0 | nu > 1){
     stop("nu must be be between 0 and 1.")
   }
-  # if (!(is.null(X1))) {
   X1 <- as.matrix(X1); p1 <- ncol(X1)
   if (length(colnames(X1)) == p1) {
     name1 = colnames(X1)
@@ -73,11 +72,7 @@ estR <- function(X1, type1, X2 = NULL, type2 = NULL, Xlist = NULL, types = NULL,
     R1 <- R.final[1:p1, 1:p1]
     R2 <- R.final[(p1 + 1):(p1 + p2), (p1 + 1):(p1 + p2)]
     R12 <- R.final[1:p1, (p1 + 1):(p1 + p2)]
-    # return(list(type = c(type1, type2), R1 = R1, R2 = R2, R12 = R12, R = R.final))
-    return(R12)
+    return(list(type = c(type1, type2), R1 = R1, R2 = R2, R12 = R12, R = R.final))
   }
-  # } else {
-  #
-  # }
 }
 

### 2021-06-03 man/examples/estimateR_ex.R

@@ -1,5 +1,4 @@
 
-library(MASS)
 library(ggplot2)
 ### Data setting
 n <- 1000; p1 <- 15; p2 <- 10 # sample size and dimensions for two datasets.
@@ -31,7 +30,7 @@ X1 <- simdata$X1; X2 <- simdata$X2; Sigma12_tt <- simdata$Sigma12
 R1_tt_org <- estR(X1, "binary", method = "original")
 R2_tt_org <- estR(X2, "continuous", method = "original")
 R12_tt_org <- estR(X1, type1 = "binary", X2, type2 = "continuous",
-                              method = "original");
+                              method = "original")$R12
 cp1 <- "exp"; cp2 <- "cube"
 for (tp1 in c("continuous", "binary", "ternary", "trunc")) {
   for (tp2 in c("continuous", "binary", "ternary", "trunc")) {
@@ -60,7 +59,7 @@ for (tp1 in c("continuous", "binary", "ternary", "trunc")) {
       assign(paste("R2", cp1, cp2, tp1, tp2, md, sep = "_"),
              estR(X2, tp2, method = md))
       assign(paste("R12", cp1, cp2, tp1, tp2, md, sep = "_"),
-             estR(X1 = X1, type1 = tp1, X2 = X2, type2 = tp2, method = md))
+             estR(X1 = X1, type1 = tp1, X2 = X2, type2 = tp2, method = md)$R12)
       PlotCompare(list(cbind(c(Sigma1), c(get(paste("R1", cp1, cp2, tp1, tp2, md, sep = "_")))),
                        cbind(c(Sigma2), c(get(paste("R2", cp1, cp2, tp1, tp2, md, sep = "_")))),
                        cbind(c(Sigma12), c(get(paste("R12", cp1, cp2, tp1, tp2, md, sep = "_"))))

### 2021-06-03 tests/testthat/test_KendallTau.R

@@ -1,0 +1,12 @@
+context("Test KendallTau")
+library(latentcor)
+
+test_that("Kendall' tau is symmetric for two inputs", {
+  X1 = rnorm(10); X2 = rnorm(10)
+  expect_equal(KendallTau(X1, X2), KendallTau(X2, X1))
+})
+#> Test passed 😸
+
+#> Test passed 🌈
+
+#> Test passed 🎉

### 2021-06-03 paper/paper.Rmd
Fixed

@@
 
-Compare to MixedCCA, standalone, new functionality, memory footprint.
+Compare to MixedCCA, memory footprint.
 
@@
 $$
-*Theorem 1* Let $W_{1}\in\cal{R}^{p_1}$, $W_{2}\in\cal{R}^{p_2}$, $W_{3}\in\cal{R}^{p_3}$ be such that $W=(W_{1}, W_{2}, W_{3})\sim NPN(0,\Sigma,f)$ with $p=p_{1}+p_{2}+p_{3}$.
-Let $X=(X_{1}, X_{2}, X_{3})\in\cal{R}^{p}$ satisfy $X_{j}=W_{j}$ for $j=1,...,p_{1}$, $X_{j}=I(W_{j}>c_{j})$ for $j=p_{1}+1,...,p_{1}+p_{2}$ and $X_{j}=I(W_{j}>c_{j})W_{j}$ for $j=p_{1}+p_{2}+1,...,p$ with $\Delta_{j}=f(c_{j})$.
+*Theorem 1* Let $W_{1}\in\cal{R}^{p_1}$, $W_{2}\in\cal{R}^{p_2}$, $W_{3}\in\cal{R}^{p_3}$, $W_{4}\in\cal{R}^{p_4}$ be such that $W=(W_{1}, W_{2}, W_{3}, W_{4})\sim NPN(0,\Sigma,f)$ with $p=p_{1}+p_{2}+p_{3}+p_{4}$.
+Let $X=(X_{1}, X_{2}, X_{3}, X_{4})\in\cal{R}^{p}$ satisfy $X_{j}=W_{j}$ for $j=1,...,p_{1}$, $X_{j}=I(W_{j}>c_{j})$ for $j=p_{1}+1,...,p_{1}+p_{2}$, $X_{j}=I(W_{j}>c_{j})W_{j}$ for $j=p_{1}+p_{2}+1,...,p$ and $X_{j}=I(W_{j}>c_{j}^{1})+I(W_{j}>c_{j}^{2})$ with $\Delta_{j}=f(c_{j})$, $\Delta_{j}^{1}=f(c_{j}^{1})$ and $\Delta_{j}^{2}=f(c_{j}^{2})$.
 The rank-based estimator of $\Sigma$ based on the observed $n$ realizations of $X$ is the matrix $\hat{R}$ with $\hat{r}_{jj}=1$, $\hat{r}_{jk}=\hat{r}_{kj}=F^{-1}(\hat{\tau}_{jk})$ with block structure
@@
 \hat{R}=\left(\begin{array}\\
-F^{-1}_{CC}(\hat{\tau})\hspace{.2in} F^{-1}_{CB}(\hat{\tau})\hspace{.2in} F^{-1}_{CT}(\hat{\tau})\\
-F^{-1}_{BC}(\hat{\tau})\hspace{.2in} F^{-1}_{BB}(\hat{\tau})\hspace{.2in} F^{-1}_{BT}(\hat{\tau})\\
-F^{-1}_{TC}(\hat{\tau})\hspace{.2in} F^{-1}_{TB}(\hat{\tau})\hspace{.2in} F^{-1}_{TT}(\hat{\tau})
+F^{-1}_{CC}(\hat{\tau})\hspace{.2in} F^{-1}_{CB}(\hat{\tau})\hspace{.2in} F^{-1}_{CT}(\hat{\tau})\hspace{.2in} F^{-1}_{CN}(\hat{\tau})\\
+F^{-1}_{BC}(\hat{\tau})\hspace{.2in} F^{-1}_{BB}(\hat{\tau})\hspace{.2in} F^{-1}_{BT}(\hat{\tau})\hspace{.2in} F^{-1}_{BN}(\hat{\tau})\\
+F^{-1}_{TC}(\hat{\tau})\hspace{.2in} F^{-1}_{TB}(\hat{\tau})\hspace{.2in} F^{-1}_{TT}(\hat{\tau})\hspace{.2in} F^{-1}_{TN}(\hat{\tau})\\
+F^{-1}_{NC}(\hat{\tau})\hspace{.2in} F^{-1}_{NB}(\hat{\tau})\hspace{.2in} F^{-1}_{NT}(\hat{\tau})\hspace{.2in} F^{-1}_{NN}(\hat{\tau})
 \end{array}\right)
@@
  Table to show memory improvement compare to mixedCCA.
- Generate data for pairs.
- One example of original call, one for interpolation call.
+```{r}
+library(latentcor)
+### Data setting
+n <- 1000; p1 <- 1; p2 <- 1 # sample size and dimensions for two datasets.
+maxcancor <- 0.9 # true canonical correlation
+
+### Correlation structure within each data set
+set.seed(0)
+perm1 <- sample(1:p1, size = p1);
+Sigma1 <- autocor(p1, 0.7)[perm1, perm1]
+blockind <- sample(1:3, size = p2, replace = TRUE);
+Sigma2 <- blockcor(blockind, 0.7)
+mu <- rbinom(p1+p2, 1, 0.5)
+
+### true variable indices for each dataset
+trueidx1 <- 1
+trueidx2 <- 1
+
+# Data generation
+simdata <- GenerateData(n=n, trueidx1 = trueidx1, trueidx2 = trueidx2,
+                        maxcancor = maxcancor,
+                        Sigma1 = Sigma1, Sigma2 = Sigma2,
+                        copula1 = "exp", copula2 = "cube",
+                        muZ = mu,
+                        type1 = "binary", type2 = "continuous",
+                        c1 = rep(1, p1), c2 =  NULL
+)
+X1 <- simdata$X1; X2 <- simdata$X2; Sigma12_tt <- simdata$Sigma12
+# Estimate latent correlation matrix with original method
+R1_tt_org <- estR(X1, "binary", method = "original")
+R2_tt_org <- estR(X2, "continuous", method = "original")
+R12_tt_org <- estR(X1, type1 = "binary", X2, type2 = "continuous",
+                              method = "original")$R12
+# Estimate latent correlation matrix with original method
+R1_tt_org <- estR(X1, "binary", method = "approx")
+R2_tt_org <- estR(X2, "continuous", method = "approx")
+R12_tt_org <- estR(X1, type1 = "binary", X2, type2 = "continuous",
+                              method = "approx")$R12
+```
 

### 2021-06-03 paper/paper.Rmd
Typo.

@@
 # Estimate latent correlation matrix with original method
-R1_tt_org <- estR(X1, "binary", method = "original")
-R2_tt_org <- estR(X2, "continuous", method = "original")
-R12_tt_org <- estR(X1, type1 = "binary", X2, type2 = "continuous",
+R1_b_org <- estR(X1, "binary", method = "original")
+R2_c_org <- estR(X2, "continuous", method = "original")
+R12_bc_org <- estR(X1, type1 = "binary", X2, type2 = "continuous",
                               method = "original")$R12
 # Estimate latent correlation matrix with original method
-R1_tt_org <- estR(X1, "binary", method = "approx")
-R2_tt_org <- estR(X2, "continuous", method = "approx")
-R12_tt_org <- estR(X1, type1 = "binary", X2, type2 = "continuous",
+R1_b_approx <- estR(X1, "binary", method = "approx")
+R2_c_approx <- estR(X2, "continuous", method = "approx")
+R12_bc_approx <- estR(X1, type1 = "binary", X2, type2 = "continuous",
                               method = "approx")$R12

### 2021-06-03 paper/paper.Rmd
typo

@@
   rticles::joss_article: default
-  rticles::joss_article(): default
 authors:

### 2021-06-03 paper/paper.tex
typo

@@
 \CommentTok{\# Estimate latent correlation matrix with original method}
-\NormalTok{R1\_tt\_org }\OtherTok{\textless{}{-}} \FunctionTok{estR}\NormalTok{(X1, }\StringTok{"binary"}\NormalTok{, }\AttributeTok{method =} \StringTok{"original"}\NormalTok{)}
-\NormalTok{R2\_tt\_org }\OtherTok{\textless{}{-}} \FunctionTok{estR}\NormalTok{(X2, }\StringTok{"continuous"}\NormalTok{, }\AttributeTok{method =} \StringTok{"original"}\NormalTok{)}
-\NormalTok{R12\_tt\_org }\OtherTok{\textless{}{-}} \FunctionTok{estR}\NormalTok{(X1, }\AttributeTok{type1 =} \StringTok{"binary"}\NormalTok{, X2, }\AttributeTok{type2 =} \StringTok{"continuous"}\NormalTok{,}
+\NormalTok{R1\_b\_org }\OtherTok{\textless{}{-}} \FunctionTok{estR}\NormalTok{(X1, }\StringTok{"binary"}\NormalTok{, }\AttributeTok{method =} \StringTok{"original"}\NormalTok{)}
+\NormalTok{R2\_c\_org }\OtherTok{\textless{}{-}} \FunctionTok{estR}\NormalTok{(X2, }\StringTok{"continuous"}\NormalTok{, }\AttributeTok{method =} \StringTok{"original"}\NormalTok{)}
+\NormalTok{R12\_bc\_org }\OtherTok{\textless{}{-}} \FunctionTok{estR}\NormalTok{(X1, }\AttributeTok{type1 =} \StringTok{"binary"}\NormalTok{, X2, }\AttributeTok{type2 =} \StringTok{"continuous"}\NormalTok{,}
                               \AttributeTok{method =} \StringTok{"original"}\NormalTok{)}\SpecialCharTok{$}\NormalTok{R12}
 \CommentTok{\# Estimate latent correlation matrix with original method}
-\NormalTok{R1\_tt\_org }\OtherTok{\textless{}{-}} \FunctionTok{estR}\NormalTok{(X1, }\StringTok{"binary"}\NormalTok{, }\AttributeTok{method =} \StringTok{"approx"}\NormalTok{)}
-\NormalTok{R2\_tt\_org }\OtherTok{\textless{}{-}} \FunctionTok{estR}\NormalTok{(X2, }\StringTok{"continuous"}\NormalTok{, }\AttributeTok{method =} \StringTok{"approx"}\NormalTok{)}
-\NormalTok{R12\_tt\_org }\OtherTok{\textless{}{-}} \FunctionTok{estR}\NormalTok{(X1, }\AttributeTok{type1 =} \StringTok{"binary"}\NormalTok{, X2, }\AttributeTok{type2 =} \StringTok{"continuous"}\NormalTok{,}
+\NormalTok{R1\_b\_approx }\OtherTok{\textless{}{-}} \FunctionTok{estR}\NormalTok{(X1, }\StringTok{"binary"}\NormalTok{, }\AttributeTok{method =} \StringTok{"approx"}\NormalTok{)}
+\NormalTok{R2\_c\_approx }\OtherTok{\textless{}{-}} \FunctionTok{estR}\NormalTok{(X2, }\StringTok{"continuous"}\NormalTok{, }\AttributeTok{method =} \StringTok{"approx"}\NormalTok{)}
+\NormalTok{R12\_bc\_approx }\OtherTok{\textless{}{-}} \FunctionTok{estR}\NormalTok{(X1, }\AttributeTok{type1 =} \StringTok{"binary"}\NormalTok{, X2, }\AttributeTok{type2 =} \StringTok{"continuous"}\NormalTok{,}
                               \AttributeTok{method =} \StringTok{"approx"}\NormalTok{)}\SpecialCharTok{$}\NormalTok{R12}

### 2021-06-04 paper/paper.Rmd
commit

@@
 Comparing to *MixedCCA*, which estimates latent correlation for canonical correlation analysis, our new package provides a standalone version for latent correlation estimation.
-Also we add new functionality for latent correlation between ternary/continous, ternary/binary, ternary/truncated and ternary/ternary cases.
+Also we add new functionality for latent correlation between ternary/continuous, ternary/binary, ternary/truncated and ternary/ternary cases.
 
@@
 The R package *stats* [@team2013r] have some functionality to calculate different type of correlations (Pearson, Kendall and Spearman).
-The R package *pcaPP* [@croux2013robust] provides a fast calculation for Kendall's \tau.
+The R package *pcaPP* [@croux2013robust] provides a fast calculation for Kendall's $\tau$.
 The R package *MixedCCA* [@yoon2020sparse] have functionality for latent correlation estimation as an intermediate step for canonical correlation analysis on mixed data.

### 2021-06-04 R/GenData.R
Simplify data generation

@@ -1,9 +1,10 @@
 
 #' Construct a correlation matrix
+#' @rdname Sigma
 #' Functions to create autocorrelation matrix (p by p) with parameter rho and block correlation matrix (p by p) using group index (of length p) and (possibly) different parameter rho for each group.
-#' @rdname CorrStructure
 #' @param p Specified matrix dimension.
 #' @param rho Correlation value(s), must be between -0.99 and 0.99. Should be a scalar for \code{autocor}, and either a scalar or a vector of the same length as the maximal \code{blockind} K for \code{blockcor}.
+#' @return  Correlation matrix
 #' @export
 autocor <- function(p, rho){
   if (abs(rho) > 0.99){ stop("correlation rho must be between -0.99 and 0.99.") }
@@ -13,9 +14,10 @@ autocor <- function(p, rho){
 
 
 #' Construct a correlation matrix
-#' @rdname CorrStructure
+#' @rdname Sigma
 #' @param blockind Block index 1,\dots, K for a positive integer K specifying which variable belongs to which block, the matrix dimension is equal to \code{length(blockind)}.
 #' @param rho Correlation value(s), must be between -0.99 and 0.99. Should be a scalar for \code{autocor}, and either a scalar or a vector of the same length as the maximal \code{blockind} K for \code{blockcor}.
+#' @return Correlation matrix \code{Sigma}
 #' @examples
 #' # For p = 8,
 #' # auto correlation matrix
@@ -48,74 +50,71 @@ blockcor <- function(blockind, rho){
   return(Sigma)
 }
 
-#' Mixed type simulation data generator for sparse CCA
-#'
-#' \code{GenerateData} is used to generate two sets of data of mixed types for sparse CCA under the Gaussian copula model.
-#'
+#' Mixed type simulation data generator
+#' \code{GenData} is used to generate two sets of data of mixed types for sparse CCA under the Gaussian copula model.
 #' @param n Sample size
-#' @param trueidx1 True canonical direction of length p1 for \code{X1}. It will be automatically normalized such that \eqn{w_1^T \Sigma_1 w_1 = 1}.
[…]

### 2021-06-04 man/examples/GenerateData_ex.R

@@ -1,153 +1,0 @@
-### Simple example
-
-# Data setting
-n <- 100; p1 <- 15; p2 <- 10 # sample size and dimensions for two datasets.
-maxcancor <- 0.9 # true canonical correlation
-
-# Correlation structure within each data set
-set.seed(0)
-perm1 <- sample(1:p1, size = p1);
-Sigma1 <- autocor(p1, 0.7)[perm1, perm1]
-blockind <- sample(1:3, size = p2, replace = TRUE);
-Sigma2 <- blockcor(blockind, 0.7)
-mu <- rbinom(p1+p2, 1, 0.5)
-
-# true variable indices for each dataset
-trueidx1 <- c(rep(1, 3), rep(0, p1-3))
-trueidx2 <- c(rep(1, 2), rep(0, p2-2))
-
-# Data generation
-simdata <- GenerateData(n=n, trueidx1 = trueidx1, trueidx2 = trueidx2, maxcancor = maxcancor,
-                        Sigma1 = Sigma1, Sigma2 = Sigma2,
-                        copula1 = "exp", copula2 = "cube",
-                        muZ = mu,
-                        type1 = "trunc", type2 = "trunc",
-                        c1 = rep(1, p1), c2 =  rep(0, p2)
-)
-X1 <- simdata$X1
-X2 <- simdata$X2
-
-# Data generation
-simdata <- GenerateData(n=n, trueidx1 = trueidx1, trueidx2 = trueidx2, maxcancor = maxcancor,
-                        Sigma1 = Sigma1, Sigma2 = Sigma2,
-                        copula1 = "exp", copula2 = "cube",
-                        muZ = mu,
-                        type1 = "trunc", type2 = "continuous",
-                        c1 = rep(1, p1), c2 =  rep(0, p2)
-)
-X1 <- simdata$X1
-X2 <- simdata$X2
-
-# Data generation
-simdata <- GenerateData(n=n, trueidx1 = trueidx1, trueidx2 = trueidx2, maxcancor = maxcancor,
-                        Sigma1 = Sigma1, Sigma2 = Sigma2,
-                        copula1 = "exp", copula2 = "cube",
-                        muZ = mu,
-                        type1 = "binary", type2 = "binary",
-                        c1 = rep(1, p1), c2 =  rep(0, p2)
-)
-X1 <- simdata$X1
-X2 <- simdata$X2
-
-# Data generation
[…]

### 2021-06-05 R/GenData.R
Simplify data generation

@@ -1,10 +1,10 @@
 
 #' Construct a correlation matrix
-#' @rdname Sigma
 #' Functions to create autocorrelation matrix (p by p) with parameter rho and block correlation matrix (p by p) using group index (of length p) and (possibly) different parameter rho for each group.
+#' @rdname Sigma
 #' @param p Specified matrix dimension.
 #' @param rho Correlation value(s), must be between -0.99 and 0.99. Should be a scalar for \code{autocor}, and either a scalar or a vector of the same length as the maximal \code{blockind} K for \code{blockcor}.
-#' @return  Correlation matrix
+#' @return  Correlation matrix \code{Sigma}
 #' @export
 autocor <- function(p, rho){
   if (abs(rho) > 0.99){ stop("correlation rho must be between -0.99 and 0.99.") }
@@ -59,6 +59,8 @@ blockcor <- function(blockind, rho){
 #' @param type1 Type of the first dataset \code{X1}. Could be "continuous", "trunc", "binary", "ternary".
 #' @param type2 Type of the second dataset \code{X2}. Could be "continuous", "trunc", "binary", "ternary".
 #' @param muZ Mean of latent multivariate normal.
+#' @param p1 Dimension of variables belong to \code{type1}.
+#' @param p2 Dimension of variables belong to \code{type2}.
 #' @param c1 Constant threshold for \code{X1} needed for "trunc", "binary", "ternary" and "ordinal" data type - the default is NULL.
 #' @param c2 Constant threshold for \code{X2} needed for "trunc", "binary", "ternary" and "ordinal" data type - the default is NULL.
 #' @return \code{GenData} returns a list containing
@@ -68,8 +70,8 @@ blockcor <- function(blockind, rho){
 #'       \item{X1: }{observed numeric data matrix (n by p1).}
 #'       \item{X2: }{observed numeric data matrix (n by p2).}
 #'       \item{type: }{a vector containing types of two datasets.}
-#'       \item{c1: }{a matrix of thresholds for \code{X1} for "trunc", "binary", "ternary" and "ordinal" data type.}
[…]

### 2021-06-05 R/estR.R

@@ -75,4 +75,3 @@ estR <- function(X1, type1, X2 = NULL, type2 = NULL, method = "approx", use.near
     return(list(type = c(type1, type2), R1 = R1, R2 = R2, R12 = R12, R = R.final))
   }
 }
-

### 2021-06-05 man/examples/estimateR_ex.R

@@ -6,31 +6,19 @@ maxcancor <- 0.9 # true canonical correlation
 
 ### Correlation structure within each data set
 set.seed(0)
-perm1 <- sample(1:p1, size = p1);
-Sigma1 <- autocor(p1, 0.7)[perm1, perm1]
-blockind <- sample(1:3, size = p2, replace = TRUE);
-Sigma2 <- blockcor(blockind, 0.7)
-mu <- rbinom(p1+p2, 1, 0.5)
-
-### true variable indices for each dataset
-trueidx1 <- c(rep(1, 3), rep(0, p1-3))
-trueidx2 <- c(rep(1, 2), rep(0, p2-2))
+perm1 <- sample(1:(p1 + p2), size = p1 + p2);
+Sigma <- autocor(p1 + p2, 0.7)[perm1, perm1]
+# blockind <- sample(1:3, size = p2, replace = TRUE);
+# Sigma2 <- blockcor(blockind, 0.7)
+mu <- rbinom(p1 + p2, 1, 0.5)
 
 # Data generation
-simdata <- GenerateData(n=n, trueidx1 = trueidx1, trueidx2 = trueidx2,
-                        maxcancor = maxcancor,
-                        Sigma1 = Sigma1, Sigma2 = Sigma2,
-                        copula1 = "exp", copula2 = "cube",
-                        muZ = mu,
-                        type1 = "binary", type2 = "continuous",
-                        c1 = rep(1, p1), c2 =  NULL
-)
-X1 <- simdata$X1; X2 <- simdata$X2; Sigma12_tt <- simdata$Sigma12
+simdata <- GenData(n = n, copula1 = "exp", copula2 = "cube", type1 = "binary", type2 = "continuous",
+                   muZ = mu, Sigma = Sigma, p1 = p1, p2 = p2, c1 = rep(1, p1), c2 =  NULL)
+X1 <- simdata$X1; X2 <- simdata$X2; Sigma_tt <- simdata$Sigma
 # Estimate latent correlation matrix with original method
-R1_tt_org <- estR(X1, "binary", method = "original")
-R2_tt_org <- estR(X2, "continuous", method = "original")
-R12_tt_org <- estR(X1, type1 = "binary", X2, type2 = "continuous",
-                              method = "original")$R12
+R_tt_org <- estR(X1, type1 = "binary", X2, type2 = "continuous",
+                              method = "original")$R
 cp1 <- "exp"; cp2 <- "cube"
[…]

### 2021-06-05 paper/paper.Rmd
Commit

--- 2021-06-03 paper/paper.Rmd
@@
 # Estimate latent correlation matrix with original method
-R1_b_org <- estR(X1, "binary", method = "original")
-R2_c_org <- estR(X2, "continuous", method = "original")
-R12_bc_org <- estR(X1, type1 = "binary", X2, type2 = "continuous",
+R1_tt_org <- estR(X1, "binary", method = "original")
+R2_tt_org <- estR(X2, "continuous", method = "original")
+R12_tt_org <- estR(X1, type1 = "binary", X2, type2 = "continuous",
                               method = "original")$R12
 # Estimate latent correlation matrix with original method
-R1_b_approx <- estR(X1, "binary", method = "approx")
-R2_c_approx <- estR(X2, "continuous", method = "approx")
-R12_bc_approx <- estR(X1, type1 = "binary", X2, type2 = "continuous",
+R1_tt_org <- estR(X1, "binary", method = "approx")
+R2_tt_org <- estR(X2, "continuous", method = "approx")
+R12_tt_org <- estR(X1, type1 = "binary", X2, type2 = "continuous",
                               method = "approx")$R12

### 2021-06-05 paper/paper.tex
Commit

--- 2021-06-03 paper/paper.tex
@@
 \renewcommand{\headrulewidth}{0pt}
-\fancyhead[L]{\hspace{-0.75cm}\includegraphics[width=5.5cm]{C:/Users/sharkmanhmz/Documents/R/R-4.1.0/library/rticles/rmarkdown/templates/joss/resources/JOSS-logo.png}}
+\fancyhead[L]{\hspace{-0.75cm}\includegraphics[width=5.5cm]{C:/Users/mingz/OneDrive/Documents/R/win-library/4.1/rticles/rmarkdown/templates/joss/resources/JOSS-logo.png}}
 \fancyhead[C]{}
@@
 % For Pandoc 2.11+
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
  {% don't indent paragraphs
@@
 \newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}}
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
@@
 package \emph{pcaPP} (Croux, Filzmoser, \& Fritz, 2013) provides a fast
-calculation for Kendall's \tau. The R package \emph{MixedCCA} (Yoon,
+calculation for Kendall's \(\tau\). The R package \emph{MixedCCA} (Yoon,
 Carroll, \& Gaynanova, 2020) have functionality for latent correlation

### 2021-06-06 R/GenData.R
Generate figures for paper.

@@ -69,10 +69,6 @@ blockcor <- function(blockind, rho){
 #'       \item{Z2: }{latent numeric data matrix (n by p2).}
 #'       \item{X1: }{observed numeric data matrix (n by p1).}
 #'       \item{X2: }{observed numeric data matrix (n by p2).}
-#'       \item{type: }{a vector containing types of two datasets.}
-#'       \item{c1: }{a matrix of thresholds for \code{X1} for "trunc", "binary", "ternary" and data type.}
-#'       \item{c2: }{a matrix of thresholds for \code{X2} for "trunc", "binary", "ternary" and data type.}
-#'       \item{Sigma: }{true latent correlation matrix of \code{Z1} and \code{Z2} ((p1+p2) by (p1+p2)).}
 #' }
 #' @export
 #' @importFrom MASS mvrnorm
@@ -88,7 +84,7 @@ blockcor <- function(blockind, rho){
 #'  c1 = matrix(rep(1:2, p1), nrow = 2, ncol = p1), c2 = rep(0, p2))
 #'
 #'
-GenData <- function(n, copula1 = "no", copula2 = "no", type1 = "continuous", type2 = "continuous", muZ = NULL, Sigma, p1, p2, c1 = NULL, c2 = NULL){
+GenData <- function(n, type1, type2, sigma = NULL, p1 = 1, p2 = 1, copula1 = NULL, copula2 = NULL, muZ = NULL, Sigma = NULL, c1 = NULL, c2 = NULL){
   if((type1 != "continuous") & is.null(c1)){
     stop("c1 has to be defined for truncated continuous, binary, ternary or ordinal data type.")
   }
@@ -111,20 +107,17 @@ GenData <- function(n, copula1 = "no", copula2 = "no", type1 = "continuous", typ
   Z2 <- as.matrix(dat[, (p1+1):(p1 + p2)])
 
   # Three different types of copula
-  if(copula1 != "no"){
     if(copula1 == "exp"){
       Z1 <- exp(Z1)
     }else if(copula1 == "cube"){
       Z1 <- Z1^3
     }
-  }
-  if(copula2 != "no"){
+
     if(copula2 == "exp"){
       Z2 <- exp(Z2)
     }else if(copula2 == "cube"){
       Z2 <- Z2^3
     }
-  }
 
   if(type1 != "continuous"){
[…]

### 2021-06-06 man/examples/estimateR_ex.R

@@ -19,6 +19,7 @@ X1 <- simdata$X1; X2 <- simdata$X2; Sigma_tt <- simdata$Sigma
 # Estimate latent correlation matrix with original method
 R_tt_org <- estR(X1, type1 = "binary", X2, type2 = "continuous",
                               method = "original")$R
+
 cp1 <- "exp"; cp2 <- "cube"
 for (tp1 in c("continuous", "binary", "ternary", "trunc")) {
   for (tp2 in c("continuous", "binary", "ternary", "trunc")) {
@@ -36,7 +37,7 @@ for (tp1 in c("continuous", "binary", "ternary", "trunc")) {
       }
       simdata <- GenData(n=n, copula1 = cp1, copula2 = cp2, type1 = tp1, type2 = tp2,
                          muZ = mu, Sigma = Sigma, p1= p1, p2 = p2, c1 = c1, c2 = c2)
-      X1 <- simdata$X1; X2 <- simdata$X2; Sigma <- simdata$Sigma
+      X1 <- simdata$X1; X2 <- simdata$X2
       assign(paste("R", cp1, cp2, tp1, tp2, md, sep = "_"),
              estR(X1 = X1, type1 = tp1, X2 = X2, type2 = tp2, method = md)$R)
       PlotPair(datapair = cbind(c(Sigma), c(get(paste("R", cp1, cp2, tp1, tp2, md, sep = "_")))),

### 2021-06-11 R/GenData.R
Finish all tests.

@@ -6,54 +6,14 @@
 #' @param rho Correlation value(s), must be between -0.99 and 0.99. Should be a scalar for \code{autocor}, and either a scalar or a vector of the same length as the maximal \code{blockind} K for \code{blockcor}.
 #' @return  Correlation matrix \code{Sigma}
 #' @export
-autocor <- function(p, rho){
-  if (abs(rho) > 0.99){ stop("correlation rho must be between -0.99 and 0.99.") }
-  Sigma <- rho^abs(outer(1:p, 1:p, "-"))
-  return(Sigma)
-}
-
-
-#' Construct a correlation matrix
-#' @rdname Sigma
-#' @param blockind Block index 1,\dots, K for a positive integer K specifying which variable belongs to which block, the matrix dimension is equal to \code{length(blockind)}.
-#' @param rho Correlation value(s), must be between -0.99 and 0.99. Should be a scalar for \code{autocor}, and either a scalar or a vector of the same length as the maximal \code{blockind} K for \code{blockcor}.
-#' @return Correlation matrix \code{Sigma}
-#' @examples
-#' # For p = 8,
-#' # auto correlation matrix
-#' autocor(8, 0.8)
-#' # block correlation matrix: two blocks with the same correlation within each block
-#' blockcor(c(rep(1,3), rep(2,5)), 0.8)
-#' # block correlation matrix: two blocks with different correlation within each block
-#' blockcor(c(rep(1,3), rep(2,5)), c(0.8, 0.3))
-#'
-#' @export
-blockcor <- function(blockind, rho){
-  if(max(abs(rho)) > 0.99){ stop("correlation rho must be between -0.99 and 0.99.") }
-
-  p <- length(blockind)
-  blk <- unique(blockind)
-  if (length(rho) != length(blk)){
-    if(length(rho) == 1){
-      rho <- rep(rho, length(blk))
-    } else {
-      stop("rho and number of groups must match.")
-    }
-  }
-  Sigma <- matrix(0, p, p)
-
-  for (j in 1:length(blk)){
-    coef <- which(blockind %in% blk[j])
-    Sigma[coef, coef] <- rho[j]
-  }
-  diag(Sigma) = 1
[…]

### 2021-06-11 R/estR.R

@@ -36,12 +36,12 @@
 #' @example man/examples/estimateR_ex.R
 #'
 
-estR <- function(X1, type1, X2 = NULL, type2 = NULL, method = "approx", use.nearPD = TRUE, nu = 0.01, tol = 1e-6, verbose = FALSE, ratio = 0.9){
+estR = function(X1, type1, X2 = NULL, type2 = NULL, method = "approx", use.nearPD = TRUE, nu = 0.01, tol = 1e-6, verbose = FALSE, ratio = 0.9){
   # shrinkage method
   if(nu < 0 | nu > 1){
     stop("nu must be be between 0 and 1.")
   }
-  X1 <- as.matrix(X1); p1 <- ncol(X1)
+  X1 = as.matrix(X1); p1 = ncol(X1)
   if (length(colnames(X1)) == p1) {
     name1 = colnames(X1)
   } else {
@@ -49,13 +49,14 @@ estR <- function(X1, type1, X2 = NULL, type2 = NULL, method = "approx", use.near
   }
   if (is.null(X2)) {
     if (p1 == 1) {
-      out = as.matrix(1)
+      R = as.matrix(1)
     } else {
-      out = estimateR(X1 = X1, type1 = type1, method = method, tol = tol, ratio = ratio)
+      R = estimateR(X1 = X1, type1 = type1, method = method, tol = tol, ratio = ratio)
     }
-    return(out)
+    colnames(R) = rownames(R) = make.names(c(name1))
+    return(R)
   } else {
-    X2 <- as.matrix(X2); p2 <- ncol(X2)
+    X2 = as.matrix(X2); p2 = ncol(X2)
     if (length(colnames(X2)) == p2) {
       name2 = colnames(X2)
     } else {
@@ -67,15 +68,15 @@ estR <- function(X1, type1, X2 = NULL, type2 = NULL, method = "approx", use.near
       R1 = estimateR(X1 = X1, type1 = type1, method = method, tol = tol, ratio = ratio)
       R2 = estimateR(X1 = X2, type1 = type2, method = method, tol = tol, ratio = ratio)
       R12 = estimateR(X1 = X1, type1 = type1, X2 = X2, type2 = type2, method = method, tol = tol, ratio = ratio)
-      Rall <- rbind(cbind(R1, R12), cbind(t(R12), R2))
+      Rall = rbind(cbind(R1, R12), cbind(t(R12), R2))
[…]

### 2021-06-11 R/fromZtoX.R

@@ -1,0 +1,21 @@
+
+fromZtoX = function(Z, copula, type, c) {
+  if(copula == "exp"){
+    Z = exp(Z)
+  }else if(copula == "cube"){
+    Z = Z^3
+  }
+  if(type == "continuous") {
+    X = Z
+  } else if (ncol(c) != ncol(Z)) {
+    stop("The length of threshold vector c does not match with the dimension of the data p.")
+  } else if(type == "trunc") {
+    X = ifelse(Z > matrix(c, nrow = nrow(Z), ncol = ncol(Z), byrow = T), Z, 0)
+  } else if (type == "binary") {
+    X = ifelse(Z > matrix(c, nrow = nrow(Z), ncol = ncol(Z), byrow = T), 1, 0)
+  } else if (type == "ternary") {
+    X = ifelse(Z >= matrix(apply(c, 2, max), nrow = nrow(Z), ncol = ncol(Z), byrow = T), 2, 1)
+    X[Z <= matrix(apply(c, 2, min), nrow = nrow(Z), ncol = ncol(Z), byrow = T)] = 0
+  }
+  return(X)
+}

### 2021-06-11 man/examples/estimateR_ex.R

@@ -1,6 +1,5 @@
 ### Data setting
 n <- 1000; p1 <- 15; p2 <- 10 # sample size and dimensions for two datasets.
-maxcancor <- 0.9 # true canonical correlation
 
 ### Correlation structure within each data set
 set.seed(0)

### 2021-06-11 tests/testthat/test_GenData.R

@@ -1,0 +1,43 @@
+
+library(latentcor)
+
+### Data setting
+n = 100; p = 2 # sample size and dimensions for two datasets.
+
+test_that("autocor is symmetric.", {
+  expect_equal(autocor(p, 0.7), t(autocor(p, 0.7)))
+})
+
+Sigma = autocor(p, 0.7)
+mu = rbinom(p, 1, 0.5)
+dat = MASS::mvrnorm(n, mu = mu, Sigma = Sigma) # generate a data matrix of size.
+
+test_that("truncated data have lower bound zero.", {
+  expect_equal(apply(fromZtoX(Z = dat, copula = "cube", type = "trunc", c = matrix(rep(0, p), ncol = p)), 2, min), rep(0, p))
+})
+
+test_that("binary data either zero or one.", {
+  expect_equal(sort(unique(c(fromZtoX(Z = dat, copula = "cube", type = "binary", c = matrix(rep(0, p), ncol = p))))), c(0, 1))
+})
+
+test_that("ternary data should be zero, one or two.", {
+  expect_equal(sort(unique(c(fromZtoX(Z = dat, copula = "cube", type = "ternary", c = matrix(rep(0:1, p), ncol = p))))), c(0, 1, 2))
+})
+
+p1 = 1; p2 = 2
+Sigma = autocor(p1 + p2, 0.7)
+mu = rbinom(p1 + p2, 1, 0.5)
+
+test_that("truncated data have lower bound zero.", {
+  expect_equal(apply(GenData(n = n, copula1 = "cube", copula2 = "cube", type1 = "ternary", type2 = "trunc",muZ = mu, Sigma = Sigma, p1 = p1, p2 = p2,c1 = matrix(rep(0:1, p1), nrow = 2, ncol = p1), c2 = matrix(rep(0, p2), ncol = p2))$X2, 2, min), rep(0, p2))
+})
+
+test_that("truncated data have lower bound zero.", {
+  expect_equal(sort(unique(c(GenData(n = n, copula1 = "cube", copula2 = "cube", type1 = "ternary", type2 = "binary",muZ = mu, Sigma = Sigma, p1 = p1, p2 = p2,c1 = matrix(rep(0:1, p1), nrow = 2, ncol = p1), c2 = matrix(rep(0, p2), ncol = p2))$X2))), c(0, 1))
+})
+
+test_that("ternary data should be zero, one or two.", {
[…]

### 2021-06-11 tests/testthat/test_R_adj.R

@@ -1,0 +1,16 @@
+
+library(latentcor)
+
+R = autocor(3, 0.7)
+
+test_that("Adjust nearPD won't destroy symmetry.", {
+  expect_equal(R_adj(R = R, use.nearPD = TRUE, verbose = FALSE, nu = .5),
+               t(R_adj(R = R, use.nearPD = TRUE, verbose = FALSE, nu = .5)))
+})
+
+R = diag(1, 3); R[3, 3] = 0
+
+test_that("Adjust nearPD won't destroy symmetry.", {
+  expect_equal(R_adj(R = R, use.nearPD = TRUE, verbose = FALSE, nu = .5),
+               t(R_adj(R = R, use.nearPD = TRUE, verbose = FALSE, nu = .5)))
+})

### 2021-06-11 tests/testthat/test_estR.R

@@ -1,0 +1,223 @@
+
+library(latentcor)
+
+### Data setting
+n <- 1000; p1 <- 3; p2 <- 2 # sample size and dimensions for two datasets.
+maxcancor <- 0.9 # true canonical correlation
+
+### Correlation structure within each data set
+perm1 <- sample(1:(p1 + p2), size = p1 + p2);
+Sigma <- autocor(p1 + p2, 0.7)[perm1, perm1]
+mu <- rbinom(p1 + p2, 1, 0.5)
+# Data generation
+simdata <- GenData(n=n, type1 = "continuous", type2 = "binary", p1 = p1, p2 = p2, copula1 = "cube",
+                   copula2 = "cube",  muZ = mu, Sigma = Sigma,
+                   c1 = NULL, c2 =  rep(0, p2))
+X1 <- simdata$X1; X2 <- simdata$X2
+colnames(X1) = c("X11", "X12", "X13"); colnames(X2) = c("X21", "X22")
+
+test_that("estR is symmetric.", {
+  expect_equal(estR(X1 = X1, type1 = "continuous", method = "original", use.nearPD = TRUE, nu = 0.5, tol = 1e-6, ratio = .9),
+               t(estR(X1 = X1, type1 = "continuous", method = "original", use.nearPD = TRUE, nu = 0.5, tol = 1e-6, ratio = .9)))
+})
+
+test_that("estR is symmetric.", {
+  expect_equal(estR(X1 = X1, type1 = "continuous", method = "ml", use.nearPD = TRUE, nu = 0.5, tol = 1e-6, ratio = .9),
+               t(estR(X1 = X1, type1 = "continuous", method = "ml", use.nearPD = TRUE, nu = 0.5, tol = 1e-6, ratio = .9)))
+})
+
+test_that("estR is symmetric.", {
+  expect_equal(estR(X1 = X2, type1 = "binary", method = "original", use.nearPD = TRUE, nu = 0.5, tol = 1e-6, ratio = .9),
+               t(estR(X1 = X2, type1 = "binary", method = "original", use.nearPD = TRUE, nu = 0.5, tol = 1e-6, ratio = .9)))
+})
+
+test_that("estR is symmetric.", {
+  expect_equal(estR(X1 = X2, type1 = "binary", method = "ml", use.nearPD = TRUE, nu = 0.5, tol = 1e-6, ratio = .9),
[…]

### 2021-06-11 tests/testthat/test_estimateR.R

@@ -6,7 +6,6 @@ n <- 1000; p1 <- 3; p2 <- 2 # sample size and dimensions for two datasets.
 maxcancor <- 0.9 # true canonical correlation
 
 ### Correlation structure within each data set
-set.seed(0)
 perm1 <- sample(1:(p1 + p2), size = p1 + p2);
 Sigma <- autocor(p1 + p2, 0.7)[perm1, perm1]
 mu <- rbinom(p1 + p2, 1, 0.5)
@@ -17,8 +16,63 @@ simdata <- GenData(n=n, type1 = "continuous", type2 = "binary", p1 = p1, p2 = p2
 X1 <- simdata$X1; X2 <- simdata$X2
 
 test_that("estimateR is symmetric.", {
+  expect_equal(estimateR(X1 = X1, type1 = "continuous", method = "original", tol = 1e-6, ratio = .9),
+               t(estimateR(X1 = X1, type1 = "continuous", method = "original", tol = 1e-6, ratio = .9)))
+})
+
+test_that("estimateR is symmetric.", {
+  expect_equal(estimateR(X1 = X1, type1 = "continuous", method = "ml", tol = 1e-6, ratio = .9),
+               t(estimateR(X1 = X1, type1 = "continuous", method = "ml", tol = 1e-6, ratio = .9)))
+})
+
+test_that("estimateR is symmetric.", {
+  expect_equal(estimateR(X1 = X2, type1 = "binary", method = "original", tol = 1e-6, ratio = .9),
+               t(estimateR(X1 = X2, type1 = "binary", method = "original", tol = 1e-6, ratio = .9)))
+})
+
+test_that("estimateR is symmetric.", {
+  expect_equal(estimateR(X1 = X2, type1 = "binary", method = "ml", tol = 1e-6, ratio = .9),
+               t(estimateR(X1 = X2, type1 = "binary", method = "ml", tol = 1e-6, ratio = .9)))
+})
+
+simdata <- GenData(n=n, type1 = "trunc", type2 = "ternary", p1 = p1, p2 = p2, copula1 = "cube",
+                   copula2 = "cube",  muZ = mu, Sigma = Sigma,
+                   c1 = rep(0, p1), c2 =  matrix(rep(0:1, p2), nrow = 2, ncol = p2))
+X1 <- simdata$X1; X2 <- simdata$X2
+
+test_that("estimateR is symmetric.", {
[…]

### 2021-06-12 man/examples/estimateR_ex.R
Add example to README.md

@@ -18,6 +18,33 @@ R_nc_org <- estR(X1 = X1, type1 = "ternary", X2 = X2, type2 = "continuous",
 R_nc_approx <- estR(X1 = X1, type1 = "ternary", X2 = X2, type2 = "continuous",
                     method = "approx")$R
 
+cp1 <- "exp"; cp2 <- "cube"
+for (tp1 in c("continuous", "binary", "ternary", "trunc")) {
+  for (tp2 in c("continuous", "binary", "ternary", "trunc")) {
+    for (md in c("original", "ml", "approx")) {
+      c1 <- c2 <- NULL
+      if (tp1 == "binary" | tp1 == "trunc") {
+        c1 <- rep(1, p1)
+      } else if (tp1 == "ternary") {
+        c1 <- matrix(rep(1:2, p1), nrow = 2, ncol = p1)
+      }
+      if (tp2 == "binary" | tp2 == "trunc") {
+        c2 <- rep(0, p2)
+      } else if (tp2 == "ternary") {
+        c2 <- matrix(rep(0:1, p2), nrow = 2, ncol = p2)
+      }
+      simdata <- GenData(n=n, copula1 = cp1, copula2 = cp2, type1 = tp1, type2 = tp2,
+                         muZ = mu, Sigma = Sigma, p1= p1, p2 = p2, c1 = c1, c2 = c2)
+      X1 <- simdata$X1; X2 <- simdata$X2
+      assign(paste("R", cp1, cp2, tp1, tp2, md, sep = "_"),
+             estR(X1 = X1, type1 = tp1, X2 = X2, type2 = tp2, method = md)$R)
+      PlotPair(datapair = cbind(c(Sigma), c(get(paste("R", cp1, cp2, tp1, tp2, md, sep = "_")))),
+               namepair = c("Sigma", paste("R", cp1, cp2, tp1, tp2, md, sep = "_")),
+               title = "Latent correlation (True vs. Estimated)")
+    }
+  }
+}
+
 # Plot ternary/continuous case estimation via original method.
 PlotPair(datapair = cbind(c(Sigma), c(R_nc_org)), namepair = c("Sigma", "R_nc_org"),
          title = "Latent correlation (True vs. Estimated)")

### 2021-07-17 R/GenData.R
Fix bugs and documentations.

@@ -3,19 +3,19 @@
 #' @description \code{GenData} is used to generate data of mixed types for the Gaussian copula model. It can generate single, pair or multiple variables.
 #' @param n a positive interger to specify sample size. The default value is 100.
 #' @param types A length of p vector as types for each variable. Could be "con" (continuous), "bin" (binary), "tru" (truncated) or "ter" (ternary).
+#'              The number of variables p is determined based on supplied length of types (or something along those lines).
 #'              The default value is c("ter", "con") which specifies a pair of variables, the first one is ternary, the second one is continuous.
-#' @param rhos A length of \code{sum(lower.tri(diag(p)))} vector as true latent correlations for between variables.
-#'             The default value is 0.5 which means correlations between any two variables are 0.5.
-#' @param copulas A length of p vector as copula types for each variable, e.g. U = f(Z). Which could be "no", "expo" or "cube".
+#' @param rhos A vector with length of 1 as true correlation for between any two variables (e.g. off-diagonal elements of correlation matrix, so that the correlation matrix is \code{matrix(c(1, rhos, rhos, 1), 2, 2)}) or length of \code{sum(lower.tri(diag(p)))} as lower triangular elements of correlation matrix (e.g. \code{rhos = c(.3, .5, .7)} means the correlation matrix is \code{matrix(c(1, .3, .5, .3, 1, .7, .5, .7, 1), 3, 3)}). The default value is 0.5 which means correlations between any two variables are 0.5.
+#' @param copulas A vector with length of 1 for copula type of all variables or p for copula types for each variable, e.g. U = f(Z). Which could be "no", "expo" (exponential) or "cube" (cube).
 #'                The default value is "no" which means no copula transformation for any variables.
 #' @param XP A length of p list of proportions of zeros (and ones for ternary) corresponding to each variable.
+#'           If null, the following values are automatically generated as elements of \code{XP} list for the corresponding data types:
 #'           For continuous variable, the corresponding proportion should be NA;
 #'           for binary or truncated variable, the corresponding proportion should be a number between 0 and 1 represents the proportion of zeros;
[…]

### 2021-07-17 R/estR.R

@@ -1,27 +1,27 @@
-#' @title Calculate Kendall's Tau matrix then estimate latent correlation matrix
-#' @description Estimation of latent correlation matrix from observed data of (possibly) mixed types (continuous/biary/truncated continuous) based on the latent Gaussian copula model.
+#' @title Estimate latent correlation for mixed types.
+#' @description Estimation of latent correlation matrix from observed data of (possibly) mixed types (continuous/binary/truncated/ternary) based on the latent Gaussian copula model. Missing values (NA) are allowed. The estimation is based on pairwise complete observations.
 #' @rdname estR
 #' @aliases estR
-#' @param X A numeric data matrix (n by p).
-#' @param types A vector with length of p which specifies types of variables in \code{X} correspondingly.
-#'              Must be one of "con" (continuous), "bin" (binary), "tru" (truncated) or "ter" (ternary). The default value is c("ter", "con") which means the first variable is ternary, second variable is continuous.
-#' @param method The calculation method of latent correlation. Either "original" method or "approx". If \code{method = "approx"}, multilinear approximation method is used, which is much faster than the original method.
-#'               If \code{method = "original"}, optimization of the bridge inverse function is used. The default is "approx".
+#' @param X A numeric data matrix (n by p). Missing values (NA) are allowed. The estimation is based on pairwise complete observations.
+#' @param types A vector with length of 1 for the type of all variables or p which specifies types of variables in \code{X} correspondingly. Must be one of "con" (continuous), "bin" (binary), "tru" (truncated) or "ter" (ternary). The default value is "con" which means all variables are continuous.
+#' @param method The calculation method of latent correlation. Either "original" method or "approx". If \code{method = "approx"}, multilinear approximation method is used, which is much faster than the original method. If \code{method = "original"}, optimization of the bridge inverse function is used. The default is "approx".
 #' @param nu Shrinkage parameter for correlation matrix, must be between 0 and 1, the default value is 0.01.
[…]

### 2021-07-17 R/internal.R

@@ -4,25 +4,25 @@ fromZtoX = function(z, type, copula, xp) {
   type_switch = switch(type, "con" = function(u, xp) {x = u; return(x)},
                        "bin" = function(u, xp) {q = quantile(u, xp); x = ifelse(u > q, 1, 0); return(x)},
                        "tru" = function(u, xp) {q = quantile(u, xp); x = ifelse(u > q, u, q) - q; return(x)},
-                       "ter" = function(u, xp) {q = quantile(u, cumsum(xp)); x = rep(0, length(u)); x[u > q[1] & u <= q[2]] = 1; x[u > q[2]] = 2; return(x)},
-                       "qua" = function(u, xp) {q = quantile(u, cumsum(xp)); x = rep(0, length(u)); x[u > q[1] & u <= q[2]] = 1; x[u > q[2] & u <= q[3]] = 2;
-                       x[u > q[3]] = 3; return(x)},
-                       "qui" = function(u, xp) {q = quantile(u, cumsum(xp)); x = rep(0, length(u)); x[u > q[1] & u <= q[2]] = 1; x[u > q[2] & u <= q[3]] = 2;
-                       x[u > q[3] & u <= q[4]] = 3; x[u > q[4]] = 4; return(x)},
-                       "sen" = function(u, xp) {q = quantile(u, cumsum(xp)); x = rep(0, length(u)); x[u > q[1] & u <= q[2]] = 1; x[u > q[2] & u <= q[3]] = 2;
-                       x[u > q[3] & u <= q[4]] = 3; x[u > q[4] & u <= q[5]] = 4; x[u > q[5]] = 5; return(x)},
-                       "sep" = function(u, xp) {q = quantile(u, cumsum(xp)); x = rep(0, length(u)); x[u > q[1] & u <= q[2]] = 1; x[u > q[2] & u <= q[3]] = 2;
-                       x[u > q[3] & u <= q[4]] = 3; x[u > q[4] & u <= q[5]] = 4; x[u > q[5] & u <= q[6]] = 5; x[u > q[6]] = 6; return(x)},
-                       "oct" = function(u, xp) {q = quantile(u, cumsum(xp)); x = rep(0, length(u)); x[u > q[1] & u <= q[2]] = 1; x[u > q[2] & u <= q[3]] = 2;
[…]

### 2021-07-18 R/estR.R
Add example for vignette.

@@ -11,7 +11,7 @@
 #' @param showplot Logical indicator. \code{showplot = TRUE} to generate an ggplot object \code{plotR} in output list as the heatmap of latent correlation matrix \code{R}. \code{plotR = NULL} if \code{showplot = FALSE}.
 #' @return \code{estR} returns
 #' \itemize{
-#'       \item{zratios: }{A list of zratios. Each element correponds to the zratios for one variable. zratios for continuous variables is NA (no zratios); zratios for binary/truncated variables is the proportions of zeros; zratios for ternary variables is the accumulated proportions of zeros and ones (e.g. first value is proportion of zeros, second value is proportion of zeros and ones)}
+#'       \item{zratios: }{A list of zratios. Each element correponds to the zratios for one variable. zratios for continuous variables is NA (no zratios); zratios for binary/truncated variables is the proportions of zeros; zratios for ternary variables is the cumulative proportions of zeros and ones (e.g. first value is proportion of zeros, second value is proportion of zeros and ones)}
 #'       \item{K: }{Kendall Tau (Tau-a) Matrix of \code{X} (p x p)}
 #'       \item{R: }{Estimated latent correlation matrix of whole \code{X} (p x p)}
 #'       \item{plotR: }{Heatmap plot for latent correlation matrix \code{R}, return NULL if \code{showplot = FALSE}}

### 2021-07-22 R/estR.R
Minor adjustments to estR documentation

@@ -2,25 +2,27 @@
 #' @description Estimation of latent correlation matrix from observed data of (possibly) mixed types (continuous/binary/truncated/ternary) based on the latent Gaussian copula model. Missing values (NA) are allowed. The estimation is based on pairwise complete observations.
 #' @rdname estR
 #' @aliases estR
-#' @param X A numeric data matrix (n by p). Missing values (NA) are allowed. The estimation is based on pairwise complete observations.
-#' @param types A vector with length of 1 for the type of all variables or p which specifies types of variables in \code{X} correspondingly. Must be one of "con" (continuous), "bin" (binary), "tru" (truncated) or "ter" (ternary). The default value is "con" which means all variables are continuous.
-#' @param method The calculation method of latent correlation. Either "original" method or "approx". If \code{method = "approx"}, multilinear approximation method is used, which is much faster than the original method. If \code{method = "original"}, optimization of the bridge inverse function is used. The default is "approx".
-#' @param nu Shrinkage parameter for correlation matrix, must be between 0 and 1, the default value is 0.01.
+#' @param X A numeric data matrix (n by p), where n is number of samples, and p is number of variables. Missing values (NA) are allowed, in which case the estimation is based on pairwise complete observations.
+#' @param types A vector of length p indicating the type of each of the p variables in \code{X}. Each element must be one of \code{"con"} (continuous), \code{"bin"} (binary), \code{"tru"} (truncated) or \code{"ter"} (ternary). If the vector has length 1, then all p variables are assumed to be of the same type that is supplied. The default value is \code{"con"} which means all variables are continuous.
+#' @param method The calculation method for latent correlations. Either \code{"original"} or \code{"approx"}. If \code{method = "approx"}, multilinear approximation method is used, which is much faster than the original method. If \code{method = "original"}, optimization of the bridge inverse function is used. The default is \code{"approx"}.
+#' @param nu Shrinkage parameter for the correlation matrix, must be between 0 and 1, the default value is 0.01.
 #' @param tol Desired accuracy when calculating the solution of bridge function. The default value is 1e-8.
[…]

### 2021-07-22 R/GenData.R
Minor documentation clarifications for GenData

@@ -1,28 +1,28 @@
 #'
 #' @title Mixed type simulation data generator
-#' @description \code{GenData} is used to generate data of mixed types for the Gaussian copula model. It can generate single, pair or multiple variables.
-#' @param n a positive interger to specify sample size. The default value is 100.
-#' @param types A length of p vector as types for each variable. Could be "con" (continuous), "bin" (binary), "tru" (truncated) or "ter" (ternary).
-#'              The number of variables p is determined based on supplied length of types (or something along those lines).
-#'              The default value is c("ter", "con") which specifies a pair of variables, the first one is ternary, the second one is continuous.
-#' @param rhos A vector with length of 1 as true correlation for between any two variables (e.g. off-diagonal elements of correlation matrix, so that the correlation matrix is \code{matrix(c(1, rhos, rhos, 1), 2, 2)}) or length of \code{sum(lower.tri(diag(p)))} as lower triangular elements of correlation matrix (e.g. \code{rhos = c(.3, .5, .7)} means the correlation matrix is \code{matrix(c(1, .3, .5, .3, 1, .7, .5, .7, 1), 3, 3)}). The default value is 0.5 which means correlations between any two variables are 0.5.
-#' @param copulas A vector with length of 1 for copula type of all variables or p for copula types for each variable, e.g. U = f(Z). Which could be "no", "expo" (exponential) or "cube" (cube).
-#'                The default value is "no" which means no copula transformation for any variables.
-#' @param XP A length of p list of proportions of zeros (and ones for ternary) corresponding to each variable.
-#'           If null, the following values are automatically generated as elements of \code{XP} list for the corresponding data types:
-#'           For continuous variable, the corresponding proportion should be NA;
-#'           for binary or truncated variable, the corresponding proportion should be a number between 0 and 1 represents the proportion of zeros;
-#'           for ternary variable, the corresponding proportion should be a pair of numbers between 0 and 1, the first number indicates the the proportion of zeros, the second number indicates the proportion of ones. The sum of a pair of numbers should between 0 and 1.
-#'           The default value for continuous variable is NA, for binary/truncated variable is 0.5, for ternary variable is c(0.3, 0.5).
[…]

### 2021-08-08 R/internal.R
Eventurally I found the bug which made real example have problem.

@@ -271,7 +271,7 @@ r_sol = function(K, zratio1, zratio2, comb, tol, ratio) {
   )
   K.len = length(K); out = rep(NA, K.len);
   zratio1 = as.matrix(zratio1); zratio2 = as.matrix(zratio2)
-  for (i in K.len) {
+  for (i in 1:K.len) {
     f = function(r)(bridge_switch(r = r, zratio1 = zratio1[ , i], zratio2 = zratio2[ , i]) - K[i])^2
     op = tryCatch(optimize(f, lower = -0.999, upper = 0.999, tol = tol)[1], error = function(e) 100)
     if(op == 100) {

### 2021-08-08 R/estR.R
Re-encode values for binary/truncated/ternary variables. Add mtcars as example.

@@ -50,12 +50,11 @@ estR = function(X, types = "con", method = c("approx", "original"), nu = 0.001, 
   if (length(types) == 1) {
     types = rep(types, p)
   } else if (length(types) != p) {
-    stop("Length of types should be either 1 or p which specifies types of p variables.")
+    stop("Length of types should be either 1 for all variables or the same as number of variables (columns of X).")
   }
+  # recoding binary, truncated and ternary values.
+  X = encodeX(X, types)
   method = match.arg(method, several.ok = FALSE)
-  if (length(types) != p) {
-    stop("types should have the same length as the number of variables (columns of X).")
-    }
   if (length(colnames(X)) == p) {
     name = colnames(X)
   } else {
@@ -93,10 +92,10 @@ estR = function(X, types = "con", method = c("approx", "original"), nu = 0.001, 
   # Check if the matrix is semi-pos.definite
   R_min_eig = min(eigen(R)$values)
   if (R_min_eig < 0) {
-    message("Using Matrix::nearPD since Minimum eigenvalue of latent correlation matrix is ", R_min_eig, "smaller than 0.")
-    R = as.matrix(Matrix::nearPD(R, corr = TRUE, maxit = 1000)$mat)
+    message("Using Matrix::nearPD since Minimum eigenvalue of latent correlation matrix is ", R_min_eig, " smaller than 0.")
+    R = as.matrix(Matrix::nearPD(R, corr = TRUE)$mat)
   }
-  # Do adjustmnet by nu - makes it stricly positive definite like ridge
+  # Do adjustmnet by nu - makes it strictly positive definite like ridge
   R = (1 - nu) * R + nu * diag(nrow(R))
   colnames(K) = rownames(K) = colnames(R) = rownames(R) = make.names(c(name))
   plotR = NULL

### 2021-08-08 R/internal.R

@@ -50,14 +50,33 @@ n_x = function(x, n) {
   return(n_x)
 }
 
+encodeX = function(X, types) {
+  colmin_mat = matrix(apply(X, 2, function(x) min(x, na.rm = TRUE)), nrow = nrow(X), ncol = ncol(X), byrow = TRUE)
+  colmax_mat = matrix(apply(X, 2, function(x) max(x, na.rm = TRUE)), nrow = nrow(X), ncol = ncol(X), byrow = TRUE)
+  if (sum(types == "bin") > 0) {
+  X_bin = X[ , types == "bin"]
+  X[ , types == "bin"][X_bin == colmin_mat[ , types == "bin"]] = 0; X[ , types == "bin"][X_bin == colmax_mat[ , types == "bin"]] = 1
+  }
+  if (sum(types == "tru") > 0) {
+  X_tru = X[ , types == "tru"]
+  X[ , types == "tru"] = X_tru - colmin_mat[ , types == "tru"]
+  }
+  if (sum(types == "ter") > 0) {
+  X_ter = X[ , types == "ter"]
+  X[ , types == "ter"][X_ter == colmin_mat[ , types == "ter"]] = 0; X[ , types == "ter"][X_ter == colmax_mat[ , types == "ter"]] = 2
+  X[ , types == "ter"][(!(is.na(X_ter))) & (X_ter != colmin_mat[ , types == "ter"]) & (X_ter != colmax_mat[ , types == "ter"])] = 1
+  }
+  return(as.matrix(X))
+}
+
 zratios = function(X, types) {
   X = as.matrix(X); out = vector(mode = "list", length = ncol(X))
   for (type in unique(types)) {
     zratios_switch = switch(type, "con" = function(X) zratios = rep(NA, ncol(as.matrix(X))),
-                            "bin" = function(X) zratios = colMeans(as.matrix(X == 0)),
-                            "tru" = function(X) zratios = colMeans(as.matrix(X == 0)),
+                            "bin" = function(X) zratios = colMeans(as.matrix(X == 0), na.rm = TRUE),
+                            "tru" = function(X) zratios = colMeans(as.matrix(X == 0), na.rm = TRUE),
                             "ter" = function(X) {X0 = X == 0; X1 = X == 1;
-                              zratios = rbind(colMeans(as.matrix(X0)), colMeans(as.matrix(X0 + X1)))
[…]

### 2021-08-08 tests/testthat/test_encodeX.R

@@ -1,0 +1,7 @@
+library(latentcor)
+
+test_that("Always recode binary/ternary/truncated", {
+  types = c("con", "bin", "tru", "ter")
+  X = GenData(types = types)$X; X[sample(length(X), 10)] = NA
+  expect_equal(encodeX(X = X, types = types), X)
+})

### 2021-08-08 R/estR.R
Finish example for mtcars.

@@ -52,14 +52,14 @@ estR = function(X, types = "con", method = c("approx", "original"), nu = 0.001, 
   } else if (length(types) != p) {
     stop("Length of types should be either 1 for all variables or the same as number of variables (columns of X).")
   }
-  # recoding binary, truncated and ternary values.
-  X = encodeX(X, types)
   method = match.arg(method, several.ok = FALSE)
   if (length(colnames(X)) == p) {
     name = colnames(X)
   } else {
     name = paste0("X", 1:p)
   }
+  # recoding binary, truncated and ternary values.
+  X = encodeX(X, types)
   R = matrix(0, p, p); cp = rbind(row(R)[lower.tri(R)], col(R)[lower.tri(R)]); cp.col = ncol(cp)
   if (any(is.na(X))) {
     K_a.lower = sapply(seq(p), function(i) Kendalltau(X[ , cp[ , i]]))
@@ -97,7 +97,7 @@ estR = function(X, types = "con", method = c("approx", "original"), nu = 0.001, 
   }
   # Do adjustmnet by nu - makes it strictly positive definite like ridge
   R = (1 - nu) * R + nu * diag(nrow(R))
-  colnames(K) = rownames(K) = colnames(R) = rownames(R) = make.names(c(name))
+  colnames(K) = rownames(K) = colnames(R) = rownames(R) = colnames(Rpointwise) = rownames(Rpointwise) = make.names(c(name))
   plotR = NULL
   if (showplot) {
     plotR = heatmaply(R, dendrogram = "none", main = "Latent Correlation", margins = c(80,80,80,80),

### 2021-08-08 R/internal.R

@@ -51,6 +51,7 @@ n_x = function(x, n) {
 }
 
 encodeX = function(X, types) {
+  X = matrix(unlist(X), nrow = nrow(X), ncol = ncol(X))
   colmin_mat = matrix(apply(X, 2, function(x) min(x, na.rm = TRUE)), nrow = nrow(X), ncol = ncol(X), byrow = TRUE)
   colmax_mat = matrix(apply(X, 2, function(x) max(x, na.rm = TRUE)), nrow = nrow(X), ncol = ncol(X), byrow = TRUE)
   if (sum(types == "bin") > 0) {
@@ -289,7 +290,7 @@ r_sol = function(K, zratio1, zratio2, comb, tol, ratio) {
   #  }
   )
   K.len = length(K); out = rep(NA, K.len);
-  zratio1 = as.matrix(zratio1); zratio2 = as.matrix(zratio2)
+  zratio1 = matrix(zratio1, ncol = K.len); zratio2 = matrix(zratio2, ncol = K.len)
   for (i in 1:K.len) {
     f = function(r)(bridge_switch(r = r, zratio1 = zratio1[ , i], zratio2 = zratio2[ , i]) - K[i])^2
     op = tryCatch(optimize(f, lower = -0.999, upper = 0.999, tol = tol)[1], error = function(e) 100)
@@ -306,8 +307,8 @@ r_sol = function(K, zratio1, zratio2, comb, tol, ratio) {
 r_ml = function(K, zratio1, zratio2, comb, tol, ratio) {
   ipol_switch = switch(comb, "10" = BCipol, "11" = BBipol, "20" = TCipol, "21" = TBipol, "22" = TTipol,
                        "30" = NCipol, "31" = NBipol, "32" = NTipol, "33" = NNipol)
-  zratio1 = as.matrix(zratio1); zratio1.row = nrow(zratio1)
-  zratio2 = as.matrix(zratio2); zratio2.row = nrow(zratio2)
+  zratio1 = matrix(zratio1, ncol = length(K)); zratio1.row = nrow(zratio1)
+  zratio2 = matrix(zratio2, ncol = length(K)); zratio2.row = nrow(zratio2)
   if (zratio1.row > 1) {zratio1[1:(zratio1.row - 1), ] = zratio1[1:(zratio1.row - 1), ] / zratio1[2:zratio1.row, ]}
[…]

### 2021-08-10 R/estR.R
Adjust plots

@@ -100,8 +100,8 @@ estR = function(X, types = "con", method = c("approx", "original"), nu = 0.001, 
   colnames(K) = rownames(K) = colnames(R) = rownames(R) = colnames(Rpointwise) = rownames(Rpointwise) = make.names(c(name))
   plotR = NULL
   if (showplot) {
-    plotR = heatmaply(R, dendrogram = "none", main = "Latent Correlation", margins = c(80,80,80,80),
-                      grid_color = "white", grid_width = 0.00001, label_names = c("Horizontal axis:", "Vertical axis:", "Latent correlation:"))
+    plotR = heatmaply(R, dendrogram = "none", main = paste0("Estimated latent correlation (", method, ")"), margins = c(80,80,80,80),
+                      grid_color = "white", grid_width = 0.00001, label_names = c("Horizontal axis:", "Vertical axis:", "Latent correlation:"), limits = c(-1, 1))
   }
   return(list(zratios = zratios, K = K, R = R, Rpointwise = Rpointwise, plotR = plotR))
 }

### 2021-08-10 R/estR.R
Revise plot.

@@ -101,7 +101,7 @@ estR = function(X, types = "con", method = c("approx", "original"), nu = 0.001, 
   plotR = NULL
   if (showplot) {
     plotR = heatmaply(R, dendrogram = "none", main = paste0("Estimated latent correlation (", method, ")"), margins = c(80,80,80,80),
-                      grid_color = "white", grid_width = 0.00001, label_names = c("Horizontal axis:", "Vertical axis:", "Latent correlation:"), limits = c(-1, 1))
+                      grid_color = "white", grid_width = 0.00001, label_names = c("Horizontal axis:", "Vertical axis:", "Latent correlation:"))
   }
   return(list(zratios = zratios, K = K, R = R, Rpointwise = Rpointwise, plotR = plotR))
 }

### 2021-08-10 R/estR.R
Put all heatmap together.

@@ -101,7 +101,7 @@ estR = function(X, types = "con", method = c("approx", "original"), nu = 0.001, 
   plotR = NULL
   if (showplot) {
     plotR = heatmaply(R, dendrogram = "none", main = paste0("Estimated latent correlation (", method, ")"), margins = c(80,80,80,80),
-                      grid_color = "white", grid_width = 0.00001, label_names = c("Horizontal axis:", "Vertical axis:", "Latent correlation:"))
+                      grid_color = "white", grid_width = 0.00001, label_names = c("Horizontal axis:", "Vertical axis:", "Latent correlation:"), limits = c(-1, 1))
   }
   return(list(zratios = zratios, K = K, R = R, Rpointwise = Rpointwise, plotR = plotR))
 }

### 2021-08-26 R/estR.R
expand estR documentation on tol, nu and ratio

@@ -4,10 +4,10 @@
 #' @aliases estR
 #' @param X A numeric data matrix (n by p), where n is number of samples, and p is number of variables. Missing values (NA) are allowed, in which case the estimation is based on pairwise complete observations.
 #' @param types A vector of length p indicating the type of each of the p variables in \code{X}. Each element must be one of \code{"con"} (continuous), \code{"bin"} (binary), \code{"tru"} (truncated) or \code{"ter"} (ternary). If the vector has length 1, then all p variables are assumed to be of the same type that is supplied. The default value is \code{"con"} which means all variables are continuous.
-#' @param method The calculation method for latent correlations. Either \code{"original"} or \code{"approx"}. If \code{method = "approx"}, multilinear approximation method is used, which is much faster than the original method. If \code{method = "original"}, optimization of the bridge inverse function is used. The default is \code{"approx"}.
-#' @param nu Shrinkage parameter for the correlation matrix, must be between 0 and 1, the default value is 0.001.
-#' @param tol Desired accuracy when calculating the solution of bridge function. The default value is 1e-8.
-#' @param ratio The boundary value for multilinear interpolation, must be between 0 and 1, the default value is 0.9. If \code{method = "original"}, this parameter is ignored.
+#' @param method The calculation method for latent correlations. Either \code{"original"} or \code{"approx"}. If \code{method = "approx"}, multilinear approximation method is used, which is much faster than the original method, see Yoon et al. (2021). If \code{method = "original"}, optimization of the bridge inverse function is used. The default is \code{"approx"}.
+#' @param nu Shrinkage parameter for the correlation matrix, must be between 0 and 1. Guarantees that the minimal eigenvalue of returned correlation matrix is greater or equal to \code{nu}. When \code{nu = 0}, no shrinkage is performed, the returned correlation matrix will be semi-positive definite but not necessarily strictly positive definite. When \code{nu = 1}, the identity matrix is returned (not recommended).  The default (recommended) value is 0.001.
+#' @param tol When \code{method = "original"}, specifies the desired accuracy of the bridge function inversion via uniroot optimization and is passed to \code{\link{optimize}}. The default value is 1e-8. When \code{method = "approx"}, this parameter is ignored.
[…]

### 2021-08-26 R/latentcor.R
Rename

@@ -1,7 +1,7 @@
 #' @title Estimate latent correlation for mixed types.
 #' @description Estimation of latent correlation matrix from observed data of (possibly) mixed types (continuous/binary/truncated/ternary) based on the latent Gaussian copula model. Missing values (NA) are allowed. The estimation is based on pairwise complete observations.
-#' @rdname estR
-#' @aliases estR
+#' @rdname latentcor
+#' @aliases latentcor
 #' @param X A numeric data matrix (n by p), where n is number of samples, and p is number of variables. Missing values (NA) are allowed, in which case the estimation is based on pairwise complete observations.
 #' @param types A vector of length p indicating the type of each of the p variables in \code{X}. Each element must be one of \code{"con"} (continuous), \code{"bin"} (binary), \code{"tru"} (truncated) or \code{"ter"} (ternary). If the vector has length 1, then all p variables are assumed to be of the same type that is supplied. The default value is \code{"con"} which means all variables are continuous.
 #' @param method The calculation method for latent correlations. Either \code{"original"} or \code{"approx"}. If \code{method = "approx"}, multilinear approximation method is used, which is much faster than the original method, see Yoon et al. (2021). If \code{method = "original"}, optimization of the bridge inverse function is used. The default is \code{"approx"}.
@@ -9,7 +9,7 @@
 #' @param tol When \code{method = "original"}, specifies the desired accuracy of the bridge function inversion via uniroot optimization and is passed to \code{\link{optimize}}. The default value is 1e-8. When \code{method = "approx"}, this parameter is ignored.
 #' @param ratio When \code{method = "approx"}, specifies the boundary value for multilinear interpolation, must be between 0 and 1. When \code{ratio = 0}, no linear interpolation is performed (the slowest execution) which is equivalent to \code{method = "original"}. When \code{ratio = 1}, linear interpolation is always performed (the fastest execution) but may lead to high approximation errors. The default (recommended) value is 0.9 which controls the approximation error and has fast execution, see Yoon et al. (2021) for details. When \code{method = "original"}, this parameter is ignored.
[…]

### 2021-08-27 man/examples/latentcor_ex.R
Update function name to latentcor

@@ -1,0 +1,9 @@
+# Data generation
+X = GenData()$X
+# Estimate latent correlation matrix with original method
+R_nc_org = latentcor(X = X, types = c("ter", "con"), method = "original")$R
+# Estimate latent correlation matrix with aprroximation method
+R_nc_approx = latentcor(X = X, types = c("ter", "con"), method = "approx")$R
+# Heatmap for latent correlation matrix.
+Heatmap_R_nc_approx = latentcor(X = X, types = c("ter", "con"), method = "approx",
+                                showplot = TRUE)$plotR

### 2021-08-27 tests/testthat/test_latentcor.R

@@ -1,0 +1,146 @@
+
+library(latentcor)
+
+test_that("n_x is 0 if no tie", {
+  x = unique(rnorm(10))
+  expect_equal(n_x(x = x, length(x)), 0)
+})
+
+test_that("n_x is not 1 if tie", {
+  x = unique(rnorm(10)); x[3] = x[4]
+  expect_equal(n_x(x = x, length(x)), 1)
+})
+
+# Data generation
+test_that("latentcor is symmetric.", {
+  X = GenData(types = c("con", "con"), rhos = 1)$X
+  expect_equal(latentcor(X = X, types = c("con", "con"), method = "original", nu = 0.5, tol = 1e-8, ratio = .9, showplot = TRUE)$R,
+               t(latentcor(X = X, types = c("con", "con"), method = "original", nu = 0.5, tol = 1e-8, ratio = .9, showplot = FALSE)$R))
+})
+
+test_that("latentcor is symmetric.", {
+  X = GenData(types = c("con", "con"))$X; colnames(X) = c("var1", "var2")
+  expect_equal(latentcor(X = X, types = c("con", "con"), method = "approx", nu = 0.5, tol = 1e-8, ratio = .9, showplot = TRUE)$R,
+               t(latentcor(X = X, types = c("con", "con"), method = "approx", nu = 0.5, tol = 1e-8, ratio = .9, showplot = FALSE)$R))
+})
+
+test_that("latentcor is symmetric.", {
+  X = GenData(types = c("bin", "bin"), rhos = 1)$X
+  expect_equal(latentcor(X = X, types = c("bin", "bin"), method = "original", nu = 0.5, tol = 1e-8, ratio = .9, showplot = TRUE)$R,
+               t(latentcor(X = X, types = c("bin", "bin"), method = "original", nu = 0.5, tol = 1e-8, ratio = .9, showplot = FALSE)$R))
+})
+
+test_that("latentcor is symmetric.", {
+  X = GenData(types = c("bin", "bin"))$X; colnames(X) = c("var1", "var2")
+  expect_equal(latentcor(X = X, types = c("bin", "bin"), method = "approx", nu = 0.5, tol = 1e-8, ratio = .9, showplot = TRUE)$R,
+               t(latentcor(X = X, types = c("bin", "bin"), method = "approx", nu = 0.5, tol = 1e-8, ratio = .9, showplot = FALSE)$R))
+})
+
+test_that("latentcor is symmetric.", {
[…]

### 2021-08-27 R/latentcor.R
rebuild website

@@ -19,7 +19,7 @@
 #' }
 #'
 #' @references
-#' Fan J., Liu H., Ning Y. and Zou H. (2017) "High dimensional semiparametric latent graphicalmodel for mixed data" \doi{10.1111/rssb.12168}.
+#' Fan J., Liu H., Ning Y. and Zou H. (2017) "High dimensional semiparametric latent graphical model for mixed data" \doi{10.1111/rssb.12168}.
 #'
 #' Yoon G., Carroll R.J. and Gaynanova I. (2020) "Sparse semiparametric canonical correlation analysis for data of mixed types" \doi{10.1093/biomet/asaa007}.
 #'

### 2021-08-27 R/get_types.R
Almost done

@@ -1,0 +1,22 @@
+get_types = function(X) {
+  X = as.matrix(X);
+  n = nrow(X); p = ncol(X); X_num = matrix(NA, n, p); types = rep(NA, p)
+  for (i in 1:p) {
+    X_fac[ , i] = factor(X[ , i])
+    level = levels(X_fac[ , i])
+    X_num[ , i] = as.numeric(X_fac[ , i]) - 1
+    if (length(level) <= 1) {
+      stop("No variation in ", i, "th variable (", i, "th column of input data).")
+    } else if (length(level) == 2) {
+      types[i] = "bin"
+    } else if (length(level) == 3) {
+      types[i] = "ter"
+    } else {
+      if (length(X_num[ , i] == min(X_num[ , i])) / length(X_num[ , i]) > 0.1) {
+        types[i] = "tru"
+      } else {
+        types[i] = "con"
+      }
+    }
+  }
+}

### 2021-08-28 R/GenData.R
Update function names, finish get_types.

@@ -12,7 +12,7 @@
 #'
 #' @param showplot Logical indicator. If TRUE, generates the plot of the data when number of variables p is no more than 3. The default value is FALSE.
 #'
-#' @return \code{GenData} returns a list containing
+#' @return \code{gen_data} returns a list containing
 #' \itemize{
 #'       \item{X: }{Generated data matrix (n by p) of observed variables.}
 #'       \item{plotX: }{Visualization of the data matrix X.
@@ -31,11 +31,11 @@
 #' @examples
 #' # Generate single continuous variable with exponential transformation (always greater than 0)
 #' # and show histogram.
-#' simdata = GenData(n = 100, copulas = "expo", types = "con", showplot = FALSE)
+#' simdata = gen_data(n = 100, copulas = "expo", types = "con", showplot = FALSE)
 #' X = simdata$X; plotX = simdata$plotX
 #' # Generate a pair of variables (ternary and continuous) with default proportions
 #' # and without copula transformation.
-#' simdata = GenData()
+#' simdata = gen_data()
 #' X = simdata$X
 #' # Generate 3 variables (binary, ternary and truncated)
 #' # corresponding copulas for each variables are "no" (no transformation),
@@ -44,7 +44,7 @@
 #' # and 40% of ones, truncated variable has 50% of zeros.
 #' # Then show the 3D scatter plot (data points project on either 0 or 1 on Axis X1;
 #' # on 0, 1 or 2 on Axas X2; on positive domain on Axis X3)
-#' simdata = GenData(n = 100, rhos = c(.3, .4, .5), copulas = c("no", "cube", "cube"),
+#' simdata = gen_data(n = 100, rhos = c(.3, .4, .5), copulas = c("no", "cube", "cube"),
 #'           types = c("bin", "ter", "tru"), XP = list(.3, c(.2, .4), .5), showplot = TRUE)
 #' X = simdata$X; plotX = simdata$plotX
 #' # Check the proportion of zeros for the binary variable.
@@ -54,7 +54,7 @@
 #' # Check the proportion of zeros for the truncated variable.
 #' sum(simdata$X[ , 3] == 0)
 
-GenData = function(n = 100, types = c("ter", "con"), rhos = .5, copulas = "no", XP = NULL, showplot = FALSE) {
+gen_data = function(n = 100, types = c("ter", "con"), rhos = .5, copulas = "no", XP = NULL, showplot = FALSE) {
[…]

### 2021-08-28 R/get_types.R

@@ -1,22 +1,0 @@
-get_types = function(X) {
-  X = as.matrix(X);
-  n = nrow(X); p = ncol(X); X_num = matrix(NA, n, p); types = rep(NA, p)
-  for (i in 1:p) {
-    X_fac[ , i] = factor(X[ , i])
-    level = levels(X_fac[ , i])
-    X_num[ , i] = as.numeric(X_fac[ , i]) - 1
-    if (length(level) <= 1) {
-      stop("No variation in ", i, "th variable (", i, "th column of input data).")
-    } else if (length(level) == 2) {
-      types[i] = "bin"
-    } else if (length(level) == 3) {
-      types[i] = "ter"
-    } else {
-      if (length(X_num[ , i] == min(X_num[ , i])) / length(X_num[ , i]) > 0.1) {
-        types[i] = "tru"
-      } else {
-        types[i] = "con"
-      }
-    }
-  }
-}

### 2021-08-28 R/internal.R

@@ -28,6 +28,40 @@ fromZtoX = function(z, type, copula, xp) {
   return(x)
 }
 
+ord = function(x, ordinal) {
+  x_ord = x
+  for (i in 1:length(ordinal)) {
+    x_ord[x == ordinal[i]] = i - 1
+  }
+  return(x_ord)
+}
+
+get_types = function(X) {
+  n = nrow(X); p = ncol(X); types = rep(NA, p)
+  for (i in 1:p) {
+    level = unique(X[ , i])
+    if (length(level) <= 1) {
+      stop("No variation in ", i, "th variable (", i, "th column of input data).")
+    } else if (length(level) == 2) {
+      types[i] = "bin"
+    } else if (length(level) == 3) {
+      types[i] = "ter"
+    } else if (length(level) > 3 & length(level) <= 10) {
+      message("ordinal levels between 4 and 10 will be approximated by continuous type.")
+      types[i] = "con"
+    } else {
+      if (length(X[ , i] == min(X[ , i], rm.na = TRUE)) / length(X[ , i]) > 0.15) {
+        types[i] = "tru"
+      } else if (length(X[ , i] == max(X[ , i], rm.na = TRUE)) / length(X[ , i]) > 0.15) {
+        types[i] = "utr"
+      } else {
+        types[i] = "con"
+      }
+    }
+  }
+  return(list(X = X, types = types))
+}
+
 Kendalltau = function(X) {
   X = na.omit(X)
   n = nrow(X); n0 = n * (n - 1) / 2
@@ -53,22 +87,28 @@ n_x = function(x, n) {
 
 encodeX = function(X, types) {
   X = matrix(unlist(X), nrow = nrow(X), ncol = ncol(X))
+  negate = rep(1, length(types))
   colmin_mat = matrix(apply(X, 2, function(x) min(x, na.rm = TRUE)), nrow = nrow(X), ncol = ncol(X), byrow = TRUE)
   colmax_mat = matrix(apply(X, 2, function(x) max(x, na.rm = TRUE)), nrow = nrow(X), ncol = ncol(X), byrow = TRUE)
   if (sum(types == "bin") > 0) {
-  X_bin = X[ , types == "bin"]
-  X[ , types == "bin"][X_bin == colmin_mat[ , types == "bin"]] = 0; X[ , types == "bin"][X_bin == colmax_mat[ , types == "bin"]] = 1
[…]

### 2021-08-28 R/latentcor.R

@@ -3,12 +3,13 @@
 #' @rdname latentcor
 #' @aliases latentcor
 #' @param X A numeric data matrix (n by p), where n is number of samples, and p is number of variables. Missing values (NA) are allowed, in which case the estimation is based on pairwise complete observations.
-#' @param types A vector of length p indicating the type of each of the p variables in \code{X}. Each element must be one of \code{"con"} (continuous), \code{"bin"} (binary), \code{"tru"} (truncated) or \code{"ter"} (ternary). If the vector has length 1, then all p variables are assumed to be of the same type that is supplied. The default value is \code{"con"} which means all variables are continuous.
 #' @param method The calculation method for latent correlations. Either \code{"original"} or \code{"approx"}. If \code{method = "approx"}, multilinear approximation method is used, which is much faster than the original method, see Yoon et al. (2021). If \code{method = "original"}, optimization of the bridge inverse function is used. The default is \code{"approx"}.
 #' @param nu Shrinkage parameter for the correlation matrix, must be between 0 and 1. Guarantees that the minimal eigenvalue of returned correlation matrix is greater or equal to \code{nu}. When \code{nu = 0}, no shrinkage is performed, the returned correlation matrix will be semi-positive definite but not necessarily strictly positive definite. When \code{nu = 1}, the identity matrix is returned (not recommended).  The default (recommended) value is 0.001.
 #' @param tol When \code{method = "original"}, specifies the desired accuracy of the bridge function inversion via uniroot optimization and is passed to \code{\link{optimize}}. The default value is 1e-8. When \code{method = "approx"}, this parameter is ignored.
[…]

### 2021-08-28 tests/testthat/test_encodeX.R

@@ -2,6 +2,6 @@ library(latentcor)
 
 test_that("Always recode binary/ternary/truncated", {
   types = c("con", "bin", "tru", "ter")
-  X = GenData(types = types)$X; X[sample(length(X), 10)] = NA
-  expect_equal(encodeX(X = cbind(X[ , 1], X[ , 2:4] - 5), types = types), X)
+  X = gen_data(types = types)$X; X[sample(length(X), 10)] = NA
+  expect_equal(encodeX(X = cbind(X[ , 1], X[ , 2:4] - 5), types = types)$X, X)
 })

### 2021-08-28 tests/testthat/test_gen_data.R

@@ -16,13 +16,13 @@ test_that("ternary data should be zero, one or two.", {
 })
 
 test_that("binary data have lower bound zero.", {
-  expect_equal(sort(unique(c(as.matrix(GenData(types = "bin", copulas = "cube", XP = list(.5))$X)))), c(0, 1))
+  expect_equal(sort(unique(c(as.matrix(gen_data(types = "bin", copulas = "cube", XP = list(.5))$X)))), c(0, 1))
 })
 
 test_that("truncated data have lower bound zero.", {
-  expect_equal(min(as.matrix(GenData(types = "tru", copulas = "cube", XP = list(.5))$X)), 0)
+  expect_equal(min(as.matrix(gen_data(types = "tru", copulas = "cube", XP = list(.5))$X)), 0)
 })
 
 test_that("ternary data should be zero, one or two.", {
-  expect_equal(sort(unique(c(as.matrix(GenData(types = "ter", copulas = "cube", XP = list(c(.3, .5)))$X)))), c(0, 1, 2))
+  expect_equal(sort(unique(c(as.matrix(gen_data(types = "ter", copulas = "cube", XP = list(c(.3, .5)))$X)))), c(0, 1, 2))
 })

### 2021-08-28 tests/testthat/test_latentcor.R

@@ -13,134 +13,134 @@ test_that("n_x is not 1 if tie", {
 
 # Data generation
 test_that("latentcor is symmetric.", {
-  X = GenData(types = c("con", "con"), rhos = 1)$X
+  X = gen_data(types = c("con", "con"), rhos = 1)$X
   expect_equal(latentcor(X = X, types = c("con", "con"), method = "original", nu = 0.5, tol = 1e-8, ratio = .9, showplot = TRUE)$R,
                t(latentcor(X = X, types = c("con", "con"), method = "original", nu = 0.5, tol = 1e-8, ratio = .9, showplot = FALSE)$R))
 })
 
 test_that("latentcor is symmetric.", {
-  X = GenData(types = c("con", "con"))$X; colnames(X) = c("var1", "var2")
+  X = gen_data(types = c("con", "con"))$X; colnames(X) = c("var1", "var2")
   expect_equal(latentcor(X = X, types = c("con", "con"), method = "approx", nu = 0.5, tol = 1e-8, ratio = .9, showplot = TRUE)$R,
                t(latentcor(X = X, types = c("con", "con"), method = "approx", nu = 0.5, tol = 1e-8, ratio = .9, showplot = FALSE)$R))
 })
 
 test_that("latentcor is symmetric.", {
-  X = GenData(types = c("bin", "bin"), rhos = 1)$X
+  X = gen_data(types = c("bin", "bin"), rhos = 1)$X
   expect_equal(latentcor(X = X, types = c("bin", "bin"), method = "original", nu = 0.5, tol = 1e-8, ratio = .9, showplot = TRUE)$R,
                t(latentcor(X = X, types = c("bin", "bin"), method = "original", nu = 0.5, tol = 1e-8, ratio = .9, showplot = FALSE)$R))
 })
 
 test_that("latentcor is symmetric.", {
-  X = GenData(types = c("bin", "bin"))$X; colnames(X) = c("var1", "var2")
+  X = gen_data(types = c("bin", "bin"))$X; colnames(X) = c("var1", "var2")
   expect_equal(latentcor(X = X, types = c("bin", "bin"), method = "approx", nu = 0.5, tol = 1e-8, ratio = .9, showplot = TRUE)$R,
                t(latentcor(X = X, types = c("bin", "bin"), method = "approx", nu = 0.5, tol = 1e-8, ratio = .9, showplot = FALSE)$R))
 })
 
[…]

### 2021-08-28 tests/testthat/test_zratio.R

@@ -2,21 +2,21 @@
 library(latentcor)
 
 test_that("zratio is NULL if type is continuous", {
-  X = GenData(types = "con", copulas = "no")$X
+  X = gen_data(types = "con", copulas = "no")$X
   expect_true(is.na(unlist(zratios(X = X, types = "con"))))
 })
 
 test_that("zratio is the proportion of zeros if type is binary", {
-  X = GenData(types = "bin", copulas = "no")$X
+  X = gen_data(types = "bin", copulas = "no")$X
   expect_equal(unlist(zratios(X = X, types = "bin")), as.numeric(colMeans(as.matrix(X == 0))))
 })
 
 test_that("zratio is the proportion of zeros if type is truncated", {
-  X = GenData(types = "tru", copulas = "no")$X
+  X = gen_data(types = "tru", copulas = "no")$X
   expect_equal(unlist(zratios(X = X, types = "tru")), as.numeric(colMeans(as.matrix(X == 0))))
 })
 
 test_that("zratio is the proportion of zeros and ones if type is ternary", {
-  X = GenData(types = "ter", copulas = "no")$X
+  X = gen_data(types = "ter", copulas = "no")$X
   expect_equal(unlist(zratios(X = X, types = "ter")), as.numeric(c(colMeans(as.matrix(X == 0)), 1 - colMeans(as.matrix(X == 2)))))
 })

### 2021-08-30 R/latentcor.R
Mover types before method.

@@ -37,7 +37,7 @@
 #' @export
 #' @example man/examples/latentcor_ex.R
 
-latentcor = function(X, method = c("approx", "original"), nu = 0.001, tol = 1e-8, ratio = 0.9, showplot = FALSE, types = NULL, ordinals = NULL){
+latentcor = function(X, types = NULL, method = c("approx", "original"), nu = 0.001, tol = 1e-8, ratio = 0.9, showplot = FALSE, ordinals = NULL){
   if(nu < 0 | nu > 1){
     stop("nu must be be between 0 and 1.")
   } else if(tol <= 0) {

### 2021-08-31 R/internal.R
Make a pre_process function to deal with data types and non-numeric cases.

@@ -36,32 +36,6 @@ ord = function(x, ordinal) {
   return(x_ord)
 }
 
-get_types = function(X) {
-  n = nrow(X); p = ncol(X); types = rep(NA, p)
-  for (i in 1:p) {
-    level = unique(X[ , i])
-    if (length(level) <= 1) {
-      stop("No variation in ", i, "th variable (", i, "th column of input data).")
-    } else if (length(level) == 2) {
-      types[i] = "bin"
-    } else if (length(level) == 3) {
-      types[i] = "ter"
-    } else if (length(level) > 3 & length(level) <= 10) {
-      message("ordinal levels between 4 and 10 will be approximated by continuous type.")
-      types[i] = "con"
-    } else {
-      if (length(X[ , i] == min(X[ , i], rm.na = TRUE)) / length(X[ , i]) > 0.15) {
-        types[i] = "tru"
-      } else if (length(X[ , i] == max(X[ , i], rm.na = TRUE)) / length(X[ , i]) > 0.15) {
-        types[i] = "utr"
-      } else {
-        types[i] = "con"
-      }
-    }
-  }
-  return(list(X = X, types = types))
-}
-
 Kendalltau = function(X) {
   X = na.omit(X)
   n = nrow(X); n0 = n * (n - 1) / 2

### 2021-08-31 R/latentcor.R

@@ -8,8 +8,7 @@
 #' @param tol When \code{method = "original"}, specifies the desired accuracy of the bridge function inversion via uniroot optimization and is passed to \code{\link{optimize}}. The default value is 1e-8. When \code{method = "approx"}, this parameter is ignored.
 #' @param ratio When \code{method = "approx"}, specifies the boundary value for multilinear interpolation, must be between 0 and 1. When \code{ratio = 0}, no linear interpolation is performed (the slowest execution) which is equivalent to \code{method = "original"}. When \code{ratio = 1}, linear interpolation is always performed (the fastest execution) but may lead to high approximation errors. The default (recommended) value is 0.9 which controls the approximation error and has fast execution, see Yoon et al. (2021) for details. When \code{method = "original"}, this parameter is ignored.
 #' @param showplot Logical indicator. \code{showplot = TRUE} generates a ggplot object \code{plotR} with the heatmap of latent correlation matrix \code{R}. \code{plotR = NULL} if \code{showplot = FALSE}.
-#' @param types A vector of length p indicating the type of each of the p variables in \code{X}. Each element must be one of \code{"con"} (continuous), \code{"bin"} (binary), \code{"tru"} (truncated) or \code{"ter"} (ternary). If the vector has length 1, then all p variables are assumed to be of the same type that is supplied. The default value is \code{"con"} which means all variables are continuous.
-#' @param ordinals A list to convert characters into ordinal numbers.
+#' @param types A vector of length p indicating the type of each of the p variables in \code{X}. Each element must be one of \code{"con"} (continuous), \code{"bin"} (binary), \code{"ter"} (ternary), \code{"tru"} (truncated) or \code{"utr"}(upper truncated). If the vector has length 1, then all p variables are assumed to be of the same type that is supplied. The default value is \code{"con"} which means all variables are continuous.
 #' @return \code{latentcor} returns
 #' \itemize{
 #'       \item{zratios: }{A list of of length p corresponding to each variable. Returns NA for continuous variable; proportion of zeros for binary/truncated variables; the cumulative proportions of zeros and ones (e.g. first value is proportion of zeros, second value is proportion of zeros and ones) for ternary variable. }
[…]

### 2021-08-31 R/pre_process.R

@@ -1,0 +1,48 @@
+#' Pre-process of raw data to deal with non-numeric variable, then get suggested variable types.
+#'
+#' @param raw_data A data frame or data matrix (n by p), where n is number of samples, and p is number of variables. Missing values (NA) are allowed, in which case the estimation is based on pairwise complete observations. Non-numeric variables are allowed, in which case users should provide ordinal orders of non-numeric values by \code{ordinals}.
+#' @param ordinals A list to convert characters into ordinal numbers. Each element of the list is corresponding to each variable. For the numeric variable, please put NA; for the non-numeric variable, please put the character in ascending order to be converted to ordinal variable. The default value is NULL for the case that all variables are numeric. See vignette for detail.
+#' @return \code{pre_process} returns
+#' \itemize{
+#'       \item{X: }{A numeric data matrix (n by p), where n is number of samples, and p is number of variables. Missing values (NA) are allowed, in which case the estimation is based on pairwise complete observations.}
+#'       \item{types: }{A vector of length p indicating the type of each of the p variables in \code{X}. Each element must be one of \code{"con"} (continuous), \code{"bin"} (binary), \code{"ter"} (ternary), \code{"tru"} (truncated) or \code{"utr"} (upper truncated). If the vector has length 1, then all p variables are assumed to be of the same type that is supplied. The default value is \code{"con"} which means all variables are continuous.}
+#' }
+#' @export
+#' @examples
+#' raw_data = data.frame(matrix(NA, 3, 3))
+#' raw_data$X1 = 1:3
+#' raw_data$X2 = c("medium", "small", "large")
+#' raw_data$X3 = 7:9
+#' pre_process = pre_process(raw_data = raw_data, ordinals = list(NA, c("small", "medium", "large")))
+#' pre_process$X
+#' pre_process$types
+#'
+
+pre_process = function(raw_data, ordinals = NULL) {
+  X = data.frame(raw_data); n = nrow(X); p = ncol(X); types = rep(NA, p)
+  for (i in 1:p) {
+    if (!(is.numeric(X[ , i]))) {
+      X[ , i] = ord(X[ , i], ordinals[[i]])
+    }
+    level = unique(X[ , i])
+    if (length(level) <= 1) {
+      stop("No variation in ", i, "th variable (", i, "th column of input data).")
+    } else if (length(level) == 2) {
+      types[i] = "bin"
[…]

### 2021-08-31 R/latentcor.R
make pre_process function and create an example for it.

@@ -3,12 +3,13 @@
 #' @rdname latentcor
 #' @aliases latentcor
 #' @param X A numeric data matrix (n by p), where n is number of samples, and p is number of variables. Missing values (NA) are allowed, in which case the estimation is based on pairwise complete observations.
+#' @param types A vector of length p indicating the type of each of the p variables in \code{X}. Each element must be one of \code{"con"} (continuous), \code{"bin"} (binary), \code{"ter"} (ternary), \code{"tru"} (truncated) or \code{"utr"}(upper truncated). If the vector has length 1, then all p variables are assumed to be of the same type that is supplied. The default value is \code{"con"} which means all variables are continuous.
 #' @param method The calculation method for latent correlations. Either \code{"original"} or \code{"approx"}. If \code{method = "approx"}, multilinear approximation method is used, which is much faster than the original method, see Yoon et al. (2021). If \code{method = "original"}, optimization of the bridge inverse function is used. The default is \code{"approx"}.
 #' @param nu Shrinkage parameter for the correlation matrix, must be between 0 and 1. Guarantees that the minimal eigenvalue of returned correlation matrix is greater or equal to \code{nu}. When \code{nu = 0}, no shrinkage is performed, the returned correlation matrix will be semi-positive definite but not necessarily strictly positive definite. When \code{nu = 1}, the identity matrix is returned (not recommended).  The default (recommended) value is 0.001.
 #' @param tol When \code{method = "original"}, specifies the desired accuracy of the bridge function inversion via uniroot optimization and is passed to \code{\link{optimize}}. The default value is 1e-8. When \code{method = "approx"}, this parameter is ignored.
 #' @param ratio When \code{method = "approx"}, specifies the boundary value for multilinear interpolation, must be between 0 and 1. When \code{ratio = 0}, no linear interpolation is performed (the slowest execution) which is equivalent to \code{method = "original"}. When \code{ratio = 1}, linear interpolation is always performed (the fastest execution) but may lead to high approximation errors. The default (recommended) value is 0.9 which controls the approximation error and has fast execution, see Yoon et al. (2021) for details. When \code{method = "original"}, this parameter is ignored.
 #' @param showplot Logical indicator. \code{showplot = TRUE} generates a ggplot object \code{plotR} with the heatmap of latent correlation matrix \code{R}. \code{plotR = NULL} if \code{showplot = FALSE}.
[…]

### 2021-08-31 R/pre_process.R

@@ -16,7 +16,6 @@
 #' pre_process = pre_process(raw_data = raw_data, ordinals = list(NA, c("small", "medium", "large")))
 #' pre_process$X
 #' pre_process$types
-#'
 
 pre_process = function(raw_data, ordinals = NULL) {
   X = data.frame(raw_data); n = nrow(X); p = ncol(X); types = rep(NA, p)
@@ -44,5 +43,5 @@ pre_process = function(raw_data, ordinals = NULL) {
       }
     }
   }
-  return(list(X = as.matrix(X), types = types))
+  return(list(X = data.matrix(X), types = types))
 }

### 2021-08-31 R/internal.R
fix bugs

@@ -29,7 +29,7 @@ fromZtoX = function(z, type, copula, xp) {
 }
 
 ord = function(x, ordinal) {
-  x_ord = x
+  x_ord = length(x)
   for (i in 1:length(ordinal)) {
     x_ord[x == ordinal[i]] = i - 1
   }

### 2021-08-31 R/pre_process.R

@@ -13,15 +13,19 @@
 #' raw_data$X1 = 1:3
 #' raw_data$X2 = c("medium", "small", "large")
 #' raw_data$X3 = 7:9
-#' pre_process = pre_process(raw_data = raw_data, ordinals = list(NA, c("small", "medium", "large")))
+#' pre_process = pre_process(raw_data = raw_data,
+#'               ordinals = list(NA, c("small", "medium", "large"), NA))
 #' pre_process$X
 #' pre_process$types
 
 pre_process = function(raw_data, ordinals = NULL) {
-  X = data.frame(raw_data); n = nrow(X); p = ncol(X); types = rep(NA, p)
+  raw_data = data.frame(raw_data); n = nrow(raw_data); p = ncol(raw_data); types = rep(NA, p)
+  X = matrix(NA, n, p)
   for (i in 1:p) {
-    if (!(is.numeric(X[ , i]))) {
-      X[ , i] = ord(X[ , i], ordinals[[i]])
+    if (is.numeric(raw_data[ , i])) {
+      X[ , i] = raw_data[ , i]
+    } else {
+      X[ , i] = ord(raw_data[ , i], ordinals[[i]])
     }
     level = unique(X[ , i])
     if (length(level) <= 1) {
@@ -43,5 +47,6 @@ pre_process = function(raw_data, ordinals = NULL) {
       }
     }
   }
-  return(list(X = data.matrix(X), types = types))
+  colnames(X) = colnames(raw_data)
+  return(list(X = X, types = types))
 }

### 2021-08-31 tests/testthat/test_pre_process.R

@@ -1,0 +1,11 @@
+library(latentcor)
+
+test_that("Test non-numeric conversion and types generation", {
+  raw_data = data.frame(matrix(NA, 3, 3))
+  raw_data$X1 = 1:3
+  raw_data$X2 = c("medium", "small", "large")
+  raw_data$X3 = 7:9
+  pre_process = pre_process(raw_data = raw_data, ordinals = list(NA, c("small", "medium", "large"), NA))
+  expect_equal(pre_process$X, data.matrix(data.frame(matrix(c(1, 2, 3, 1, 0, 2, 7, 8, 9), 3, 3))))
+  expect_equal(pre_process$types, c("ter", "ter", "ter"))
+})

### 2021-08-31 R/latentcor.R
Adding details to latentcor documentation

@@ -10,6 +10,16 @@
 #' @param ratio When \code{method = "approx"}, specifies the boundary value for multilinear interpolation, must be between 0 and 1. When \code{ratio = 0}, no linear interpolation is performed (the slowest execution) which is equivalent to \code{method = "original"}. When \code{ratio = 1}, linear interpolation is always performed (the fastest execution) but may lead to high approximation errors. The default (recommended) value is 0.9 which controls the approximation error and has fast execution, see Yoon et al. (2021) for details. When \code{method = "original"}, this parameter is ignored.
 #' @param showplot Logical indicator. \code{showplot = TRUE} generates a ggplot object \code{plotR} with the heatmap of latent correlation matrix \code{R}. \code{plotR = NULL} if \code{showplot = FALSE}.
 #' @param ordinals A list to convert characters into ordinal numbers. Each element of the list is corresponding to each variable. For the numeric variable, please put NA; for the non-numeric variable, please put the character in ascending order to be converted to ordinal variable. The default value is NULL for the case that all variables are numeric. See vignette for detail.
+#' @details
+#' The function estimates latent correlation by calculating inverse bridge function (Fan et al., 2017) evaluated at the value of sample Kendall's tau (\eqn{\hat \tau}). The bridge function F connects Kendall's tau to latent correlation r so that \eqn{F(r) = E(\hat \tau)}. The form of function F depends on the variable types (continuous/binary/truncated/ternary), but is exact. The exact form of inverse is not available, so has to be evaluated numerically for each pair of variables leading to \code{Rpointwise}.
+#'
+#' When \code{method = "original"}, the inversion is done numerically by solving \deqn{minimize_r (F(r) - \hat \tau)^2} using \code{\link{optimize}}. The parameter \code{tol} is used to control the accuracy of the solution.
+#'
+#' When \code{method = "approx"}, the inversion is done approximately by interpolating previously calculated and stored values of \eqn{F^{-1}(\hat \tau)}. This is significantly faster than the original method (Yoon et al., 2021) for binary/ternary/truncated cases, however the approximation errors may be non-negligible on some regions of the space. The parameter \code{ratio} controls the region where the interpolation is performed, with original inversion method applied outside of that region, see Yoon et al. 2021 and vignette for more details.
+#'
[…]

### 2021-09-01 R/get_types.R
Minor document clarification on get_types

@@ -1,6 +1,6 @@
-#' Get types of each variable in data matrix.
+#' Automatically determine types of each variable (continuous/binary/ternary/truncated) in a data matrix.
 #'
-#' @param X A numeric data matrix (n by p), where n is number of samples, and p is number of variables. Missing values (NA) are allowed, in which case the estimation is based on pairwise complete observations.
+#' @param X A numeric data matrix (n by p), where n is number of samples, and p is number of variables. Missing values (NA) are allowed.
 #' @return \code{get_types} returns
 #' \itemize{
 #'       \item{types: }{A vector of length p indicating the type of each of the p variables in \code{X}. Each element must be one of \code{"con"} (continuous), \code{"bin"} (binary), \code{"ter"} (ternary) or \code{"tru"} (truncated).}
@@ -13,7 +13,7 @@
 get_types = function(X) {
   X = data.matrix(X)
   if (!(is.numeric(X))) {
-    stop("Input data matrix should be numerical.")
+    stop("Input data matrix should be numeric.")
   }
   n = nrow(X); p = ncol(X); types = rep(NA, p)
   for (i in 1:p) {

### 2021-09-01 R/latentcor.R
Add documentation for default type NULL.

@@ -3,7 +3,7 @@
 #' @rdname latentcor
 #' @aliases latentcor
 #' @param X A numeric data matrix (n by p), where n is number of samples, and p is number of variables. Missing values (NA) are allowed, in which case the estimation is based on pairwise complete observations.
-#' @param types A vector of length p indicating the type of each of the p variables in \code{X}. Each element must be one of \code{"con"} (continuous), \code{"bin"} (binary), \code{"ter"} (ternary) or \code{"tru"} (truncated). If the vector has length 1, then all p variables are assumed to be of the same type that is supplied. The default value is \code{"con"} which means all variables are continuous.
+#' @param types A vector of length p indicating the type of each of the p variables in \code{X}. Each element must be one of \code{"con"} (continuous), \code{"bin"} (binary), \code{"ter"} (ternary) or \code{"tru"} (truncated). If the vector has length 1, then all p variables are assumed to be of the same type that is supplied. The default value is \code{"con"} which means all variables are continuous. The default value NULL to call \code{get_types} to generate recommended types automatically.
 #' @param method The calculation method for latent correlations. Either \code{"original"} or \code{"approx"}. If \code{method = "approx"}, multilinear approximation method is used, which is much faster than the original method, see Yoon et al. (2021). If \code{method = "original"}, optimization of the bridge inverse function is used. The default is \code{"approx"}.
 #' @param nu Shrinkage parameter for the correlation matrix, must be between 0 and 1. Guarantees that the minimal eigenvalue of returned correlation matrix is greater or equal to \code{nu}. When \code{nu = 0}, no shrinkage is performed, the returned correlation matrix will be semi-positive definite but not necessarily strictly positive definite. When \code{nu = 1}, the identity matrix is returned (not recommended).  The default (recommended) value is 0.001.
 #' @param tol When \code{method = "original"}, specifies the desired accuracy of the bridge function inversion via uniroot optimization and is passed to \code{\link{optimize}}. The default value is 1e-8. When \code{method = "approx"}, this parameter is ignored.

### 2021-09-01 R/latentcor.R
Correct the bug in latentcor if types = NULL

@@ -3,7 +3,7 @@
 #' @rdname latentcor
 #' @aliases latentcor
 #' @param X A numeric data matrix (n by p), where n is number of samples, and p is number of variables. Missing values (NA) are allowed, in which case the estimation is based on pairwise complete observations.
-#' @param types A vector of length p indicating the type of each of the p variables in \code{X}. Each element must be one of \code{"con"} (continuous), \code{"bin"} (binary), \code{"ter"} (ternary) or \code{"tru"} (truncated). If the vector has length 1, then all p variables are assumed to be of the same type that is supplied. The default value is \code{"con"} which means all variables are continuous.
+#' @param types A vector of length p indicating the type of each of the p variables in \code{X}. Each element must be one of \code{"con"} (continuous), \code{"bin"} (binary), \code{"ter"} (ternary) or \code{"tru"} (truncated). If the vector has length 1, then all p variables are assumed to be of the same type that is supplied. The default value is \code{NULL}, and the variable types are determined automatically using function \code{\link{get_types}}. As automatic determination of variable types takes extra time, it is recommended to supply the types explicitly when they are known in advance.
 #' @param method The calculation method for latent correlations. Either \code{"original"} or \code{"approx"}. If \code{method = "approx"}, multilinear approximation method is used, which is much faster than the original method, see Yoon et al. (2021). If \code{method = "original"}, optimization of the bridge inverse function is used. The default is \code{"approx"}.
 #' @param nu Shrinkage parameter for the correlation matrix, must be between 0 and 1. Guarantees that the minimal eigenvalue of returned correlation matrix is greater or equal to \code{nu}. When \code{nu = 0}, no shrinkage is performed, the returned correlation matrix will be semi-positive definite but not necessarily strictly positive definite. When \code{nu = 1}, the identity matrix is returned (not recommended).  The default (recommended) value is 0.001.
 #' @param tol When \code{method = "original"}, specifies the desired accuracy of the bridge function inversion via uniroot optimization and is passed to \code{\link{optimize}}. The default value is 1e-8. When \code{method = "approx"}, this parameter is ignored.
@@ -48,28 +48,37 @@
 #' @example man/examples/latentcor_ex.R
 
[…]

### 2021-09-01 R/get_types.R
Adjust logic of automatic truncated type assignment to only focus on zero-inflation

@@ -1,16 +1,19 @@
 #' Automatically determine types of each variable (continuous/binary/ternary/truncated) in a data matrix.
 #'
 #' @param X A numeric data matrix (n by p), where n is number of samples, and p is number of variables. Missing values (NA) are allowed.
+#' @param tru_prop A scalar between 0 and 1 indicating the minimal proportion of zeros that should be present in a variable to be treated as \code{"tru"} (truncated type or zero-inflated) rather than as \code{"con"} (continuous type). The default value is 0.05 (any variable with more than 5\% of zero values among n samples is treated as truncated or zero-inflated)
+#'
+#'
 #' @return \code{get_types} returns
 #' \itemize{
-#'       \item{types: }{A vector of length p indicating the type of each of the p variables in \code{X}. Each element must be one of \code{"con"} (continuous), \code{"bin"} (binary), \code{"ter"} (ternary) or \code{"tru"} (truncated).}
+#'       \item{types: }{A vector of length p indicating the type of each of the p variables in \code{X}. Each element is one of \code{"con"} (continuous), \code{"bin"} (binary), \code{"ter"} (ternary) or \code{"tru"} (truncated).}
 #' }
 #' @export
 #' @examples
 #' X = gen_data(types = c("ter", "con"))$X
-#' get_types(X = X)
-
-get_types = function(X) {
+#' get_types(X)
+#'
+get_types = function(X, tru_prop = 0.05) {
   X = data.matrix(X)
   if (!(is.numeric(X))) {
     stop("Input data matrix should be numeric.")
@@ -20,22 +23,24 @@ get_types = function(X) {
   # Determine the type of each variable
   for (i in 1:p) {
     # Extract the number of unique levels
-    level = unique(X[ , i])
-    if (length(level) <= 1) {
+    levels = unique(X[ , i])
+    if (length(levels) <= 1) {
       # Only one level means no variation
       stop("No variation in ", i, "th variable (", i, "th column of input data).")
-    } else if (length(level) == 2) {
+    } else if (length(levels) == 2) {
       # Two levels means binary
       types[i] = "bin"
-    } else if (length(level) == 3) {
+    } else if (length(levels) == 3) {
       # Three levels means ternary
       types[i] = "ter"
-    } else if (length(level) > 3 & length(level) <= 10) {
-      # Since ordinals of more than 3 levels are not supported, alert the user those are treated as continuous
[…]

### 2021-09-01 R/latentcor.R
Add timing example to latentcor documentation

@@ -4,7 +4,7 @@
 #' @aliases latentcor
 #' @param X A numeric data matrix (n by p), where n is number of samples, and p is number of variables. Missing values (NA) are allowed, in which case the estimation is based on pairwise complete observations.
 #' @param types A vector of length p indicating the type of each of the p variables in \code{X}. Each element must be one of \code{"con"} (continuous), \code{"bin"} (binary), \code{"ter"} (ternary) or \code{"tru"} (truncated). If the vector has length 1, then all p variables are assumed to be of the same type that is supplied. The default value is \code{NULL}, and the variable types are determined automatically using function \code{\link{get_types}}. As automatic determination of variable types takes extra time, it is recommended to supply the types explicitly when they are known in advance.
-#' @param method The calculation method for latent correlations. Either \code{"original"} or \code{"approx"}. If \code{method = "approx"}, multilinear approximation method is used, which is much faster than the original method, see Yoon et al. (2021). If \code{method = "original"}, optimization of the bridge inverse function is used. The default is \code{"approx"}.
+#' @param method The calculation method for latent correlations. Either \code{"original"} or \code{"approx"}. If \code{method = "approx"}, multilinear approximation method is used, which is much faster than the original method, see Yoon et al. (2021) for timing comparisons for various variable types. If \code{method = "original"}, optimization of the bridge inverse function is used. The default is \code{"approx"}.
 #' @param nu Shrinkage parameter for the correlation matrix, must be between 0 and 1. Guarantees that the minimal eigenvalue of returned correlation matrix is greater or equal to \code{nu}. When \code{nu = 0}, no shrinkage is performed, the returned correlation matrix will be semi-positive definite but not necessarily strictly positive definite. When \code{nu = 1}, the identity matrix is returned (not recommended).  The default (recommended) value is 0.001.
 #' @param tol When \code{method = "original"}, specifies the desired accuracy of the bridge function inversion via uniroot optimization and is passed to \code{\link{optimize}}. The default value is 1e-8. When \code{method = "approx"}, this parameter is ignored.
[…]

### 2021-09-01 man/examples/latentcor_ex.R

@@ -1,9 +1,21 @@
-# Data generation
+# Example 1 - truncated data type, same type for all variables
+# Generate data
+X = gen_data(n = 300, types = rep("tru", 5))$X
+# Estimate latent correlation matrix with original method and check the timing
+start_time = proc.time()
+R_nc_org = latentcor(X = X, types = "tru", method = "original")$R
+proc.time() - start_time
+# Estimate latent correlation matrix with approximation method and check the timing
+start_time = proc.time()
+R_nc_approx = latentcor(X = X, types = "tru", method = "approx")$R
+proc.time() - start_time
+# Heatmap for latent correlation matrix.
+Heatmap_R_nc_approx = latentcor(X = X, types = "tru", method = "approx",
+                                showplot = TRUE)$plotR
+
+# Example 2 - ternary/continuous case
 X = gen_data()$X
 # Estimate latent correlation matrix with original method
 R_nc_org = latentcor(X = X, types = c("ter", "con"), method = "original")$R
 # Estimate latent correlation matrix with aprroximation method
 R_nc_approx = latentcor(X = X, types = c("ter", "con"), method = "approx")$R
-# Heatmap for latent correlation matrix.
-Heatmap_R_nc_approx = latentcor(X = X, types = c("ter", "con"), method = "approx",
-                                showplot = TRUE)$plotR

### 2021-09-03 R/latentcor.R
Both numeric matrix and data frame are OK

@@ -2,7 +2,7 @@
 #' @description Estimation of latent correlation matrix from observed data of (possibly) mixed types (continuous/binary/truncated/ternary) based on the latent Gaussian copula model. Missing values (NA) are allowed. The estimation is based on pairwise complete observations.
 #' @rdname latentcor
 #' @aliases latentcor
-#' @param X A numeric data matrix (n by p), where n is number of samples, and p is number of variables. Missing values (NA) are allowed, in which case the estimation is based on pairwise complete observations.
+#' @param X A numeric matrix or numeric data frame (n by p), where n is number of samples, and p is number of variables. Missing values (NA) are allowed, in which case the estimation is based on pairwise complete observations.
 #' @param types A vector of length p indicating the type of each of the p variables in \code{X}. Each element must be one of \code{"con"} (continuous), \code{"bin"} (binary), \code{"ter"} (ternary) or \code{"tru"} (truncated). If the vector has length 1, then all p variables are assumed to be of the same type that is supplied. The default value is \code{NULL}, and the variable types are determined automatically using function \code{\link{get_types}}. As automatic determination of variable types takes extra time, it is recommended to supply the types explicitly when they are known in advance.
 #' @param method The calculation method for latent correlations. Either \code{"original"} or \code{"approx"}. If \code{method = "approx"}, multilinear approximation method is used, which is much faster than the original method, see Yoon et al. (2021) for timing comparisons for various variable types. If \code{method = "original"}, optimization of the bridge inverse function is used. The default is \code{"approx"}.
 #' @param nu Shrinkage parameter for the correlation matrix, must be between 0 and 1. Guarantees that the minimal eigenvalue of returned correlation matrix is greater or equal to \code{nu}. When \code{nu = 0}, no shrinkage is performed, the returned correlation matrix will be semi-positive definite but not necessarily strictly positive definite. When \code{nu = 1}, the identity matrix is returned (not recommended).  The default (recommended) value is 0.001.
@@ -57,11 +57,10 @@ latentcor = function(X, types = NULL, method = c("approx", "original"), nu = 0.0
     stop("ratio must be between 0 and 1.")
   }
 
-  X = data.matrix(X)
   if (!(is.numeric(X))) {
     stop("Input data matrix should be numeric.")
   }
-  p = ncol(X); name = colnames(X)
+  X = data.matrix(X); p = ncol(X); name = colnames(X)
 
 
   if (is.null(types)) {

### 2021-09-03 R/latentcor.R
check data type rather than coerce into numeric directly.

@@ -57,11 +57,11 @@ latentcor = function(X, types = NULL, method = c("approx", "original"), nu = 0.0
     stop("ratio must be between 0 and 1.")
   }
 
+  X = as.matrix(X)
   if (!(is.numeric(X))) {
     stop("Input data matrix should be numeric.")
   }
-  X = data.matrix(X); p = ncol(X); name = colnames(X)
-
+  p = ncol(X); name = colnames(X)
 
   if (is.null(types)) {
     # if variable types are not supplied, get the types by applying get_types

### 2021-09-03 R/get_types.R
Split vignette into two parts.

@@ -14,7 +14,7 @@
 #' get_types(X)
 #'
 get_types = function(X, tru_prop = 0.05) {
-  X = data.matrix(X)
+  X = as.matrix(X)
   if (!(is.numeric(X))) {
     stop("Input data matrix should be numeric.")
   }

### 2021-09-13 supplement/unbias_check.R
Generate plot

@@ -100,3 +100,41 @@ for (tp1 in c("con", "bin", "ter", "tru")) {
 
 pdf(file = "", width = , height = )
 dev.off()
+
+## Speed comparison
+# p = 20
+types_20 = rep(c("con", "bin", "ter", "tru"), 5)
+X_20 = gen_data(types = types_20)$X
+# p = 40
+types_40 = rep(c("con", "bin", "ter", "tru"), 10)
+X_40 = gen_data(types = types_40)$X
+# p = 100
+types_100 = rep(c("con", "bin", "ter", "tru"), 25)
+X_100 = gen_data(types = types_100)$X
+# p = 200
+types_200 = rep(c("con", "bin", "ter", "tru"), 50)
+X_200 = gen_data(types = types_200)$X
+# p = 400
+types_400 = rep(c("con", "bin", "ter", "tru"), 100)
+X_400 = gen_data(types = types_400)$X
+
+library(microbenchmark)
+library(latentcor)
+library(parallel)
+library(doFuture)
+types = list(types_20, types_40, types_100, types_200, types_400)
+X = list(X_20, X_40, X_100, X_200, X_400)
+md = c("original", "approx")
+registerDoFuture()
+plan(multicore, workers = detectCores())
+timing =
+foreach (i = 1:length(types), .combine = rbind) %dopar%  {
+  value = rep(NA, 2)
+  for (j in 1:length(md)) {
+    value[j] = median(microbenchmark(latentcor(X = X[[i]], types = types[[i]], method = md[j]), times = 5L)$time)
+  }
+  timing_vector = value
+}
+df = data.frame(c("log10(20)", "log10(40)", "log10(100)", "log10(200)", "log10(400)"), log10(cbind(timing)))
+print(ggplot(df, aes(x = df[ , 1])) + geom_line(y = df[ , 2], color = "blue") + geom_line(y = df[ , 3], color = "red")
+      + ggtitle("Speed comparison") + xlab("log10 of dimension") + ylab("log10 of computation times"))

### 2021-09-13 supplement/unbias_check.R
Create speed plot

@@ -102,6 +102,8 @@ pdf(file = "", width = , height = )
 dev.off()
 
 ## Speed comparison
+library(microbenchmark)
+library(latentcor)
 # p = 20
 types_20 = rep(c("con", "bin", "ter", "tru"), 5)
 X_20 = gen_data(types = types_20)$X
@@ -118,23 +120,20 @@ X_200 = gen_data(types = types_200)$X
 types_400 = rep(c("con", "bin", "ter", "tru"), 100)
 X_400 = gen_data(types = types_400)$X
 
-library(microbenchmark)
-library(latentcor)
-library(parallel)
-library(doFuture)
-types = list(types_20, types_40, types_100, types_200, types_400)
-X = list(X_20, X_40, X_100, X_200, X_400)
-md = c("original", "approx")
-registerDoFuture()
-plan(multicore, workers = detectCores())
-timing =
-foreach (i = 1:length(types), .combine = rbind) %dopar%  {
-  value = rep(NA, 2)
-  for (j in 1:length(md)) {
-    value[j] = median(microbenchmark(latentcor(X = X[[i]], types = types[[i]], method = md[j]), times = 5L)$time)
-  }
-  timing_vector = value
-}
-df = data.frame(c("log10(20)", "log10(40)", "log10(100)", "log10(200)", "log10(400)"), log10(cbind(timing)))
-print(ggplot(df, aes(x = df[ , 1])) + geom_line(y = df[ , 2], color = "blue") + geom_line(y = df[ , 3], color = "red")
-      + ggtitle("Speed comparison") + xlab("log10 of dimension") + ylab("log10 of computation times"))
+timing = microbenchmark(latentcor(X = X_20, types = types_20, method = "original"), latentcor(X = X_20, types = types_20),
+                        latentcor(X = X_40, types = types_40, method = "original"), latentcor(X = X_40, types = types_40),
+                        latentcor(X = X_100, types = types_100, method = "original"), latentcor(X = X_100, types = types_100),
+                        latentcor(X = X_200, types = types_200, method = "original"), latentcor(X = X_200, types = types_200),
+                        latentcor(X = X_400, types = types_400, method = "original"), latentcor(X = X_400, types = types_400), times = 5L, unit = "s")
+time_mat = matrix(print(timing)$median, ncol = 2, byrow = TRUE)
+dim = rep(c(log10(20), log10(40), log10(100), log10(200), log10(400)), each = 2)
[…]

### 2021-09-13 supplement/unbias_check.R
Done with speed comparison section in vignette latentcor_math

@@ -134,6 +134,6 @@ timing_plot = print(ggplot(time, aes(x = dim, y = time, color = method, group = 
       + ggtitle("Speed comparison") + xlab("log10 of dimension") + ylab("log10 of computation times (in seconds)")
 + scale_x_continuous(breaks=c(log10(20), log10(40), log10(100), log10(200), log10(400)), labels=c("log10(20)", "log10(40)", "log10(100)", "log10(200)", "log10(400)")))
 save(time, file = "timing.rda")
-pdf(file = "timing_plot.pdf", width = 7, height = 5)
+pdf(file = "./supplement/timing_plot.pdf", width = 21, height = 15)
 timing_plot
 dev.off()

### 2021-10-26 R/latentcor.R
Add use.nearPD as user-defined option.

@@ -5,6 +5,7 @@
 #' @param X A numeric matrix or numeric data frame (n by p), where n is number of samples, and p is number of variables. Missing values (NA) are allowed, in which case the estimation is based on pairwise complete observations.
 #' @param types A vector of length p indicating the type of each of the p variables in \code{X}. Each element must be one of \code{"con"} (continuous), \code{"bin"} (binary), \code{"ter"} (ternary) or \code{"tru"} (truncated). If the vector has length 1, then all p variables are assumed to be of the same type that is supplied. The default value is \code{NULL}, and the variable types are determined automatically using function \code{\link{get_types}}. As automatic determination of variable types takes extra time, it is recommended to supply the types explicitly when they are known in advance.
 #' @param method The calculation method for latent correlations. Either \code{"original"} or \code{"approx"}. If \code{method = "approx"}, multilinear approximation method is used, which is much faster than the original method, see Yoon et al. (2021) for timing comparisons for various variable types. If \code{method = "original"}, optimization of the bridge inverse function is used. The default is \code{"approx"}.
+#' @param use.nearPD Logical indicator. \code{use.nearPD = TRUE} checks semi-positive definiteness of latent correlation matrix, get nearest positive definite matrix with shrinkage adjustment by \code{nu}. output \code{R} is the same as \code{Rpointwise} if \code{use.nearPD = FALSE}. Default value is TRUE.
 #' @param nu Shrinkage parameter for the correlation matrix, must be between 0 and 1. Guarantees that the minimal eigenvalue of returned correlation matrix is greater or equal to \code{nu}. When \code{nu = 0}, no shrinkage is performed, the returned correlation matrix will be semi-positive definite but not necessarily strictly positive definite. When \code{nu = 1}, the identity matrix is returned (not recommended).  The default (recommended) value is 0.001.
 #' @param tol When \code{method = "original"}, specifies the desired accuracy of the bridge function inversion via uniroot optimization and is passed to \code{\link{optimize}}. The default value is 1e-8. When \code{method = "approx"}, this parameter is ignored.
[…]

### 2021-10-26 R/get_types.R
Minor correction to make get_types work for NAs.

@@ -23,7 +23,8 @@ get_types = function(X, tru_prop = 0.05) {
   # Determine the type of each variable
   for (i in 1:p) {
     # Extract the number of unique levels
-    levels = unique(X[ , i])
+    x = X[ , i]
+    levels = unique(x[!(is.na(x))])
     if (length(levels) <= 1) {
       # Only one level means no variation
       stop("No variation in ", i, "th variable (", i, "th column of input data).")
@@ -39,7 +40,7 @@ get_types = function(X, tru_prop = 0.05) {
         # Since ordinals of more than 3 levels are not supported, alert the user those are treated as continuous
         message("ordinal levels between 4 and 10 will be approximated by either continuous or truncated type.")
       }
-      if ((min(X[ , i], na.rm = TRUE) == 0) & (mean(X[ , i] == 0, na.rm = TRUE) > tru_prop)) {
+      if ((min(x, na.rm = TRUE) == 0) & (mean(x == 0, na.rm = TRUE) > tru_prop)) {
         # If the minimal value is zero and there are at least tru_prop zeros -> truncated (zero-inflated)
         types[i] = "tru"
       } else {

### 2022-05-21 R/internal.R
Temporarily disable approx due to chebpol depricated.

@@ -312,13 +312,14 @@ r_sol = function(K, zratio1, zratio2, comb, tol, ratio) {
 }
 
 r_ml = function(K, zratio1, zratio2, comb, tol, ratio) {
-  ipol_switch = get(paste("ipol", comb, sep = "_"))
+#  ipol_switch = get(paste("ipol", comb, sep = "_"))
   zratio1 = matrix(zratio1, ncol = length(K)); zratio1.row = nrow(zratio1)
   zratio2 = matrix(zratio2, ncol = length(K)); zratio2.row = nrow(zratio2)
   if (zratio1.row > 1) {zratio1[1:(zratio1.row - 1), ] = zratio1[1:(zratio1.row - 1), ] / zratio1[2:zratio1.row, ]}
   if (zratio2.row > 1) {zratio2[1:(zratio2.row - 1), ] = zratio2[1:(zratio2.row - 1), ] / zratio2[2:zratio2.row, ]}
   if (any(is.na(zratio2))) {zratio2 = NULL}
-  out = ipol_switch(rbind(K, zratio1, zratio2)) / 10^7
+#  out = ipol_switch(rbind(K, zratio1, zratio2)) / 10^7
+  out <- NULL
   return(out)
 }
 

### 2022-05-21 R/latentcor.R

@@ -44,11 +44,10 @@
 #' @importFrom heatmaply heatmaply
 #' @importFrom Matrix nearPD
 #' @importFrom pcaPP cor.fk
-#' @importFrom chebpol ipol
 #' @export
 #' @example man/examples/latentcor_ex.R
 
-latentcor = function(X, types = NULL, method = c("approx", "original"), use.nearPD = TRUE, nu = 0.001, tol = 1e-8, ratio = 0.9, showplot = FALSE){
+latentcor = function(X, types = NULL, method = c("original", "approx"), use.nearPD = TRUE, nu = 0.001, tol = 1e-8, ratio = 0.9, showplot = FALSE){
   # Check the supplied parameters are compatible with what is expected
   if(nu < 0 | nu > 1){
     stop("nu must be between 0 and 1.")

### 2022-05-21 man/examples/latentcor_ex.R

@@ -5,17 +5,17 @@ X = gen_data(n = 300, types = rep("tru", 5))$X
 start_time = proc.time()
 R_nc_org = latentcor(X = X, types = "tru", method = "original")$R
 proc.time() - start_time
-# Estimate latent correlation matrix with approximation method and check the timing
-start_time = proc.time()
-R_nc_approx = latentcor(X = X, types = "tru", method = "approx")$R
-proc.time() - start_time
-# Heatmap for latent correlation matrix.
-Heatmap_R_nc_approx = latentcor(X = X, types = "tru", method = "approx",
-                                showplot = TRUE)$plotR
+# # Estimate latent correlation matrix with approximation method and check the timing
+# start_time = proc.time()
+# R_nc_approx = latentcor(X = X, types = "tru", method = "approx")$R
+# proc.time() - start_time
+# # Heatmap for latent correlation matrix.
+# Heatmap_R_nc_approx = latentcor(X = X, types = "tru", method = "approx",
+#                                 showplot = TRUE)$plotR
 
 # Example 2 - ternary/continuous case
 X = gen_data()$X
 # Estimate latent correlation matrix with original method
 R_nc_org = latentcor(X = X, types = c("ter", "con"), method = "original")$R
-# Estimate latent correlation matrix with aprroximation method
-R_nc_approx = latentcor(X = X, types = c("ter", "con"), method = "approx")$R
+# # Estimate latent correlation matrix with aprroximation method
+# R_nc_approx = latentcor(X = X, types = c("ter", "con"), method = "approx")$R

### 2022-05-21 tests/testthat/test_latentcor.R

@@ -20,14 +20,14 @@ test_that("latentcor is symmetric.", {
                latentcor(X = X, method = "original", nu = 0.5, tol = 1e-8, ratio = .9, showplot = TRUE)$R)
 })
 
-test_that("latentcor is symmetric.", {
-  X = gen_data(types = c("con", "con"))$X; colnames(X) = c("var1", "var2")
-  expect_equal(latentcor(X = X, types = c("con", "con"), method = "approx", nu = 0.5, tol = 1e-8, ratio = .9, showplot = TRUE)$R,
-               t(latentcor(X = X, types = c("con", "con"), method = "approx", nu = 0.5, tol = 1e-8, ratio = .9, showplot = FALSE)$R))
-  expect_equal(latentcor(X = X, types = c("con", "con"), method = "approx", nu = 0.5, tol = 1e-8, ratio = .9, showplot = TRUE)$R,
-               latentcor(X = X, method = "approx", nu = 0.5, tol = 1e-8, ratio = .9, showplot = TRUE)$R)
-
-})
+# test_that("latentcor is symmetric.", {
+#   X = gen_data(types = c("con", "con"))$X; colnames(X) = c("var1", "var2")
+#   expect_equal(latentcor(X = X, types = c("con", "con"), method = "approx", nu = 0.5, tol = 1e-8, ratio = .9, showplot = TRUE)$R,
+#                t(latentcor(X = X, types = c("con", "con"), method = "approx", nu = 0.5, tol = 1e-8, ratio = .9, showplot = FALSE)$R))
+#   expect_equal(latentcor(X = X, types = c("con", "con"), method = "approx", nu = 0.5, tol = 1e-8, ratio = .9, showplot = TRUE)$R,
+#                latentcor(X = X, method = "approx", nu = 0.5, tol = 1e-8, ratio = .9, showplot = TRUE)$R)
+#
+# })
 
 test_that("latentcor is symmetric.", {
   X = gen_data(types = c("bin", "bin"), rhos = 1)$X
@@ -37,13 +37,13 @@ test_that("latentcor is symmetric.", {
                latentcor(X = X, method = "original", nu = 0.5, tol = 1e-8, ratio = .9, showplot = TRUE)$R)
 })
 
-test_that("latentcor is symmetric.", {
-  X = gen_data(types = c("bin", "bin"))$X; colnames(X) = c("var1", "var2")
[…]

### 2022-05-21 tests/testthat/test_ml.R

@@ -1,59 +1,0 @@
-
-library(latentcor)
-
-K = runif(1, -.5, .5)
-zratio1 = .5; zratio2 = .5
-
-test_that("r should not exceed one.", {
-  expect_lte(r_ml(K = K, zratio1 = zratio1, zratio2 = NA, comb = "10", tol = 1e-8, ratio = .9), 1)
-  expect_lte(r_ml(K = K, zratio1 = zratio1, zratio2 = zratio2, comb = "11", tol = 1e-8, ratio = .9), 1)
-  expect_lte(r_ml(K = K, zratio1 = zratio1, zratio2 = NA, comb = "20", tol = 1e-8, ratio = .9), 1)
-  expect_lte(r_ml(K = K, zratio1 = zratio1, zratio2 = zratio2, comb = "21", tol = 1e-8, ratio = .9), 1)
-  expect_lte(r_ml(K = K, zratio1 = zratio1, zratio2 = zratio2, comb = "22", tol = 1e-8, ratio = .9), 1)
-})
-
-zratio1 = matrix(c(.3, .7), ncol = 1)
-
-test_that("r should not exceed one.", {
-  expect_lte(r_ml(K = K, zratio1 = zratio1, zratio2 = NA, comb = "30", tol = 1e-8, ratio = .9), 1)
-  expect_lte(r_ml(K = K, zratio1 = zratio1, zratio2 = zratio2, comb = "31", tol = 1e-8, ratio = .9), 1)
-  expect_lte(r_ml(K = K, zratio1 = zratio1, zratio2 = zratio2, comb = "32", tol = 1e-8, ratio = .9), 1)
-})
-
-zratio2 = matrix(c(.2, .8), ncol = 1)
-test_that("r should not exceed one.", {
-  expect_lte(r_ml(K = K, zratio1 = zratio1, zratio2 = zratio2, comb = "33", tol = 1e-8, ratio = .9), 1)
-})
-
-K = sort(c(runif(1, -1, 0), runif(1, 0, 1)))
-zratio1 = .5; zratio2 = .5
-
-test_that("r is increasing in tau.", {
-  expect_lte(r_ml(K = K[1], zratio1 = zratio1, zratio2 = NA, comb = "10", tol = 1e-8, ratio = .9),
-             r_ml(K = K[2], zratio1 = zratio1, zratio2 = NA, comb = "10", tol = 1e-8, ratio = .9))
[…]

### 2022-07-14 R/interpolant.R
Put ipol into latentcor

@@ -1,0 +1,18 @@
+#' Evaluate an interpolant in a point
+#'
+#' An interpolant is a function returned by \code{\link{ipol}} which has prespecified values in some
+#' points, and which fills in between with some reasonable values.
+#'
+#' @name interpolant
+#' @param x The argument of the function. A function of more then one variable takes a
+#' vector. \code{x} can also be a matrix of column vectors.
+#' @param threads The number of threads to use for evaluation. All  interpolants created by
+#' \code{ipol} are parallelized. If given a matrix argument \code{x}, the vectors can
+#' be evaluated in parallel.
+#' @param ... Other parameters. Currently used for simplex linear interpolants with the logical argument
+#' \code{epol} which makes the interpolant extrapolate to points outside the domain.
+#' The stalker spline has the argument \code{blend=c("linear","cubic","sigmoid")} where a
+#' blending function can be chosen as described in a vignette. The \code{"multilinear"} interpolant also
+#' has such a blending function.
+#' @return A numeric. If more than one point was evaluated, a vector.
+NULL

### 2022-07-14 R/ipol.R

@@ -1,0 +1,81 @@
+#' Methods for creating multivariate interpolations on hypercubes.
+#' 
+#' The primary method is \code{\link{ipol}} which
+#' dispatches to some other method.  All the generated
+#' \link{interpolant}s accept as an argument a matrix of column
+#' vectors. The generated functions also accept an argument
+#' \code{threads=getOption('ipol.threads')} to utilize more than
+#' one CPU if a matrix of column vectors is evaluated.  The option
+#' \code{ipol.threads} is initialized from the environment variable
+#' \code{IPOL_THREADS} upon loading of the package. It defaults to \code{1}.
+#'
+#' The interpolants are ordinary R-objects and can be saved with \code{save()} and loaded
+#' later with \code{load()} or serialized/unserialized with other tools, just like any R-object.
+#' However, they contain calls to functions in the package, and while the author will make efforts
+#' to ensure that generated interpolants are compatible with future versions of \pkg{ipol},
+#' I can issue no such absolute guarantee.
+#'
+#' 
+#' @name interpolation
+#' @aliases interpolation
+#' @seealso \link{ipol}, \link{interpolant}
+#' @useDynLib ipol, .registration=TRUE, .fixes='C_'
+#'
+NULL
+#' 
+#' Create an interpolating function from given values. Several interpolation methods are
+#' supported.
+#' 
+#' @param val array or function. Function values on a grid.
+#' @param grid list. Each element is a vector of ordered grid-points for a
+#' dimension.
+#' @param ... Further arguments to the function, if \code{is.function(val)}. And some
+#' extra arguments for interpolant creation described in section Details.
+#' @return A \code{function(x, threads=getOption('chebpol.threads'))} defined on a hypercube, an \link{interpolant}
+#' for the given function. The argument \code{x} can be a matrix of column
+#' vectors which are evaluated in parallel in a number of threads.  The
+#' function yields values for arguments outside the hypercube as well, though
+#' it will typically be a poor approximation.  \code{threads} is an integer
+#' specifying the number of parallel threads which should be used when
+#' evaluating a matrix of column vectors.
+#' @import stats geometry
+#' @examples
+#' 
+#' ## evenly spaced grid-points
[…]

### 2022-07-14 R/multilinear.R

@@ -1,0 +1,40 @@
+#' Multilinear interpolation on a grid
+#' 
+#' Multilinear interpolation on an arbitrary Cartesian product.
+#' 
+#' A call \code{fun <- mlappx(val,grid)} creates a multilinear interpolant on
+#' the grid.  The value on the grid points will be exact, the value between the
+#' grid points is a convex combination of the values in the corners of the
+#' hypercube surrounding it. 
+#' 
+#' If \code{val} is a function it will be evaluated on the grid.
+#' 
+#' @aliases mlappx mlappxf
+#' @param val Array or function. Function values on a grid, or the function
+#' itself. If it is the values, the \code{dim}-attribute must be appropriately
+#' set.
+#' @param grid A list.  Each element is a vector of ordered grid-points for a
+#' dimension.  These need not be Chebyshev-knots, nor evenly spaced.
+#' @param ... Further arguments to the function, if \code{is.function(val)}.
+#' @return A \code{function(x)} defined on the hypercube, approximating the
+#' given function.  The function yields values for arguments outside the
+#' hypercube as well, as a linear extension.
+#' @keywords internal
+mlappx.real <- function(val, grid, ...) {
+  if(is.numeric(grid)) grid <- list(grid)
+  if(any(sapply(grid,is.unsorted))) {
+    if(!is.function(val)) stop('Grid points must be ordered in increasing order')
+    grid <- lapply(grid,sort)
+  }
+  gl <- prod(sapply(grid,length))
+  if(length(val) != gl)
+    stop("length of values ",length(val)," do not match size of grid ",gl)
+  val <- as.numeric(val)
+  vectorfun(function(x,threads=getOption('ipol.threads'),
+                     blend=c('linear','cubic','sigmoid','parodic','square')) {
+    blend <- switch(match.arg(blend),linear=0L,sigmoid=1L,parodic=2L,cubic=3L,square=4L)
+    .Call(C_evalmlip,grid,val,x,threads,blend)
+  },
+  arity=length(grid), 
+  domain=lapply(grid,range))
+}

### 2022-07-14 R/tools.R

@@ -1,0 +1,65 @@
+.onLoad <- function(libname,pkgname) {
+  if(is.na(thr <- as.integer(Sys.getenv('IPOL_THREADS')))) thr <- 1L
+  options(ipol.threads=thr)
+}
+
+vectorfun <- function(fun,arity,domain=NULL) {
+  force(arity)
+  f <- function(x) {
+    mc <- match.call(expand.dots=TRUE)
+    mc[[1L]] <- quote(list)
+    arglist <- eval.parent(mc)
+    x <- arglist[[1]]
+    if(is.matrix(x) || length(x) == arity) return(do.call(fun,arglist))
+    if(arity == 1) {arglist[[1]] <- matrix(x,1); return(do.call(fun,arglist))}
+    # try to coerce argument to matrix
+    if(length(x) %% arity == 0) {
+      numvec <- length(x) / arity
+      if(numvec > 1)
+        warning(sprintf('Coercing vector of length %d into %dx%d matrix',length(x),arity,numvec))
+      dim(x) <- c(arity,numvec)
+      arglist[[1]] <- x
+      do.call(fun,arglist)
+    } else
+      stop(sprintf('Function should take %d arguments, you supplied a vector of length %d',arity,length(x)))
+  }
+  formals(f) <- formals(fun)
+  structure(f,arity=arity,domain=as.data.frame(domain),
+            ipol.version=utils::packageVersion('ipol'))
+}
+
+defaultblend <- c('linear','cubic','sigmoid','parodic','square','mean')
+blenddef <- function(fun,blend=defaultblend) {
+  if(is.null(blend)) return(fun)
+  blend <- match.arg(blend)
+  pos <- match(blend,defaultblend)
+  f <- formals(fun)
+  if(!('blend' %in% names(f))) return(f)
+  f[['blend']] <- as.call(c(list(as.name('c')),c(defaultblend[pos],defaultblend[-pos])))
+  formals(fun) <- f
+  ffun <- get('fun',environment(fun))
+  ff <- formals(ffun)
+  if(!('blend' %in% names(ff))) return(fun)
+  ff[['blend']] <- f[['blend']]
+  formals(ffun) <- ff
+  assign('fun',ffun,envir=environment(fun))
+  fun
+}
+
+separgs <- function(fun,...) {
+  fun <- match.fun(fun)
+  f <- function() {
+    mc <- match.call()
[…]

### 2022-07-14 src/ipol.c

@@ -1,0 +1,263 @@
+#include "ipol.h"
+#define UNUSED(x) (void)(x)
+
+
+// return index of the smallest element larger or equal to val
+// binary search is slower than linear for small n, but we don't optimize that here
+// small grids anyway run faster
+// we never return 0, if val < arr[0] we return 1
+// if val > arr[n-1], we return n-1
+// this is suited for our particular application
+static inline int binsearch(const int n,double *arr, const double val) {
+  int toosmall, toolarge, test;
+  toosmall = 0;
+  toolarge = n-1;
+  while(toosmall+1 < toolarge) {
+    test = (toosmall + toolarge) >> 1;
+    if(arr[test] >= val) {
+      toolarge = test;
+    } else if(arr[test] < val) {
+      toosmall = test;
+    } 
+  }
+  return toolarge;
+}
+
+static void C_predmlip(const int rank, double **grid, int *dims, double *values, double *output) {
+  int stride[rank+1];
+  stride[0] = 1;
+  for(int i = 1; i <= rank; ++i) stride[i] = dims[i-1]*stride[i-1];
+  int N = stride[rank];
+  int dim[rank];
+  for(int i = 0; i < rank; ++i) {
+    dim[i] = 0;
+  }
+  /* Loop over entire grid. For each point, predict it as the mean of all its immediate neighbours. */
+  for(int j = 0; j < N; j++,dim[0]++) {
+    for(int i = 0; i < rank-1; i++) {
+      if(dim[i] >= dims[i]) {
+	dim[i] -= dims[i];
+	dim[i+1]++;
+      } else 
+	break;
+    }
+    //    printf("j %d: ",j);for(int i = 0; i < rank; i++) printf("%d ",dim[i]); printf("\n");
+    double pred = 0;
+    int w = 0;
+    for(int i = 0; i < rank; i++) {
+      if(dim[i] <= 0 || dim[i]+1 >= dims[i]) continue;
+      double *gr = grid[i];
+      double r = (gr[dim[i]] - gr[dim[i]-1])/(gr[dim[i]+1] - gr[dim[i]-1]);
+      w++;
+      pred += r*values[j+stride[i]] + (1-r)*values[j-stride[i]];
+    }
+    if(w == 0)
+      output[j] = values[j];
+    else
[…]

### 2022-07-14 src/ipol.h

@@ -1,0 +1,35 @@
+#ifndef _IPOL
+#define _IPOL
+#include <math.h>
+#include <Rmath.h>
+#include <R.h>
+#include <Rdefines.h>
+#include <R_ext/Rdynload.h>
+#include <R_ext/BLAS.h>
+#include <R_ext/Lapack.h>
+#include <R_ext/Applic.h>
+#include <R_ext/Visibility.h>
+
+static R_INLINE double blendfun(double w,int blend) {
+  switch(blend) {
+  case 0:
+    // Linear, nothing to do here
+    break;
+  case 1:
+    w = w<0.5 ? 0.5*exp(2-1/w) : 1-0.5*exp(2-1/(1-w));  // smooth sigmoid blending
+    break;
+  case 2:
+    w = w<0.5 ? 0.5*exp(4-1/(w*w)) : 1-0.5*exp(4-1/((1-w)*(1-w)));  // parodic sigmoid blending
+    break;
+  case 3:
+    w = (-2*w + 3)*w*w; // cubic blending
+    break;
+  case 4:
+    w = (w < 0.5) ? 0 : 1;  // discontinuous blending
+  break;
+  case 5:
+    w = 0.5; // mean blending
+  }
+  return w;
+}
+#endif

### 2022-07-14 R/ipol.R
Update package name to latentcor for all new added files.

@@ -1,5 +1,5 @@
 #' Methods for creating multivariate interpolations on hypercubes.
-#' 
+#'
 #' The primary method is \code{\link{ipol}} which
 #' dispatches to some other method.  All the generated
 #' \link{interpolant}s accept as an argument a matrix of column
@@ -15,17 +15,17 @@
 #' to ensure that generated interpolants are compatible with future versions of \pkg{ipol},
 #' I can issue no such absolute guarantee.
 #'
-#' 
+#'
 #' @name interpolation
 #' @aliases interpolation
 #' @seealso \link{ipol}, \link{interpolant}
-#' @useDynLib ipol, .registration=TRUE, .fixes='C_'
+#' @useDynLib latentcor, .registration=TRUE, .fixes='C_'
 #'
 NULL
-#' 
+#'
 #' Create an interpolating function from given values. Several interpolation methods are
 #' supported.
-#' 
+#'
 #' @param val array or function. Function values on a grid.
 #' @param grid list. Each element is a vector of ordered grid-points for a
 #' dimension.
@@ -40,7 +40,7 @@ NULL
 #' evaluating a matrix of column vectors.
 #' @import stats geometry
 #' @examples
-#' 
+#'
 #' ## evenly spaced grid-points
 #' su <- seq(0,1,length.out=10)
 #' ## irregularly spaced grid-points
@@ -50,18 +50,18 @@ NULL
 #' ## test it, since exp is convex, the linear approximation lies above
 #' ## the exp between the grid points
 #' ml1(su) - exp(su)
-#' 
+#'
 #' ## multi dimensional approximation
 #' f <- function(x) 10/(1+25*mean(x^2))
 #' # a 3-dimensional 10x10x10 grid, first and third coordinate are non-uniform
 #' grid <- list(s, su, sort(1-s))
-#' 
+#'
 #' # make multilinear spline.
 #' ml2 <- ipol(array(apply(expand.grid(grid), 1, f), c(10, 10, 10)), grid=grid)
 #' # make 7 points in R3 to test them on
 #' m <- matrix(runif(3*7),3)
 #' rbind(true=apply(m,2,f), ml=ml2(m))
-#' 
+#'
 #' @export
 ipol <- function(val, grid=NULL, ...) {
   args <- list(...)

### 2022-07-14 src/ipol.c

@@ -18,7 +18,7 @@ static inline int binsearch(const int n,double *arr, const double val) {
       toolarge = test;
     } else if(arr[test] < val) {
       toosmall = test;
-    } 
+    }
   }
   return toolarge;
 }
@@ -38,7 +38,7 @@ static void C_predmlip(const int rank, double **grid, int *dims, double *values,
       if(dim[i] >= dims[i]) {
 	dim[i] -= dims[i];
 	dim[i+1]++;
-      } else 
+      } else
 	break;
     }
     //    printf("j %d: ",j);for(int i = 0; i < rank; i++) printf("%d ",dim[i]); printf("\n");
@@ -82,7 +82,7 @@ static SEXP R_mlippred(SEXP sgrid, SEXP values) {
   return resvec;
 }
 
-static double C_evalmlip(const int rank, double *x, double **grid, int *dims, 
+static double C_evalmlip(const int rank, double *x, double **grid, int *dims,
 			 double *values, int blend) {
 
   double weight[rank];
@@ -143,9 +143,9 @@ static SEXP R_evalmlip(SEXP sgrid, SEXP values, SEXP x, SEXP Rthreads, SEXP Sble
   double *grid[rank];
   int blend = 0;
   if(!IS_NUMERIC(values)) error("values must be numeric");
-  if(!IS_NUMERIC(x)) error("argument x must be numeric");  
+  if(!IS_NUMERIC(x)) error("argument x must be numeric");
   if(isMatrix(x) ? (nrows(x) != rank) : (LENGTH(x) != rank))
-    error("grid has dimension %d, you supplied a length %d vector", 
+    error("grid has dimension %d, you supplied a length %d vector",
 	  rank, isMatrix(x) ? nrows(x) : LENGTH(x));
 
   /* Make some pointers into the grid data */
@@ -254,10 +254,10 @@ R_CallMethodDef callMethods[] = {
 };
 
 
-void attribute_visible R_init_ipol(DllInfo *info) {
+void attribute_visible R_init_latentcor(DllInfo *info) {
   if(info != NULL) {}; // avoid warning about unused parameter
   /* register our routines */
   R_registerRoutines(info,NULL,callMethods,NULL,NULL);
   R_useDynamicSymbols(info, FALSE);
-  R_RegisterCCallable("ipol", "evalmlip", (DL_FUNC) C_evalmlip);
+  R_RegisterCCallable("latentcor", "evalmlip", (DL_FUNC) C_evalmlip);
 }

### 2022-08-05 supplement/ex_mtcars.R
Add some supplement

@@ -1,0 +1,50 @@
+# Script for example of mtcars
+library(ggplot2)
+mtcars$cyl = as.character(mtcars$cyl)
+mtcars$vs = as.character(mtcars$vs)
+mtcars$am = as.character(mtcars$am)
+hist_mpg = ggplot(mtcars, aes(x=mpg)) + geom_histogram(aes(y=..density..), binwidth = 2, color="black", fill="white")+geom_density(alpha=.2, fill="#FF6666")
+hist_cyl = ggplot(mtcars, aes(cyl)) + geom_bar(color="black", fill="white")
+hist_disp = ggplot(mtcars, aes(x=disp)) + geom_histogram(aes(y=..density..), binwidth = 40, color="black", fill="white")+geom_density(alpha=.2, fill="#FF6666")
+hist_hp = ggplot(mtcars, aes(x=hp)) + geom_histogram(aes(y=..density..), binwidth = 20, color="black", fill="white")+geom_density(alpha=.2, fill="#FF6666")
+hist_drat = ggplot(mtcars, aes(x=drat)) + geom_histogram(aes(y=..density..), binwidth = .2, color="black", fill="white")+geom_density(alpha=.2, fill="#FF6666")
+hist_wt = ggplot(mtcars, aes(x=wt)) + geom_histogram(aes(y=..density..), binwidth = .5, color="black", fill="white")+geom_density(alpha=.2, fill="#FF6666")
+hist_qsec = ggplot(mtcars, aes(x=qsec)) + geom_histogram(aes(y=..density..), binwidth = 1, color="black", fill="white")+geom_density(alpha=.2, fill="#FF6666")
+hist_vs = ggplot(mtcars, aes(vs)) + geom_bar(color="black", fill="white")
+hist_am = ggplot(mtcars, aes(am)) + geom_bar(color="black", fill="white")
+hist_gear = ggplot(mtcars, aes(gear)) + geom_bar(color="black", fill="white")
+hist_carb = ggplot(mtcars, aes(x=carb)) + geom_histogram(aes(y=..density..), binwidth = 1, color="black", fill="white")+geom_density(alpha=.2, fill="#FF6666")
+plot_all_1 = gridExtra::marrangeGrob(grobs = list(hist_mpg, hist_cyl, hist_disp, hist_hp, hist_drat, hist_wt), widths = rep(1, 2), top = NULL, layout_matrix = matrix(c(1:6), 3, 2, byrow = T))
[…]

### 2022-08-05 supplement/plot_eval.R

@@ -2,51 +2,96 @@ library(heatmaply)
 library(ggplot2)
 library(plotly)
 library(graphics)
+library(gridExtra)
+library(png)
+library(grid)
 load("~/latentcor/latentcor/supplement/all_evaluation.rda")
 grid_list = list(LatentR = seq(-0.9, 0.9, by = 0.1), TruncRate = seq(0.1, 0.9, by = 0.1))
 names(grid_list)=c("Latent Correlation", "Proportion of Zeros")
 eval_data = evaluation_BC$meanAE_diff
 rownames(eval_data) = as.character(grid_list[[1]]); colnames(eval_data) = as.character(grid_list[[2]])
-heatmaply(eval_data, dendrogram = "none", main = "Mean Absolute Approximation Error for Binary/Continuous", margins = c(80,80,80,80), ylab = names(grid_list[1]), xlab = names(grid_list[2]),
+heatmaply(eval_data, dendrogram = "none", main = "binary vs. continuous (hybrid)", margins = c(80,80,80,80), ylab = names(grid_list[1]), xlab = names(grid_list[2]),
           grid_color = "white", grid_width = 0.00001, label_names = c(paste0(names(grid_list[2]), ":"), paste0(names(grid_list[1]), ":"), "Mean Absolute Approximation Error:"))
 
 eval_data = evaluation_BB$meanAE_diff
 rownames(eval_data) = as.character(grid_list[[1]]); colnames(eval_data) = as.character(grid_list[[2]])
-heatmaply(eval_data, dendrogram = "none", main = "Mean Absolute Approximation Error for Binary/Binary", margins = c(80,80,80,80), ylab = names(grid_list[1]), xlab = names(grid_list[2]),
+heatmaply(eval_data, dendrogram = "none", main = "binary vs. binary (hybrid)", margins = c(80,80,80,80), ylab = names(grid_list[1]), xlab = names(grid_list[2]),
           grid_color = "white", grid_width = 0.00001, label_names = c(paste0(names(grid_list[2]), ":"), paste0(names(grid_list[1]), ":"), "Mean Absolute Approximation Error:"))
 
 eval_data = evaluation_TC$meanAE_diff
 rownames(eval_data) = as.character(grid_list[[1]]); colnames(eval_data) = as.character(grid_list[[2]])
[…]

### 2022-08-05 supplement/unbias_check.R

@@ -1,5 +1,6 @@
 library(latentcor)
 library(ggplot2)
+library(gridExtra)
 
 rm(list=ls())
 PlotPair = function(datapair, namepair = c("X", "Y"), title = "Plot X vs Y") {
@@ -10,67 +11,90 @@ PlotPair = function(datapair, namepair = c("X", "Y"), title = "Plot X vs Y") {
         +ggtitle(title) + xlab(namepair[1]) + ylab(namepair[2]) + xlim(-1, 1) + ylim(-1, 1))
 }
 
-
-# Data generation
-rhos = seq(-1, 1, by = .01); rhorep = rep(NA, length(rhos)); Rrep = matrix(NA, length(rhos), 3)
-for (r in 1:length(rhos)) {
-  X = GenData(n = 1000, types = c("ter", "con"), rhos = rhos[r], XP = list(c(.3, .5), NA))$X
-  R_nc_org = estR(X, types = c("ter", "con"), method = "original")$R
-  R_nc_approx = estR(X, types = c("ter", "con"), method = "approx")$R
-  rhorep[r] = rhos[r]; Rrep[r, 1] = R_nc_org[2, 1]; Rrep[r, 2] = R_nc_approx[2, 1]; Rrep[r, 3] = cor(X)[2, 1]
-}
-
-# Plot ternary/continuous case estimation via original method.
-R_nc_org = PlotPair(datapair = cbind(rhorep, Rrep[,1]), namepair = c("True latent correlation", "Estimated latent correlation (original)"),
-                    title = "Ternary vs. continuous")
-
-# Plot ternary/continuous case estimation via approximation method.
-R_nc_approx = PlotPair(datapair = cbind(rhorep, Rrep[,2]), namepair = c("True latent correlation", "Estimated latent correlation (approx)"),
-                       title = "Ternary vs. continuous")
-
-R_nc_pearson = PlotPair(datapair = cbind(rhorep, Rrep[,3]), namepair = c("True latent correlation", "Pearson correlation"),
-                        title = "Ternary vs. continuous")
-#R_nc = gridExtra::marrangeGrob(grobs = list(R_nc_org, R_nc_approx, R_nc_pearson), widths = c(10, 2, 10, 2, 10), top = NULL, layout_matrix = matrix(c(1, NA, 2, NA, 3), 1, 5))
-pdf(file = "nc_org.pdf", width = 5, height = 5)
-R_nc_org
-dev.off()
[…]

### 2022-08-06 man/examples/latentcor_ex.R
Fixed bugs

@@ -3,19 +3,19 @@
 X = gen_data(n = 300, types = rep("tru", 5))$X
 # Estimate latent correlation matrix with original method and check the timing
 start_time = proc.time()
-R_nc_org = latentcor(X = X, types = "tru", method = "original")$R
+R_org = latentcor(X = X, types = "tru", method = "original")$R
 proc.time() - start_time
-# # Estimate latent correlation matrix with approximation method and check the timing
-# start_time = proc.time()
-# R_nc_approx = latentcor(X = X, types = "tru", method = "approx")$R
-# proc.time() - start_time
-# # Heatmap for latent correlation matrix.
-# Heatmap_R_nc_approx = latentcor(X = X, types = "tru", method = "approx",
-#                                 showplot = TRUE)$plotR
+# Estimate latent correlation matrix with approximation method and check the timing
+start_time = proc.time()
+R_approx = latentcor(X = X, types = "tru", method = "approx")$R
+proc.time() - start_time
+# Heatmap for latent correlation matrix.
+Heatmap_R_approx = latentcor(X = X, types = "tru", method = "approx",
+                             showplot = TRUE)$plotR
 
 # Example 2 - ternary/continuous case
 X = gen_data()$X
 # Estimate latent correlation matrix with original method
 R_nc_org = latentcor(X = X, types = c("ter", "con"), method = "original")$R
-# # Estimate latent correlation matrix with aprroximation method
-# R_nc_approx = latentcor(X = X, types = c("ter", "con"), method = "approx")$R
+# Estimate latent correlation matrix with aprroximation method
+R_nc_approx = latentcor(X = X, types = c("ter", "con"), method = "approx")$R

### 2022-08-06 tests/testthat/test_latentcor.R

@@ -14,36 +14,36 @@ test_that("n_x is not 1 if tie", {
 # Data generation
 test_that("latentcor is symmetric.", {
   X = gen_data(types = c("con", "con"), rhos = 1)$X
-  expect_equal(latentcor(X = X, types = c("con", "con"), method = "original", nu = 0.5, tol = 1e-8, ratio = .9, showplot = TRUE)$R,
+  expect_equal(latentcor(X = X, types = c("con", "con"), method = "original", nu = 0.5, tol = 1e-8, ratio = .9, showplot = FALSE)$R,
                t(latentcor(X = X, types = c("con", "con"), method = "original", nu = 0.5, tol = 1e-8, ratio = .9, showplot = FALSE)$R))
-  expect_equal(latentcor(X = X, types = c("con", "con"), method = "original", nu = 0.5, tol = 1e-8, ratio = .9, showplot = TRUE)$R,
-               latentcor(X = X, method = "original", nu = 0.5, tol = 1e-8, ratio = .9, showplot = TRUE)$R)
+  expect_equal(latentcor(X = X, types = c("con", "con"), method = "original", nu = 0.5, tol = 1e-8, ratio = .9, showplot = FALSE)$R,
+               latentcor(X = X, method = "original", nu = 0.5, tol = 1e-8, ratio = .9, showplot = FALSE)$R)
 })
 
-# test_that("latentcor is symmetric.", {
-#   X = gen_data(types = c("con", "con"))$X; colnames(X) = c("var1", "var2")
-#   expect_equal(latentcor(X = X, types = c("con", "con"), method = "approx", nu = 0.5, tol = 1e-8, ratio = .9, showplot = TRUE)$R,
-#                t(latentcor(X = X, types = c("con", "con"), method = "approx", nu = 0.5, tol = 1e-8, ratio = .9, showplot = FALSE)$R))
-#   expect_equal(latentcor(X = X, types = c("con", "con"), method = "approx", nu = 0.5, tol = 1e-8, ratio = .9, showplot = TRUE)$R,
-#                latentcor(X = X, method = "approx", nu = 0.5, tol = 1e-8, ratio = .9, showplot = TRUE)$R)
-#
-# })
+test_that("latentcor is symmetric.", {
+  X = gen_data(types = c("con", "con"))$X; colnames(X) = c("var1", "var2")
[…]

### 2022-08-31 supplement/unbias_check.R
minor correction

@@ -21,7 +21,7 @@ for (tp1 in c("con", "bin", "ter", "tru")) {
       R_org = latentcor(X, types = c(tp1, tp2), method = "original")$R
       R_approx = latentcor(X, types = c(tp1, tp2), method = "approx")$R
       R_conapprox = latentcor(X, types = c(ifelse(tp1 == "ter", "con", tp1), ifelse(tp2 == "ter", "con", tp2)), method = "original")$R
-      rhorep[rep] = rho; Rrep[rep, 1] = R_org[2, 1]; Rrep[rep, 2] = R_approx[2, 1]; Rrep[rep, 3] = cor(X)[2, 1]; Rrep[rep, 4] = R_conapprox[2, 1]
+      Rrep[rep, 1] = R_org[2, 1]; Rrep[rep, 2] = R_approx[2, 1]; Rrep[rep, 3] = cor(X)[2, 1]; Rrep[rep, 4] = R_conapprox[2, 1]
     }
     assign(paste("R", cp1, cp2, tp1, tp2, "org", sep = "_"), Rrep[ , 1])
     assign(paste("R", cp1, cp2, tp1, tp2, "approx", sep = "_"), Rrep[ , 2])
